---
title: "210806_GTEX_AGE_HIST"
author: "Jennifer Fisher"
date: "5/17/2021"
output:
  html_document:
    toc: true
    theme: united
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# START 

    Goal/Purpose of operations: 
    Checking GTEx samples for cofounders

    Finished psedocode on: 
    210806

    System which operations were done on: 
    my lab mac 

    GitHub Repo:
    Sex_Bias_Adverse_Events


    Directory of operations:
    /home/rstudio

    Scripts being edited for operations: 
    NA

    Docker:
    jenfisher7/rstudio_cancer_dr 

    Data being used: 
    GTEx
      GTEx_Analysis_v8_Annotations_SubjectPhenotypesDS.txt
      GTEx_Analysis_v8_Annotations_SampleAttributesDS.txt
      downloaded from https://gtexportal.org/home/datasets on 210517

    Papers and tools:
    ggplot2 and tidyverse packages

# STEPS
### Set working directory 
```{r}
setwd("/home/rstudio")
```

### load in data

```{r}
library(recount3)
library(rlang)
library(dplyr)
library(dbplyr)
```

```{r}
human_projects <- readRDS("~/data/human_recount3_projects.rds")
```



```{r}
GTEX_proj_info <- subset(
    human_projects,
    file_source == "gtex"
) 
```


```{r}
for ( i in 1:nrow(GTEX_proj_info)){
  name <- paste(GTEX_proj_info[i,1],"rse", sep="_" )
  assign(name, create_rse(GTEX_proj_info[i,]))
}
```

```{r}
name <- paste(GTEX_proj_info[,1],"rse", sep="_" )
gtex_coldata<-  colData(get(name[1]))
for (i in 2:length(name)){
  coldata<-  colData(get(name[i]))
  gtex_coldata<- rbind(gtex_coldata, coldata)
}
```

### Analysis
```{r}
gtex_coldata<- as.data.frame(gtex_coldata)
```


```{r}
table(is.na(gtex_coldata$gtex.age))
table(gtex_coldata$gtex.age)
table(is.na(gtex_coldata$gtex.sex))
table(gtex_coldata$gtex.sex)
```
1= male 2= female 

```{r}
table(gtex_coldata$gtex.dthhrdy)
table(is.na(gtex_coldata$gtex.dthhrdy))
```
0=Ventilator Case
1=Violent and fast death
2=Fast death of natural causes
3=Intermediate death
4=Slow death

How many sample per tissue?
```{r}
as.data.frame(table(gtex_coldata$gtex.smts))
```
substructures of tissues
```{r}
as.data.frame(table(gtex_coldata$gtex.smtsd))
```



```{r}
gtex_coldata$gtex.sex<- factor( ifelse(gtex_coldata$gtex.sex == 1, "M","F" ), levels= c("M", "F"))
gtex_coldata$gtex.dthhrdy <- ifelse(gtex_coldata$gtex.dthhrdy == 0,  "Ventilator_Case", ifelse(gtex_coldata$gtex.dthhrdy == 1,  "Violent_and_Fast_Death", ifelse(gtex_coldata$gtex.dthhrdy == 2,  "Fast_Death_of_Natural_Causes",ifelse(gtex_coldata$gtex.dthhrdy == 3,  "Intermediate_Death" , ifelse(gtex_coldata$gtex.dthhrdy== 4,  "Slow_Death", "Not Reported")))))
```

```{r}
gtex_coldata$SEX_AGE<- paste(gtex_coldata$gtex.sex, gtex_coldata$gtex.age, sep = "_")
```

```{r}
gtex_coldata$SEX_AGE_DEATH<- paste(gtex_coldata$SEX_AGE, gtex_coldata$gtex.dthhrdy ,sep = "_")
```

```{r}
table(gtex_coldata$SEX_AGE_DEATH)
```

```{r}
gtex_coldata2<- gtex_coldata[!is.na(gtex_coldata$gtex.smgebtcht),]
gtex_coldata3<- gtex_coldata2[gtex_coldata2$gtex.smgebtcht == "TruSeq.v1",]
gtex_coldata3<- gtex_coldata3[!is.na(gtex_coldata3$gtex.subjid),]
gtex_coldata4<- gtex_coldata3[!gtex_coldata3$gtex.subjid == "GTEX-11ILO",] 
gtex_coldata<- gtex_coldata4
```


```{r}
#sample_info
library(tidyverse)
```


```{r}
# for each tissue count each 
#i and J llops
tissue<- unique(gtex_coldata$gtex.smts)
sub_tissues<- unique(gtex_coldata$gtex.smtsd)
tissue_data <- paste(tissue, "data", sep= "_")
  # Sex 
  sex_table <- as.data.frame(table(gtex_coldata[c(14,6)]))
  
  #Age
  age_table <-  as.data.frame(table(gtex_coldata[c(14,7)]))
  
  #Death
  death_table <- as.data.frame(table(gtex_coldata[c(14,8)]))
  #AGE_Sex
  age_sex_table <- as.data.frame(table(gtex_coldata[c(14,199)]))
  
  age_sex_table_v2 <- as.data.frame(table(gtex_coldata[c(14,6,7)]))
  
    #AGE_Sex_Death
  age_sex_death_table <- as.data.frame(table(gtex_coldata[c(14,200)]))
  
  age_sex_death_table_v2 <- as.data.frame(table(gtex_coldata[c(14,6,7,8)]))
  
```


```{r}
library(ggplot2)
library(viridis)
ggplot(sex_table, aes(fill=gtex.sex, y=Freq, x=gtex.smts)) + 
    geom_bar(position="stack", stat="identity") +
    theme_classic()+
    xlab("Tissue") +
    ylab("Count")+
    scale_fill_viridis(discrete = T) + 
    theme(axis.text.x = element_text(angle=90, size=3.5), axis.ticks = element_blank())+
    ggtitle("Sex for GTEx tissue") 
ggsave("~/output/GTEx_plots/210806_sex_tissue_GTEx.pdf")
```

```{r}
ggplot(sex_table, aes(fill=gtex.sex, y=Freq, x=gtex.smts)) + 
    geom_bar(position="fill", stat="identity") +
    theme_classic()+
    xlab("Tissue") +
    ylab("Fraction")+
    scale_fill_viridis(discrete = T) + 
    theme(axis.text.x = element_text(angle=90, size=3.5), axis.ticks = element_blank())+
    ggtitle("Sex for GTEx tissue") 
ggsave("~/output/GTEx_plots/210806_sex_tissue_GTEx_percent.pdf")
```


```{r}
ggplot(age_table, aes(fill=gtex.age, y=Freq, x=gtex.smts)) + 
    geom_bar(position="stack", stat="identity") +
    theme_classic()+
    xlab("Tissue") +
    ylab("Count")+
    scale_fill_viridis(discrete = T) + 
    theme(axis.text.x = element_text(angle=90, size=3.5), axis.ticks = element_blank())+
    ggtitle("Age for GTEx tissue") 
ggsave("~/output/GTEx_plots/210806_age_tissue_GTEx.pdf")
```

```{r}
ggplot(age_table, aes(fill=gtex.age, y=Freq, x=gtex.smts)) + 
    geom_bar(position="fill", stat="identity") +
    theme_classic()+
    xlab("Tissue") +
    ylab("Fraction")+
    scale_fill_viridis(discrete = T) + 
    theme(axis.text.x = element_text(angle=90, size=3.5), axis.ticks = element_blank())+
    ggtitle("Age for GTEx tissue") 
ggsave("~/output/GTEx_plots/210806_age_tissue_GTEx_precent.pdf")
```


```{r}
ggplot(age_sex_table, aes(fill=SEX_AGE, y=Freq, x=gtex.smts)) + 
    geom_bar(position="stack", stat="identity") +
    theme_classic()+
    xlab("Tissue") +
    ylab("Count")+
    scale_fill_viridis(discrete = T) + 
    theme(axis.text.x = element_text(angle=90, size=3.5), axis.ticks = element_blank())+
    ggtitle("Sex and Age for GTEx tissue") 
ggsave("~/output/GTEx_plots/210806_sex_age_tissue_GTEx.pdf")
```

```{r}
ggplot(age_sex_table, aes(fill=SEX_AGE, y=Freq, x=gtex.smts)) + 
    geom_bar(position="fill", stat="identity") +
    theme_classic()+
    xlab("Tissue") +
    ylab("Fraction")+
    scale_fill_viridis(discrete = T) + 
    theme(axis.text.x = element_text(angle=90, size=3.5), axis.ticks = element_blank())+
    ggtitle("Sex and Age for GTEx tissue") 
ggsave("~/output/GTEx_plots/210806_sex_age_tissue_GTEx_precent.pdf")
```

```{r}
ggplot(death_table, aes(fill=gtex.dthhrdy, y=Freq, x=gtex.smts)) + 
    geom_bar(position="fill", stat="identity") +
    theme_classic()+
    xlab("Tissue") +
    ylab("Fraction")+
    scale_fill_viridis(discrete = T) + 
    theme(axis.text.x = element_text(angle=90, size =3.5), axis.ticks = element_blank())+
    ggtitle("Type of Death for GTEx tissue") 
ggsave("~/output/GTEx_plots/210806_death_tissue_GTEx_precent.pdf")
```



```{r}
tissue
```


```{r}
for (i in 1:length(tissue)){
  data<- age_sex_table_v2[age_sex_table_v2$gtex.smts == tissue[i] , ]
  title<- paste("Sex and Age for GTEx", tissue[i])
  plot <- ggplot(data=data, aes(x=gtex.age, y=Freq, group=gtex.sex)) +
      geom_line(aes(linetype=gtex.sex))+
      geom_point()+
      theme_classic()+
      ggtitle(title) 
  print(plot)
  nfile<- paste("~/output/GTEx_plots/210806_age_sex_GTEx_", tissue[i],".pdf",sep="")
  ggsave(nfile)
}
```


```{r}
ggplot(data=age_sex_table_v2, aes(x=gtex.age, y=Freq, group=interaction(gtex.sex, gtex.smts))) +
      geom_line(aes(color=interaction(gtex.sex, gtex.smts)))+
      geom_point()+
      theme_classic()+
      ggtitle("Sex and Age GTEx Tissues") 
ggsave("~/output/GTEx_plots/210806_sex_age_all_tissue_GTEx.pdf")
```


```{r}
    # Sex 
  sex_sub_table <- as.data.frame(table(gtex_coldata[c(15,6)]))
  
  #Age
  age_sub_table <-  as.data.frame(table(gtex_coldata[c(15,7)]))
  
  #Death
  death_sub_table <- as.data.frame(table(gtex_coldata[c(15,8)]))
  #AGE_Sex
  age_sex_sub_table <- as.data.frame(table(gtex_coldata[c(15,199)]))
  
  age_sex_sub_table_v2 <- as.data.frame(table(gtex_coldata[c(15,6,7)]))
  
    #AGE_Sex_Death
  age_sex_death_sub_table <- as.data.frame(table(gtex_coldata[c(15,200)]))
  
  age_sex_death_sub_table_v2 <- as.data.frame(table(gtex_coldata[c(15,6,7,8)]))
  
```

```{r}
ggplot(sex_sub_table, aes(fill=gtex.sex, y=Freq, x=gtex.smtsd)) + 
    geom_bar(position="stack", stat="identity") +
    theme_classic()+
    xlab("Tissue") +
    ylab("Count")+
    scale_fill_viridis(discrete = T) + 
    theme(axis.text.x = element_text(angle=90, size=3.5), axis.ticks = element_blank())+
    ggtitle("Sex for GTEx tissue") 
ggsave("~/output/210806_sex_tissue_structures_GTEx.pdf")
```

```{r}
ggplot(sex_sub_table, aes(fill=gtex.sex, y=Freq, x=gtex.smtsd)) + 
    geom_bar(position="fill", stat="identity") +
    theme_classic()+
    xlab("Tissue") +
    ylab("Fraction")+
    scale_fill_viridis(discrete = T) + 
    theme(axis.text.x = element_text(angle=90, size=3.5), axis.ticks = element_blank())+
    ggtitle("Sex for GTEx tissue") 
ggsave("~/output/GTEx_plots/210806_sex_tissue_structures_GTEx_precent.pdf")
```

```{r}
ggplot(age_sub_table, aes(fill=gtex.age, y=Freq, x=gtex.smtsd)) + 
    geom_bar(position="stack", stat="identity") +
    theme_classic()+
    xlab("Tissue") +
    ylab("Count")+
    scale_fill_viridis(discrete = T) + 
    theme(axis.text.x = element_text(angle=90, size=3.5), axis.ticks = element_blank())+
    ggtitle("Age for GTEx tissue") 
ggsave("~/output/GTEx_plots/210806_age_tissue_structures_GTEx.pdf")
```

```{r}
ggplot(age_sub_table, aes(fill=gtex.age, y=Freq, x=gtex.smtsd)) + 
    geom_bar(position="fill", stat="identity") +
    theme_classic()+
    xlab("Tissue") +
    ylab("Fraction")+
    scale_fill_viridis(discrete = T) + 
    theme(axis.text.x = element_text(angle=90, size= 3.5), axis.ticks = element_blank())+
    ggtitle("Age for GTEx tissue") 
ggsave("~/output/210806_age_tissue_structures_GTEx.pdf")
```

```{r}
ggplot(age_sex_sub_table, aes(fill=SEX_AGE, y=Freq, x=gtex.smtsd)) + 
    geom_bar(position="stack", stat="identity") +
    theme_classic()+
    xlab("Tissue") +
    ylab("Count")+
    scale_fill_viridis(discrete = T) + 
    theme(axis.text.x = element_text(angle=90, size =3.5), axis.ticks = element_blank())+
    ggtitle("Sex and Age for GTEx tissue") 
ggsave("~/output/GTEx_plots/210806_sex_age_tissue_structures_GTEx.pdf")
```

```{r}
ggplot(age_sex_sub_table, aes(fill=SEX_AGE, y=Freq, x=gtex.smtsd)) + 
    geom_bar(position="fill", stat="identity") +
    theme_classic()+
    xlab("Tissue") +
    ylab("Fraction")+
    scale_fill_viridis(discrete = T) + 
    theme(axis.text.x = element_text(angle=90, size=3.5), axis.ticks = element_blank())+
    ggtitle("Sex and Age for GTEx tissue") 
ggsave("~/output/GTEx_plots/210806_sex_age_tissue_structures_GTEx_precent.pdf")
```

```{r}
ggplot(death_sub_table, aes(fill=gtex.dthhrdy, y=Freq, x=gtex.smtsd)) + 
    geom_bar(position="fill", stat="identity") +
    theme_classic()+
    xlab("Tissue") +
    ylab("Fraction")+
    scale_fill_viridis(discrete = T) + 
    theme(axis.text.x = element_text(angle=90, size=3.5), axis.ticks = element_blank())+
    ggtitle("Type of Death for GTEx tissue") 
ggsave("~/output/GTEx_plots/210806_death_tissue_structures_GTEx_precent.pdf")
```

```{r}
ggplot(gtex_coldata, aes(x=gtex.smts, y=gtex.smrin, fill=gtex.sex)) +
  geom_boxplot(position=position_dodge(1))+
    theme_classic()+
    xlab("Tissue") +
    ylab("RIN Score")+
    scale_fill_viridis(discrete = T) + 
    theme(axis.text.x = element_text(angle=90, size=3.5), axis.ticks = element_blank())+
    ggtitle("RIN for GTEx tissue") + geom_hline(yintercept=5, linetype="dashed", color = "red")
ggsave("~/output/210806_RIN_tissue_GTEx.pdf")
```

```{r}
autolysis_score <- as.data.frame(table(gtex_coldata[c(14, 10)]))
ggplot(autolysis_score, aes(fill=gtex.smatsscr, y=Freq, x=gtex.smts)) + 
   geom_bar(position="fill", stat="identity") +
    theme_classic()+
    xlab("Tissue") +
    ylab("Fraction")+
    scale_fill_viridis(discrete = T) + 
    theme(axis.text.x = element_text(angle=90, size=3.5), axis.ticks = element_blank())+
    ggtitle("Autolysis Score for GTEx tissue") 
ggsave("~/output/GTEx_plots/210806_autolysis_score_tissue_GTEx_precent.pdf")
```

The autolysis score was assigned by a pathologist during a visual inspection of the histology image. The assigned values ranged from 0 to 3 (None, Mild, Moderate, and Severe).

```{r}
ggplot(gtex_coldata, aes(x=gtex.smts , y= gtex.smmppd , fill=gtex.sex)) +
  geom_boxplot(position=position_dodge(1))+
    theme_classic()+
    xlab("Tissue") +
    ylab("Total number of reads aligned/map")+
    scale_fill_viridis(discrete = T) + 
    theme(axis.text.x = element_text(angle=90, size=3.5), axis.ticks = element_blank())+
    ggtitle("Total number of reads aligned/map for GTEx tissue") 
ggsave("~/output/GTEx_plots/210806_ALIGNED_tissue_GTEx.pdf")
```

```{r}
ggplot(gtex_coldata, aes(x=gtex.smts, y=gtex.smtsisch , fill=gtex.sex)) +
  geom_boxplot(position=position_dodge(1))+
    theme_classic()+
    xlab("Tissue") +
    ylab("Ischemic Time")+
    scale_fill_viridis(discrete = T) + 
    theme(axis.text.x = element_text(angle=90, size=3.5), axis.ticks = element_blank())+
    ggtitle("Ischemic Time for GTEx tissue") 
ggsave("~/output/GTEx_plots/210806_Ischemic_Time_tissue_GTEx.pdf")
```


```{r}
ggplot(gtex_coldata, aes(x=gtex.smts, y= gtex.smmaprt, fill=gtex.sex)) +
  geom_boxplot(position=position_dodge(1))+
    theme_classic()+
    xlab("Tissue") +
    ylab("Mapping Rate")+
    scale_fill_viridis(discrete = T) + 
    theme(axis.text.x = element_text(angle=90, size=3.5), axis.ticks = element_blank())+
    ggtitle("Mapping Rate for GTEx tissue") 
ggsave("~/output/GTEx_plots/210806_mapping_rate_tissue_GTEx.pdf")
```
```{r}
ggplot(gtex_coldata, aes(x=gtex.smtsd, y=gtex.smrin, fill=gtex.sex)) +
  geom_boxplot(position=position_dodge(1))+
    theme_classic()+
    xlab("Tissue") +
    ylab("RIN Score")+
    scale_fill_viridis(discrete = T) + 
    theme(axis.text.x = element_text(angle=90, size=3.5), axis.ticks = element_blank())+
    ggtitle("RIN for GTEx tissue") + geom_hline(yintercept=5, linetype="dashed", color = "red")
ggsave("~/output/GTEx_plots/210806_RIN_tissue_structures_GTEx.pdf")
```
High-quality RNA will contain an RIN of at least 8, where partially fragmented RNA will contain an RIN within the range of 6–8. Any RNA sample that has a RIN below 5 should not be subjected to further fragmentation during the ScriptSeq protocol, as it will generate smaller than desired fragments.


```{r}
autolysis_score <- as.data.frame(table(gtex_coldata[c(15, 10)]))
ggplot(autolysis_score, aes(fill=gtex.smatsscr, y=Freq, x=gtex.smtsd)) + 
   geom_bar(position="fill", stat="identity") +
    theme_classic()+
    xlab("Tissue") +
    ylab("Fraction")+
    scale_fill_viridis(discrete = T) + 
    theme(axis.text.x = element_text(angle=90, size=3.5), axis.ticks = element_blank())+
    ggtitle("Autolysis Score for GTEx tissue") 
ggsave("~/output/GTEx_plots/210806_autolysis_score_tissue_structure_GTEx_precent.pdf")
```


```{r}
ggplot(gtex_coldata, aes(x=gtex.smtsd , y= gtex.smmppd , fill=gtex.sex))+
  geom_boxplot(position=position_dodge(1))+
    theme_classic()+
    xlab("Tissue") +
    ylab("Total number of reads aligned/map")+
    scale_fill_viridis(discrete = T) + 
    theme(axis.text.x = element_text(angle=90, size=3.5), axis.ticks = element_blank())+
    ggtitle("Total number of reads aligned/map for GTEx tissue") 
ggsave("~/output/GTEx_plots/210806_aligned_tissue_structures_GTEx.pdf")
```

```{r}
ggplot(gtex_coldata, aes(x=gtex.smtsd, y=gtex.smtsisch , fill=gtex.sex))+
  geom_boxplot(position=position_dodge(1))+
    theme_classic()+
    xlab("Tissue") +
    ylab("Ischemic Time (mins)")+
    scale_fill_viridis(discrete = T) + 
    theme(axis.text.x = element_text(angle=90, size=3.5), axis.ticks = element_blank())+
    ggtitle("Ischemic Time for GTEx tissue") 
ggsave("~/output/GTEx_plots/210806_ischemic_time_tissue_structures_GTEx.pdf")
```
Interval between actual death, presumed death, or cross clamp application and final tissue stabilization

```{r}
ggplot(gtex_coldata, aes(x=gtex.smtsd, y= gtex.smmaprt, fill=gtex.sex)) +
  geom_boxplot(position=position_dodge(1))+
    theme_classic()+
    xlab("Tissue") +
    ylab("Mapping Rate")+
    scale_fill_viridis(discrete = T) + 
    theme(axis.text.x = element_text(angle=90, size=3.5), axis.ticks = element_blank())+
    ggtitle("Mapping Rate for GTEx tissue") 
ggsave("~/output/GTEx_plots/210806_mapping_rate_tissue_structures_GTEx.pdf")
```
Mapping Rate: Ratio of total mapped reads to total reads

```{r}
sex_sub_table[sex_sub_table$Freq < 10,]
```
```{r}
unique(sex_sub_table$gtex.smtsd[sex_sub_table$Freq < 10])
```
```{r}
#keep bladder aka number 8
remove_tissues<- unique(sex_sub_table$gtex.smtsd[sex_sub_table$Freq < 10])[-8]
```

```{r}
gtex_coldata_v2 <- gtex_coldata[! gtex_coldata$gtex.smtsd %in% remove_tissues,]
gtex_coldata_v2 <- gtex_coldata_v2[ gtex_coldata_v2$gtex.smrin > 5, ]
as.data.frame(table(gtex_coldata_v2[c(15,6)]))
```
```{r}
name <- paste(GTEX_proj_info[,1],"rse", sep="_" )
gtex_counts <- as.data.frame(assay(get(name[1])))
for (i in 2:length(name)){
  counts<-  as.data.frame(assay(get(name[i])))
  gtex_counts <- cbind(gtex_counts, counts)
}
```

```{r}
gtex_counts_v2 <- gtex_counts[,colnames(gtex_counts) %in% gtex_coldata_v2$external_id]
dim(gtex_counts_v2)
dim(gtex_coldata_v2)
setdiff(gtex_coldata_v2$external_id, colnames(gtex_counts_v2))
gtex_coldata_v2$external_id[duplicated(gtex_coldata_v2$external_id)]
```

```{r eval =FALSE}
saveRDS(gtex_coldata_v2,  "~/data/metadata_gtex_filter_samples.rds")
```

```{r eval = FALSE}
saveRDS(gtex_counts_v2 , "~/data/counts_gtex_filter_samples.rds")
```

```{r}
table(gtex_coldata_v2$gtex.smnabtcht)
```

covariates to look at it in PCA: ischemic time, RIN, age, and batch

```{r}
unique(gtex_coldata_v2$gtex.smtsd)
```
```{r}
ids <- gtex_coldata_v2$external_id[gtex_coldata_v2$gtex.smtsd == "Liver" ]
col_data_sub <- gtex_coldata_v2[gtex_coldata_v2$gtex.smtsd == "Liver" , ]
counts_sub <- gtex_counts_v2[, colnames(gtex_counts_v2) %in% ids ]
pca_sub<- prcomp(t(counts_sub ))
```

```{r}
library(RColorBrewer)
```

the lighter the color the higher the value. 
```{r}
#colorfunc = colorRamp(c("blue", "white","red"))
FUN <- colorRamp(c("black", "#FDE725FF" ))
#col_data_sub$gtex.smrin
MPG <- with(col_data_sub, (gtex.smrin - min(gtex.smrin)) / diff(range(gtex.smrin)))
mycolors = rgb(FUN(MPG), maxColorValue=256)
par(cex=1.0, cex.axis=0.8, cex.main=0.8)
pairs(pca_sub$x[,1:5], col=mycolors, main="Principal components analysis bi-plot\nPCs 1-5", pch=16)
```

```{r}
MPG <- with(col_data_sub, (gtex.smtsisch - min(gtex.smtsisch)) / diff(range(gtex.smtsisch)))
mycolors = rgb(FUN(MPG), maxColorValue=256)
par(cex=1.0, cex.axis=0.8, cex.main=0.8)
pairs(pca_sub$x[,1:5], col=mycolors, main="Principal components analysis bi-plot\nPCs 1-5", pch=16)
```

```{r}
cor(col_data_sub$gtex.smtsisch ,col_data_sub$gtex.smrin , method= "spearman")
```
somewhat of a relationship but not strong. 


```{r}
#MPG <- with(col_data_sub, (gtex.age - min(gtex.age)) / diff(range(gtex.age)))
age<- col_data_sub$gtex.age
#mycolors = rgb(FUN(col_data_sub$gtex.age), maxColorValue=256)
age<- ifelse(age == "20-29", "#440154FF",age)
age<- ifelse(age == "30-39", "#414487FF",age)
age<- ifelse(age == "40-49", "#2A788EFF",age)
age <- ifelse(age == "50-59", "#22A884FF",age)
age <- ifelse(age == "60-69", "#7AD151FF",age)
age <- ifelse(age == "70-79", "#FDE725FF",age)
par(cex=1.0, cex.axis=0.8, cex.main=0.8)
pairs(pca_sub$x[,1:5], col=age, main="Principal components analysis bi-plot\nPCs 1-5", pch=16)
```
some grouping by age but it is mostly RIN and Ischemic time; using all the investigated covariates for downstream models

#plots for age and batches


### Notes
*Worried bout the  kidney low RIN and high autolysis scores

*worried about about the quality of the alignment of whole blood


### Save Data
NA

### Save Figures
Done in script

# END

    Location of final scripts:
    "/home/rstudio/script"

    Location of data produced:
    "/home/rstudio/output/GTEx_plots/"

    Dates when operations were done:
    210810 and again on 220803 for project

## Versions
```{r}
sessionInfo()
```

