---
title: "220722_Meddra_mapping"
author: "Jennifer Fisher"
date: '2022-07-22'
output:
  html_document:
    toc: true
    theme: united
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# START 
    Goal/Purpose of operations: 
    The goal of this script is to mapp the Meddra terms from the lowest level to the highest level for downstream mapping of adverse events in FARES database
    
    Started psedocode on: 
    220722
    
    System which operations were done on: 
    my laptop with a docker
    
    GitHub Repo:
    230321_JLF_Sex_bias_adverse_events
    
    Docker:
    jenfisher7/rstudio_sex_bias_drugs
    
    Directory of operations: 
    /Users/jenniferfisher/Documents/230321_JLF_Sex_bias_adverse_events
    
    Scripts being edited for operations: 
    NA
    
    Data being used:description and location
    MedDRA database in July 2022 version 25.0
    https://www.meddra.org/how-to-use/basics/hierarchy
    
    Papers and tools:
    NA

# STEPS
### Set working directory 
```{r}
setwd("/home/rstudio")
```

```{r}
source("~/src/sex_bias_drug_functions.R")
library(ggplot2)
library(viridis)
```


### load in data

create the meddra term mapping 
https://www.meddra.org/how-to-use/basics/hierarchy
```{r}
#lowest_level_term
llt_meddra<- read.delim("~/data/MedDRA_25_0_English/MedAscii/llt.asc", sep="$", header= FALSE)
# preferred terms
pt_meddra <- read.delim("~/data/MedDRA_25_0_English/MedAscii/pt.asc", sep="$", header= FALSE)
#high level term
hlt_meddra<- read.delim("~/data/MedDRA_25_0_English/MedAscii/hlt.asc", sep="$", header= FALSE)
#high level group 
hlgt_meddra<- read.delim("~/data/MedDRA_25_0_English/MedAscii/hlgt.asc", sep="$", header= FALSE)
#system organ
soc_meddra <- read.delim("~/data/MedDRA_25_0_English/MedAscii/soc.asc", sep="$", header= FALSE)
```

create a mapping
1- llt, 2-pt, 3-hlt, 4-hglt, 5-soc, repeat with the names
```{r}
meddra_mapping <- llt_meddra[,1:3]
#add the pt and soc info 
info_add<- c("pt_id", "pt_name", "soc_id")
for (i in 1:nrow(meddra_mapping)){
  info_1 <- pt_meddra[meddra_mapping[i,3] == pt_meddra[,1],c(1,2,4)]
  info_add<- rbind(info_add, info_1)
}
```

```{r}
info_add<- info_add[-1,]
rownames(meddra_mapping) <- NULL
rownames(info_add) <- NULL
meddra_mapping_v2 <- cbind(meddra_mapping, info_add)
```
add the soc names
```{r}
soc_list <- c("soc_name", "soc_short")
for(i in 1:nrow(meddra_mapping_v2)){
  soc_list_v2 <- soc_meddra[meddra_mapping_v2[i,6]== soc_meddra[,1], c(2,3)]
  soc_list <- rbind(soc_list, soc_list_v2)
}
soc_list<- soc_list[-1,]
```

```{r}
meddrda_mapping_all <- cbind(meddra_mapping_v2, soc_list)
meddrda_mapping_all[1:5,]
```

```{r}
colnames(meddrda_mapping_all)<- c("llt_id", "llt_name", "pt_id", "pt_id2", "pt_name", "soc_id", "soc_name", "soc_short")
```

### Save Data
```{r}
saveRDS(meddrda_mapping_all, "~/data/MedDRA_mapping_llt_pt_soc.rds" )
```


# END
    Location of final scripts:
    src
    Location of data produced:
    ~/data
    Dates when operations were done:
    220722
    
## Versions
```{r}
sessionInfo()
```

