---
title: "230105_lioness_set_up"
author: "Jennifer Fisher"
date: "2023-01-05"
output:
  html_document:
    toc: true
    theme: united
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# START 
    Goal/Purpose of operations: 

    
    Finished psedocode on:
    230105
    
    System which operations were done on: 
    my laptop(on cheaha)
    
    GitHub Repo:
    Sex_Bias_Adverse_Events
    
    Directory of operations:
    /data/project/lasseigne_lab/JLF_scratch/Sex_Bias_Adverse_Events
    
    Scripts being edited for operations:
    sex_bias_functions.R 
    
    Data being used:
 
    
    Papers and tools: 
    NA

# STEPS

### Set working directory 
```{r}
source("/data/project/lasseigne_lab/JLF_scratch/Sex_Bias_Adverse_Events/script/sex_bias_functions.R")
library(netZooR)
```

### load in data
going to load in the ppi and transcription factor networks
```{r}
motif_table <- readRDS("/data/project/lasseigne_lab/JLF_scratch/Sex_Bias_Adverse_Events/data/panda_input_data/230104_symbol_motif_table.rds")
ppi_net <- readRDS("/data/project/lasseigne_lab/JLF_scratch/Sex_Bias_Adverse_Events/data/panda_input_data/230104_symbol_ppi_network_11_5_0.rds")
```
make a gene feature object from recount3

```{r}
#SRP118922
library(recount3)
human_projects<- available_projects()
test<- create_rse(human_projects[(human_projects$project == "SRP118922"),])
feature_info <- test@rowRanges
rm(test)
```

```{r}
saveRDS(feature_info, "/data/project/lasseigne_lab/JLF_scratch/Sex_Bias_Adverse_Events/data/recount3_feature_info.rds")
```


### Analysis
 
Commented sections are a test case for functions
```{r}
#Liver_gtex_qsmooth_expression <- readRDS("/data/project/lasseigne_lab/JLF_scratch/Sex_Bias_Adverse_Events/data/gtex_qsmooth/Liver_gtex_qsmooth_expression.rds")
```
fix the gene ids to match 
```{r}
#gene_ids<- feature_info$gene_id[feature_info$gene_id %in% rownames(Liver_gtex_qsmooth_expression)]
#gene_symbols<- feature_info$gene_name[feature_info$gene_id %in% rownames(Liver_gtex_qsmooth_expression)]
```

```{r}
#identical(rownames(Liver_gtex_qsmooth_expression), gene_ids)
```
```{r}
#rownames(Liver_gtex_qsmooth_expression)<- gene_symbols
```

```{r}
#liver_express_rm <- keep_max_duplicate(Liver_gtex_qsmooth_expression)
```


```{r}
#liver_qsmooth_expr_set <- ExpressionSet(assayData=liver_express_rm)
```

```{r}
ppi_net<- as.data.frame(ppi_net)
ppi_net[,3]<- as.numeric(ppi_net[,3])
```

```{r}
saveRDS( ppi_net, "/data/project/lasseigne_lab/JLF_scratch/Sex_Bias_Adverse_Events/data/panda_input_data/230104_symbol_numeric_ppi_network_11_5_0.rds")
```


```{r}
#reload to test function
#Liver_test_expression <- readRDS("/data/project/lasseigne_lab/JLF_scratch/Sex_Bias_Adverse_Events/data/gtex_qsmooth/Liver_gtex_qsmooth_expression.rds")
```

```{r eval =FALSE}
#panda_tissue_run(Liver_test_expression,
#                expression_file="/data/project/lasseigne_lab/JLF_scratch/Sex_Bias_Adverse_Events/output/lioness/processed_normalized_expression/test.rds",  
#                network_file= "/data/project/lasseigne_lab/JLF_scratch/Sex_Bias_Adverse_Events/output/lioness/tissue_panda/test.rds")
```

```{r}
metadata <- readRDS("/data/project/lasseigne_lab/JLF_scratch/Sex_Bias_Adverse_Events/data/metadata_gtex_filter_samples.rds")

tissues <- unique(metadata$gtex.smtsd)
#remove cell tissues
tissues<- tissues[! grepl("Cell", tissues)]
```

```{r eval=FALSE}
for( i in 1:length(tissues)){
  
  print(tissues[i])
  file_name<- paste("/data/project/lasseigne_lab/JLF_scratch/Sex_Bias_Adverse_Events/data/gtex_qsmooth/", tissues[i], "_gtex_qsmooth_expression.rds", sep = "")
  expression <- readRDS(file_name)
  file_expression <- paste("/data/project/lasseigne_lab/JLF_scratch/Sex_Bias_Adverse_Events/output/lioness/processed_normalized_expression/", tissues[i], "_gtex_processed_expression.rds", sep = "")
  file_network <- paste("/data/project/lasseigne_lab/JLF_scratch/Sex_Bias_Adverse_Events/output/lioness/tissue_panda/", tissues[i], "_gtex_panda_tissue_networks.rds", sep = "")
  
  try(panda_tissue_run(expression,
                expression_file=file_expression,  
                network_file= file_network))
}
```


### Save Data
Done in script
    
# END
    Location of final scripts:
    scripts
  
    
    Location of data produced:
    /data/project/lasseigne_lab/JLF_scratch/Sex_Bias_Adverse_Events/data/panda_input_data
    
    Dates when operations were done:
    230104
    
## Versions
```{r}
sessionInfo()
```
