---
title: "230103_network_qsmooth"
author: "Jennifer Fisher"
date: "2023-01-04"
output:
  html_document:
    toc: true
    theme: united
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# START 
    Goal/Purpose of operations: 
    Need to normalize the expression data for the network analysis.
    
    Finished psedocode on:
    230103
    
    System which operations were done on: 
    Cheaha
    
    GitHub Repo:
    230321_JLF_Sex_bias_adverse_events
    
    Docker/Singularity: 
    rstudio_sex_bias_drugs
    
    Directory of operations: 
    "/data/project/lasseigne_lab/JLF_scratch/230321_JLF_Sex_bias_adverse_events"
    
    Scripts being edited for operations:
    created a new function gtex_qsmooth_function in the sex_bias_functions.R for the data wrangling 
    
    Data being used: 
    gene expression and meta data from the 210806_GTEx_Age_Sex.Rmd script
    
    Papers and tools:
    qsmooth from the YARN R package 

# STEPS
### Set working directory 

```{r}
#change this path to match your paths 
dir_path <- "/data/project/lasseigne_lab/JLF_scratch/230321_JLF_Sex_bias_adverse_events/"
```


```{r}
source(paste0(dir_path,"script/sex_bias_drug_functions.R" ))
library(Biobase)
library(netZooR)
```


### load in data

```{r}
counts_gtex <- readRDS(paste0(dir_path,"data/counts_gtex_filter_samples.rds" ))
metadata <- readRDS(paste0(dir_path, "data/metadata_gtex_filter_samples.rds" ))
```


### Analysis

```{r}
tissues <- unique(metadata$gtex.smtsd)
#remove cell tissues
tissues <- tissues[! grepl("Cell", tissues)]
```


```{r}
counts_gtex <- as.matrix(counts_gtex)
```

```{r eval =FALSE}
for (i in 1:length(tissues)){
  file_fe<- paste(paste0(dir_path,"data/gtex_qsmooth/" ), tissues[i], "_gtex_filtered_expression_set.rds", sep = "")
  file_n<- paste(paste0(dir_path,"data/gtex_qsmooth/" ), tissues[i], "_gtex_qsmooth_expression.rds", sep = "")
  file_m <- paste(paste0(dir_path,"data/gtex_qsmooth/" ), tissues[i], "_gtex_metadata.rds", sep = "")
  print(tissues[i])
  gtex_qsmooth_function(counts_gtex, metadata, tissues[i] , file_filtered_expression=file_fe, file_norm=file_n, file_meta=file_m )
}
```

#### Output:
[1] "Adipose - Subcutaneous"

[1] "check the metadata and count data match"

[1] TRUE

[1] "make expression set object"

[1] "filter lowly expressed genes"

[1] "before"

Features  Samples 

   63856      724 

[1] "after"

Features  Samples 

   20043      724 

[1] "save filtered expression"

[1] "run and save results for qsmooth from YARN R package"

[1] "save metadata"

[1] "Adipose - Visceral (Omentum)"

[1] "check the metadata and count data match"

[1] TRUE

[1] "make expression set object"

[1] "filter lowly expressed genes"

[1] "before"

Features  Samples 

   63856      562 

[1] "after"

Features  Samples 
   20453      562 

[1] "save filtered expression"

[1] "run and save results for qsmooth from YARN R package"

[1] "save metadata"

[1] "Muscle - Skeletal"

[1] "check the metadata and count data match"

[1] TRUE

[1] "make expression set object"

[1] "filter lowly expressed genes"

[1] "before"

Features  Samples 

   63856      875 

[1] "after"

Features  Samples 

   17382      875 

[1] "save filtered expression"

[1] "run and save results for qsmooth from YARN R package"

[1] "save metadata"

[1] "Artery - Tibial"

[1] "check the metadata and count data match"

[1] TRUE

[1] "make expression set object"

[1] "filter lowly expressed genes"

[1] "before"

Features  Samples 

   63856      686 

[1] "after"

Features  Samples 

   19163      686 

[1] "save filtered expression"

[1] "run and save results for qsmooth from YARN R package"

[1] "save metadata"

[1] "Artery - Aorta"

[1] "check the metadata and count data match"

[1] TRUE

[1] "make expression set object"

[1] "filter lowly expressed genes"

[1] "before"

Features  Samples 

   63856      452 

[1] "after"

Features  Samples 

   19530      452 

[1] "save filtered expression"

[1] "run and save results for qsmooth from YARN R package"

[1] "save metadata"

[1] "Artery - Coronary"

[1] "check the metadata and count data match"

[1] TRUE

[1] "make expression set object"

[1] "filter lowly expressed genes"

[1] "before"

Features  Samples 

   63856      252 

[1] "after"

Features  Samples 

   19893      252 

[1] "save filtered expression"

[1] "run and save results for qsmooth from YARN R package"

[1] "save metadata"

[1] "Heart - Atrial Appendage"

[1] "check the metadata and count data match"

[1] TRUE

[1] "make expression set object"

[1] "filter lowly expressed genes"

[1] "before"
Features  Samples 

   63856      449 

[1] "after"

Features  Samples 

   19202      449 

[1] "save filtered expression"

[1] "run and save results for qsmooth from YARN R package"

[1] "save metadata"

[1] "Heart - Left Ventricle"

[1] "check the metadata and count data match"

[1] TRUE

[1] "make expression set object"

[1] "filter lowly expressed genes"

[1] "before"

Features  Samples 

   63856      488 

[1] "after"

Features  Samples 

   17997      488 

[1] "save filtered expression"

[1] "run and save results for qsmooth from YARN R package"

[1] "save metadata"

[1] "Breast - Mammary Tissue"

[1] "check the metadata and count data match"

[1] TRUE

[1] "make expression set object"

[1] "filter lowly expressed genes"

[1] "before"

Features  Samples 

   63856      481 

[1] "after"

Features  Samples 

   21131      481 

[1] "save filtered expression"

[1] "run and save results for qsmooth from YARN R package"

[1] "save metadata"

[1] "Skin - Sun Exposed (Lower leg)"

[1] "check the metadata and count data match"

[1] TRUE

[1] "make expression set object"

[1] "filter lowly expressed genes"

[1] "before"

Features  Samples 

   63856      775 

[1] "after"

Features  Samples 

   20587      775 
[1] "save filtered expression"

[1] "run and save results for qsmooth from YARN R package"

[1] "save metadata"

[1] "Skin - Not Sun Exposed (Suprapubic)"

[1] "check the metadata and count data match"

[1] TRUE

[1] "make expression set object"

[1] "filter lowly expressed genes"

[1] "before"

Features  Samples 

   63856      637 

[1] "after"

Features  Samples 

   20720      637 

[1] "save filtered expression"

[1] "run and save results for qsmooth from YARN R package"

[1] "save metadata"

[1] "Minor Salivary Gland"

[1] "check the metadata and count data match"

[1] TRUE

[1] "make expression set object"

[1] "filter lowly expressed genes"

[1] "before"

Features  Samples 

   63856      178 

[1] "after"

Features  Samples 

   21314      178 

[1] "save filtered expression"

[1] "run and save results for qsmooth from YARN R package"

[1] "save metadata"

[1] "Brain - Hippocampus"

[1] "check the metadata and count data match"

[1] TRUE

[1] "make expression set object"

[1] "filter lowly expressed genes"

[1] "before"



Features  Samples 

   63856      220 

[1] "after"

Features  Samples 

   20285      220 

[1] "save filtered expression"

[1] "run and save results for qsmooth from YARN R package"

[1] "save metadata"

[1] "Brain - Cortex"

[1] "check the metadata and count data match"

[1] TRUE

[1] "make expression set object"

[1] "filter lowly expressed genes"

[1] "before"

Features  Samples 

   63856      282 

[1] "after"

Features  Samples 

   20293      282 

[1] "save filtered expression"

[1] "run and save results for qsmooth from YARN R package"

[1] "save metadata"

[1] "Brain - Putamen (basal ganglia)"

[1] "check the metadata and count data match"

[1] TRUE

[1] "make expression set object"

[1] "filter lowly expressed genes"

[1] "before"

Features  Samples 

   63856      221 

[1] "after"

Features  Samples 

   20080      221 

[1] "save filtered expression"

[1] "run and save results for qsmooth from YARN R package"

[1] "save metadata"

[1] "Brain - Anterior cingulate cortex (BA24)"

[1] "check the metadata and count data match"

[1] TRUE

[1] "make expression set object"

[1] "filter lowly expressed genes"

[1] "before"

Features  Samples 

   63856      201 

[1] "after"

Features  Samples 

   20118      201 

[1] "save filtered expression"

[1] "run and save results for qsmooth from YARN R package"

[1] "save metadata"

[1] "Brain - Cerebellar Hemisphere"

[1] "check the metadata and count data match"

[1] TRUE

[1] "make expression set object"

[1] "filter lowly expressed genes"

[1] "before"

Features  Samples 

   63856      246 

[1] "after"

Features  Samples 

   20418      246 

[1] "save filtered expression"

[1] "run and save results for qsmooth from YARN R package"

[1] "save metadata"

[1] "Brain - Frontal Cortex (BA9)"

[1] "check the metadata and count data match"

[1] TRUE

[1] "make expression set object"
[1] "filter lowly expressed genes"

[1] "before"

Features  Samples 

   63856      221 

[1] "after"

Features  Samples 

   20237      221 

[1] "save filtered expression"

[1] "run and save results for qsmooth from YARN R package"

[1] "save metadata"

[1] "Brain - Spinal cord (cervical c-1)"
[1] "check the metadata and count data match"


[1] TRUE

[1] "make expression set object"

[1] "filter lowly expressed genes"

[1] "before"

Features  Samples 

   63856      171 
[1] "after"

Features  Samples 

   19807      171 

[1] "save filtered expression"

[1] "run and save results for qsmooth from YARN R package"

[1] "save metadata"

[1] "Brain - Substantia nigra"

[1] "check the metadata and count data match"

[1] TRUE

[1] "make expression set object"

[1] "filter lowly expressed genes"

[1] "before"
Features  Samples 

   63856      154 

[1] "after"

Features  Samples 

   20130      154 

[1] "save filtered expression"

[1] "run and save results for qsmooth from YARN R package"

[1] "save metadata"

[1] "Brain - Nucleus accumbens (basal ganglia)"

[1] "check the metadata and count data match"

[1] TRUE

[1] "make expression set object"

[1] "filter lowly expressed genes"

[1] "before"

Features  Samples 

   63856      262 

[1] "after"

Features  Samples 

   20552      262 

[1] "save filtered expression"

[1] "run and save results for qsmooth from YARN R package"

[1] "save metadata"

[1] "Brain - Hypothalamus"
[1] "check the metadata and count data match"

[1] TRUE

[1] "make expression set object"

[1] "filter lowly expressed genes"

[1] "before"

Features  Samples 

   63856      221 

[1] "after"

Features  Samples 

   20923      221 

[1] "save filtered expression"

[1] "run and save results for qsmooth from YARN R package"

[1] "save metadata"

[1] "Brain - Cerebellum"

[1] "check the metadata and count data match"

[1] TRUE

[1] "make expression set object"

[1] "filter lowly expressed genes"

[1] "before"

Features  Samples 

   63856      279 

[1] "after"

Features  Samples 

   20512      279 

[1] "save filtered expression"

[1] "run and save results for qsmooth from YARN R package"

[1] "save metadata"

[1] "Brain - Caudate (basal ganglia)"

[1] "check the metadata and count data match"

[1] TRUE

[1] "make expression set object"

[1] "filter lowly expressed genes"

[1] "before"

Features  Samples 

   63856      273 

[1] "after"

Features  Samples 

   20480      273 

[1] "save filtered expression"

[1] "run and save results for qsmooth from YARN R package"

[1] "save metadata"

[1] "Brain - Amygdala"

[1] "check the metadata and count data match"

[1] TRUE

[1] "make expression set object"

[1] "filter lowly expressed genes"

[1] "before"

Features  Samples 

   63856      162 

[1] "after"

Features  Samples 

   19989      162 

[1] "save filtered expression"

[1] "run and save results for qsmooth from YARN R package"

[1] "save metadata"

[1] "Adrenal Gland"

[1] "check the metadata and count data match"

[1] TRUE

[1] "make expression set object"

[1] "filter lowly expressed genes"

[1] "before"

Features  Samples 

   63856      274 

[1] "after"

Features  Samples 

   19644      274 

[1] "save filtered expression"

[1] "run and save results for qsmooth from YARN R package"

[1] "save metadata"

[1] "Thyroid"

[1] "check the metadata and count data match"

[1] TRUE

[1] "make expression set object"

[1] "filter lowly expressed genes"

[1] "before"

Features  Samples 

   63856      700 

[1] "after"

Features  Samples 

   21306      700 

[1] "save filtered expression"

[1] "run and save results for qsmooth from YARN R package"

[1] "save metadata"

[1] "Lung"

[1] "check the metadata and count data match"

[1] TRUE

[1] "make expression set object"

[1] "filter lowly expressed genes"

[1] "before"

Features  Samples 

   63856      640 

[1] "after"

Features  Samples 

   21518      640 

[1] "save filtered expression"

[1] "run and save results for qsmooth from YARN R package"

[1] "save metadata"

[1] "Spleen"
[1] "check the metadata and count data match"

[1] TRUE

[1] "make expression set object"

[1] "filter lowly expressed genes"

[1] "before"

Features  Samples 

   63856      254 

[1] "after"

Features  Samples 

   20787      254 

[1] "save filtered expression"

[1] "run and save results for qsmooth from YARN R package"

[1] "save metadata"

[1] "Pancreas"

[1] "check the metadata and count data match"

[1] TRUE

[1] "make expression set object"

[1] "filter lowly expressed genes"

[1] "before"

Features  Samples 
   63856      355 

[1] "after"

Features  Samples 

   18591      355 

[1] "save filtered expression"

[1] "run and save results for qsmooth from YARN R package"

[1] "save metadata"

[1] "Esophagus - Muscularis"

[1] "check the metadata and count data match"

[1] TRUE

[1] "make expression set object"

[1] "filter lowly expressed genes"

[1] "before"

Features  Samples 

   63856      553 

[1] "after"

Features  Samples 

   19670      553 

[1] "save filtered expression"

[1] "run and save results for qsmooth from YARN R package"

[1] "save metadata"

[1] "Esophagus - Mucosa"

[1] "check the metadata and count data match"

[1] TRUE

[1] "make expression set object"

[1] "filter lowly expressed genes"

[1] "before"

Features  Samples 

   63856      624 

[1] "after"

Features  Samples 

   20035      624 

[1] "save filtered expression"

[1] "run and save results for qsmooth from YARN R package"


[1] "save metadata"

[1] "Esophagus - Gastroesophageal Junction"

[1] "check the metadata and count data match"

[1] TRUE

[1] "make expression set object"

[1] "filter lowly expressed genes"

[1] "before"

Features  Samples 

   63856      399 

[1] "after"

Features  Samples 

   19938      399 

[1] "save filtered expression"

[1] "run and save results for qsmooth from YARN R package"

[1] "save metadata"

[1] "Stomach"

[1] "check the metadata and count data match"

[1] TRUE

[1] "make expression set object"

[1] "filter lowly expressed genes"

[1] "before"

Features  Samples 

   63856      384 

[1] "after"


Features  Samples 
   20199      384 

[1] "save filtered expression"

[1] "run and save results for qsmooth from YARN R package"

[1] "save metadata"

[1] "Colon - Transverse"

[1] "check the metadata and count data match"

[1] TRUE

[1] "make expression set object"

[1] "filter lowly expressed genes"

[1] "before"

Features  Samples 

   63856      432 

[1] "after"

Features  Samples 

   21117      432 

[1] "save filtered expression"

[1] "run and save results for qsmooth from YARN R package"

[1] "save metadata"

[1] "Colon - Sigmoid"

[1] "check the metadata and count data match"
[1] TRUE

[1] "make expression set object"

[1] "filter lowly expressed genes"

[1] "before"

Features  Samples 

   63856      389 

[1] "after"
Features  Samples 


   20073      389 

[1] "save filtered expression"

[1] "run and save results for qsmooth from YARN R package"

[1] "save metadata"

[1] "Small Intestine - Terminal Ileum"

[1] "check the metadata and count data match"

[1] TRUE

[1] "make expression set object"

[1] "filter lowly expressed genes"

[1] "before"

Features  Samples 

   63856      193 

[1] "after"
Features  Samples 

   21726      193 

[1] "save filtered expression"

[1] "run and save results for qsmooth from YARN R package"

[1] "save metadata"

[1] "Nerve - Tibial"

[1] "check the metadata and count data match"

[1] TRUE

[1] "make expression set object"

[1] "filter lowly expressed genes"

[1] "before"

Features  Samples 

   63856      658 

[1] "after"

Features  Samples 

   20864      658 

[1] "save filtered expression"

[1] "run and save results for qsmooth from YARN R package"

[1] "save metadata"

[1] "Pituitary"

[1] "check the metadata and count data match"

[1] TRUE

[1] "make expression set object"

[1] "filter lowly expressed genes"

[1] "before"

Features  Samples 
   63856      301 

[1] "after"

Features  Samples 

   22343      301 

[1] "save filtered expression"

[1] "run and save results for qsmooth from YARN R package"

[1] "save metadata"

[1] "Whole Blood"

[1] "check the metadata and count data match"

[1] TRUE

[1] "make expression set object"

[1] "filter lowly expressed genes"

[1] "before"

Features  Samples 

   63856      840 

[1] "after"

Features  Samples 
   16474      840 

[1] "save filtered expression"

[1] "run and save results for qsmooth from YARN R package"

[1] "save metadata"

[1] "Liver"

[1] "check the metadata and count data match"

[1] TRUE

[1] "make expression set object"

[1] "filter lowly expressed genes"

[1] "before"

Features  Samples 

   63856      251 

[1] "after"

Features  Samples 

   18654      251 

[1] "save filtered expression"

[1] "run and save results for qsmooth from YARN R package"

[1] "save metadata"

[1] "Kidney - Cortex"

[1] "check the metadata and count data match"

[1] TRUE

[1] "make expression set object"

[1] "filter lowly expressed genes"

[1] "before"

Features  Samples 

   63856       94 

[1] "after"

Features  Samples 

   21294       94 

[1] "save filtered expression"

[1] "run and save results for qsmooth from YARN R package"

[1] "save metadata"

[1] "Bladder"

[1] "check the metadata and count data match"

[1] TRUE

[1] "make expression set object"

[1] "filter lowly expressed genes"

[1] "before"

Features  Samples 

   63856       21 

[1] "after"

Features  Samples 

   21213       21 

[1] "save filtered expression"

[1] "run and save results for qsmooth from YARN R package"

[1] "save metadata"

### Save Data
done within the script 

### Save Figures
done within the script 
    
# END
    Location of final scripts:
    scr
    
    Location of data produced:
    ./data/gtex_qsmooth/
    
    Dates when operations were done:
    230104
    
## Versions
```{r}
sessionInfo()
```