---
title: "230130_alpaca_sex_tissues"
author: "Jennifer Fisher"
date: "2023-01-30"
output:
  html_document:
    toc: true
    theme: united
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# START 
    Goal/Purpose of operations: 
    I need to convert the data format of the sex specific tissue networks outputs to longer by dataframes for alpaca differential community analysis 
    
    Finished psedocode on:
    230130
    
    System which operations were done on: 
    Cheaha
    
    GitHub Repo:
    Sex_Bias_Adverse_Events
    
    Docker/Singularity: 
    rstudio_sex_bias_drugs
    
    Directory of operations: 
    "/data/project/lasseigne_lab/JLF_scratch/Sex_Bias_Adverse_Events"
    
    Scripts being edited for operations:
    created a new function xx in the sex_bias_functions.R for the data wrangling 
    
    Data being used: 
    the sex -spcific tissue networks from ./output/alpaca/sex_specific_networks
    
    Papers and tools:
    panda and alpaca from the netzoor package

# STEPS
### Set working directory 
```{r}
#change this path to match your paths 
dir_path <- "/data/project/lasseigne_lab/JLF_scratch/Sex_Bias_Adverse_Events"
```


```{r}
source(paste0(dir_path, "/script/sex_bias_drug_functions.R"))
library(tidyverse)
library(netZooR)
```


### load in data

```{r}
metadata <- readRDS(paste0(dir_path, "/data/metadata_gtex_filter_samples.rds"))

tissues <- unique(metadata$gtex.smtsd)
#remove cell tissues
tissues<- tissues[! grepl("Cell", tissues)]
```
note processed data had the white sapce removed 
```{r}
tissues_wo_ws <- gsub(" ", "_", tissues, fixed = TRUE)
```

```{r}
panda_output_files <- list.files(paste0(dir_path, "/output/alpaca/sex_specific_networks"))
```

create input files for alpaca analysis (male then female)
```{r eval = FALSE}
for (i in seq(1, 86, by =2)){
  #print(i)
  j<- i +1
  tissue_name <- unlist(str_split(panda_output_files[i], "F_")[[1]])[1]
  tissue_name2 <- unlist(str_split(panda_output_files[j], "M_")[[1]])[1]
  if (tissue_name == tissue_name2){
    print(tissue_name)
    create_alpaca_input(panda_output_files[i], panda_output_files[j], tissue_name)
  }else{
    print("tissues dont match")
    print(tissue_name, tissue_name2)
  }
  
}
```


### Analysis

```{r}
#liver test 
#Liver__topNet_regnet_male_female_long <- readRDS("/data/project/lasseigne_lab/JLF_scratch/Sex_Bias_Adverse_Events/output/alpaca/alpaca_adjusted_inputs/Liver__topNet_regnet_male_female_long.rds")
#liver_regnet_sex_alp<- alpaca(Liver__topNet_regnet_male_female_long, "/data/project/lasseigne_lab/JLF_scratch/Sex_Bias_Adverse_Events/output/alpaca/alpaca_run/liver_test", verbose = TRUE)

```


```{r}
panda_adjusted_files <- list.files(paste0(dir_path, "/output/alpaca/alpaca_adjusted_inputs"))
```

```{r eval = FALSE}
# female specific communities
for (i in 1:length(panda_adjusted_files)){
  print(panda_adjusted_files[i])
  input<- readRDS(paste0(paste0(dir_path, "/output/alpaca/alpaca_adjusted_inputs/"),  panda_adjusted_files[i], sep = ""))
  
  tissue_name <- unlist(str_split(panda_adjusted_files[i], "__")[[1]])[1]
  print(tissue_name)
  
  file_header<- paste0(paste0(dir_path, "/output/alpaca/alpaca_run/"),  tissue_name, "_",sep = "")
  
  regnet_sex_alp<- alpaca(input, file_header, verbose = TRUE)
}
```


```{r eval = FALSE}
#change the order for male specific communities
for (i in 1:length(panda_adjusted_files)){
  print(panda_adjusted_files[i])
  input<- readRDS(paste0(paste0(dir_path, "/output/alpaca/alpaca_adjusted_inputs/"),  panda_adjusted_files[i], sep = ""))
  input_adj<- input[,c(1,2,4,3)]
  tissue_name <- unlist(str_split(panda_adjusted_files[i], "__")[[1]])[1]
  print(tissue_name)
  
  file_header<- paste0(paste0(dir_path, "/output/alpaca/alpaca_run/"),  tissue_name, "female", "male" , "_",sep = "")
  
  regnet_sex_alp_flipped<- alpaca(input_adj, file_header, verbose = TRUE)
}
```


### Save Data
Done in script
    
# END
    Location of final scripts:
    scripts
  
    
    Location of data produced:
    /data/project/lasseigne_lab/JLF_scratch/Sex_Bias_Adverse_Events/output/lioness/adjusted_lioness_outputs
    
    Dates when operations were done:
    230130
    
## Versions
```{r}
sessionInfo()
```