---
title: "230124_fares_sbae_drugs"
author: "Jennifer Fisher"
date: "2023-01-24"
output:
  html_document:
    toc: true
    theme: united
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# START 
    Goal/Purpose of operations: 
    To determine the drugs with associations to sex-bias adverse events, we need to calulate the reporting odds ratio for each drug to determine if it is more likely for a certain drug and adverse event combination is more likely to coccur in one sex or the other. 
    
    Finished psedocode on:
    230124
    
    System which operations were done on: 
    cheaha
    
    GitHub Repo:
    230321_JLF_Sex_bias_adverse_events
    
    Directory of operations:
    /data/project/lasseigne_lab/JLF_scratch/230321_JLF_Sex_bias_adverse_events
    
    Scripts being edited for operations:
    NA
    
    Data being used:
    patient_safety.csv - https://dataverse.harvard.edu/dataset.xhtml?persistentId=doi:10.7910/DVN/G9SHDA - Curated FAERS dataset containing reports from 2013Q1 to 2020Q3. A primary source of post-marketing pharmacovigilance. The reports in FAERS mainly contain demographic information (such as age and sex, no personal identifiers), drugs (drug substances), and adverse events (preferred terms in MedDRA). https://www.nature.com/articles/s43588-021-00138-4 
    
    Papers and tools: 
    NA

# STEPS

### Set working directory 
```{r}
#change this path to match your paths 
dir_path <- "/data/project/lasseigne_lab/JLF_scratch/230321_JLF_Sex_bias_adverse_events/"
```


```{r}
source(paste0(dir_path, "src/sex_bias_drug_functions.R"))
library(tidyverse)
library(readr)
library(MASS)
library(parallel)
library(BiocParallel)
library(stringr)
library(dplyr)
library(stringi)
```

### load in data
```{r}
patient_safety <- read_csv(paste0(dir_path, "data/patient_safety.csv"))
```


```{r}
#filter data
#us only
patient_safety <- patient_safety[patient_safety$country == "US", ]
#qualification 1=physician, 2=pharmacist, 3= other health professional
patient_safety <- patient_safety[patient_safety$qualify %in% c(1,2,3), ]
#gender 1- male 2- female 
patient_safety <- patient_safety[patient_safety$gender %in% c(1,2), ]
#must have adverse event
patient_safety <- patient_safety[!patient_safety$SE == "[]", ]
#must have a drug 
patient_safety <- patient_safety[!patient_safety$drugs == "[]", ]
```

```{r}
#check for repeating cases
identical(nrow(patient_safety), length(unique(patient_safety$report_id)))
```

```{r}
#look at the duplicated cases
patient_safety$report_id[duplicated(patient_safety$report_id)]
```

```{r}
patient_safety[patient_safety$report_id == 10521631,]
```
different
```{r}
patient_safety[patient_safety$report_id == 10432254,]
```
different 
```{r}
patient_safety[patient_safety$report_id == 15431328,]
```
different

there is an earlier and later versions. I should remove the earlier entry. note this 
```{r}
patient_safety <- patient_safety[! (patient_safety$report_id == 10521631 & patient_safety$receipt_date == "2015-04-17"),]
patient_safety <- patient_safety[! (patient_safety$report_id == 10432254 & patient_safety$receipt_date == "2014-08-15"),]
patient_safety <- patient_safety[! (patient_safety$report_id == 15431328 & patient_safety$receipt_date == "2018-09-21"),]
```


```{r}
saveRDS(patient_safety, paste0(dir_path, "data/patient_safety_filtered.rds"))
```


### Analysis


How many cases? 1960181
```{r}
nrow(patient_safety)
```

```{r}
test_ae <- str_replace_all(patient_safety$SE, "[[:punct:]]", "")
test_ae <- strsplit(test_ae, " ")
test_drug <- str_replace_all(patient_safety$drugs, "[[:punct:]]", "")
test_drug <- strsplit(test_drug, " ")
```

for each case reprot I want to know the combinations of drugs and adverse events
```{r}
parallel_runs<- mclapply(1:nrow(patient_safety),drug_ae_events,  mc.cores= 2)

saveRDS(parallel_runs, paste0(dir_path, "data/drug_ae_list.rds"))
```
5 min run above

I am dividing the list of data frames to divide them smaller groups to combine them.
```{r}
N <- length(parallel_runs)
sz <- 50000 
sample_list <- split(1:N, ceiling(seq_along(1:N)/sz))
```

sample_list is 40 list. 
combine the list of data frames
```{r}
combine_parallel_runs <- mclapply(1:40,combine_sub_data_frames, sample_list, parallel_runs,  mc.cores = 2)
```
~10 min run
```{r}
drug_ae_df <- bind_rows(combine_parallel_runs)
```
remove duplicated adverse events and drug pairs
```{r}
drug_ae_df2 <- drug_ae_df[!duplicated(drug_ae_df), ]
```

```{r}
saveRDS(drug_ae_df2, paste0(dir_path, "data/drug_ae_df.rds"))
```


make adjustment to the lower level terms to be the PT term. 
meddra mappings from 220722_Meddra_mapping.Rmd
```{r}
MedDRA_mapping<- readRDS(paste0(dir_path, "data/MedDRA_mapping_llt_pt_soc.rds"))
```

```{r}
lower_terms <- unique(drug_ae_df2$adverse_event[! drug_ae_df2$adverse_event %in% MedDRA_mapping$pt_id])
saveRDS(lower_terms,paste0(dir_path, "data/FARES_lower_terms.rds") )
```

lower_terms
```{r}
ae_info<- as.data.frame(matrix(nrow= length(lower_terms), ncol= 2))
ae_info[,1]<- lower_terms
for (i in 1:length(lower_terms)){
  ae_info[i,2] <- MedDRA_mapping$pt_id[grep(lower_terms[i], MedDRA_mapping$llt_id )]
}

```

```{r}
saveRDS(ae_info,paste0(dir_path, "data/FARES_lower_terms_df.rds") )
```


```{r}
drug_ae_df2$drug <- as.character(drug_ae_df2$drug)
drug_ae_df2$adverse_event <- as.character(drug_ae_df2$adverse_event)
for (i in 1:nrow(drug_ae_df2)){
  ae_id <- as.character(drug_ae_df2[i, 2])
  if (ae_id %in% lower_terms){
    drug_ae_df2[i,2] <- ae_info$V2[grep(ae_id, ae_info$V1, fixed=TRUE)]
  }
}
```


```{r}
drug_ae_df2 <- drug_ae_df2[!duplicated(drug_ae_df2), ]
```

```{r}
saveRDS(drug_ae_df2, paste0(dir_path, "data/drug_ae_df.rds"))
```

fixed the patient_safety too
```{r}
#test case
#10083123 ->10085121
#est<- "['10013709', '10028314', '10083123', '10041349', '10049816', '10050031', '10053156', '10069296']" 
#stri_replace_all_fixed(test, ae_info$V1, ae_info$V2, vectorize_all=FALSE)

patient_safety$SE <- stri_replace_all_fixed(patient_safety$SE , ae_info$V1, ae_info$V2, vectorize_all=FALSE)
```


```{r}
saveRDS(patient_safety, paste0(dir_path, "data/patient_safety_filtered_pt.rds"))
```


### Save Data
Done in script

# END
    Location of final scripts:
    src
  
    Location of data produced:
    data

    
    Dates when operations were done:
    230124
    
## Versions
```{r}
sessionInfo()
```
