---
title: "230105_panda_set_up"
author: "Jennifer Fisher"
date: "2023-01-05"
output:
  html_document:
    toc: true
    theme: united
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# START 
    Goal/Purpose of operations: 
    Adjusting all the inputs for the panda sex-specific networks
    
    Finished psedocode on:
    230105
    
    System which operations were done on: 
    my laptop(on cheaha)
    
    GitHub Repo:
    230321_JLF_Sex_bias_adverse_events
    
    Directory of operations:
    /data/project/lasseigne_lab/JLF_scratch/230321_JLF_Sex_bias_adverse_events
    
    Docker: 
    jenfisher7/rstudio_sex_bias_drugs
    
    Scripts being edited for operations:
    sex_bias_drug_functions.R 
    
    Data being used:
    data/panda_input_data/230104_symbol_motif_table.rds
    data/panda_input_data/230104_symbol_ppi_network_11_5_0.rds
    
    
    Papers and tools: 
    NA

# STEPS

### Set working directory 
```{r}
#change this path to match your paths 
dir_path <- "/data/project/lasseigne_lab/JLF_scratch/230321_JLF_Sex_bias_adverse_events/"
```


```{r}
source(paste0(dir_path, "src/sex_bias_drug_functions.R" ))
library(netZooR)
```

### load in data
going to load in the ppi and transcription factor networks
```{r}
motif_table <- readRDS(paste0(dir_path, "data/panda_input_data/230104_symbol_motif_table.rds"))
ppi_net <- readRDS(paste0(dir_path, "data/panda_input_data/230104_symbol_ppi_network_11_5_0.rds"))
```


make a gene feature object from recount3 with info about genes
```{r}
#SRP118922- random recount3 project
library(recount3)
human_projects<- available_projects()
test<- create_rse(human_projects[(human_projects$project == "SRP118922"),])
feature_info <- test@rowRanges
rm(test)
```

```{r}
saveRDS(feature_info, paste0(dir_path, "data/recount3_feature_info.rds"))
```


### Analysis
 
Commented sections are a test case for functions
```{r}
#Liver_gtex_qsmooth_expression <- readRDS("/data/project/lasseigne_lab/JLF_scratch/Sex_Bias_Adverse_Events/data/gtex_qsmooth/Liver_gtex_qsmooth_expression.rds")
```
fix the gene ids to match 
```{r}
#gene_ids<- feature_info$gene_id[feature_info$gene_id %in% rownames(Liver_gtex_qsmooth_expression)]
#gene_symbols<- feature_info$gene_name[feature_info$gene_id %in% rownames(Liver_gtex_qsmooth_expression)]
#identical(rownames(Liver_gtex_qsmooth_expression), gene_ids)
#rownames(Liver_gtex_qsmooth_expression)<- gene_symbols
#liver_express_rm <- keep_max_duplicate(Liver_gtex_qsmooth_expression)
#liver_qsmooth_expr_set <- ExpressionSet(assayData=liver_express_rm)
```

```{r}
ppi_net <- as.data.frame(ppi_net)
ppi_net[,3] <- as.numeric(ppi_net[, 3])
```

```{r}
saveRDS(ppi_net, paste0(dir_path, "data/panda_input_data/230104_symbol_numeric_ppi_network_11_5_0.rds"))
```


```{r}
#reload to test function
#Liver_test_expression <- readRDS("/data/project/lasseigne_lab/JLF_scratch/Sex_Bias_Adverse_Events/data/gtex_qsmooth/Liver_gtex_qsmooth_expression.rds")
#panda_tissue_run(Liver_test_expression,
#                expression_file="/data/project/lasseigne_lab/JLF_scratch/Sex_Bias_Adverse_Events/output/lioness/processed_normalized_expression/test.rds",  
#                network_file= "/data/project/lasseigne_lab/JLF_scratch/Sex_Bias_Adverse_Events/output/lioness/tissue_panda/test.rds")
```

Filter normalized expression for each tissue for symbols that don't replicate
```{r}
metadata <- readRDS(paste0(dir_path, "data/metadata_gtex_filter_samples.rds"))

tissues <- unique(metadata$gtex.smtsd)
#remove cell tissues
tissues<- tissues[! grepl("Cell", tissues)]
```

```{r eval=FALSE}
for( i in 1:length(tissues)){
  
  print(tissues[i])
  file_name<- paste(dir_path, "data/gtex_qsmooth/", tissues[i], "_gtex_qsmooth_expression.rds", sep = "")
  expression <- readRDS(file_name)
  file_expression <- paste(dir_path, "results/lioness/processed_normalized_expression/", tissues[i], "_gtex_processed_expression.rds", sep = "")
  
  panda_tissue_fix(expression, expression_file = file_expression)
}
```


### Save Data
Done in script
    
# END
    Location of final scripts:
    src
    
    Location of data produced:
    /data/project/lasseigne_lab/JLF_scratch/230321_JLF_Sex_bias_adverse_events/results/lioness
    
    Dates when operations were done:
    230104
    
## Versions
```{r}
sessionInfo()
```
