---
title: "230109_sex_specific_tissue_networks"
author: "Jennifer Fisher"
date: "2023-01-09"
output:
  html_document:
    toc: true
    theme: united
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# START 
    Goal/Purpose of operations: 
    To compare the communities between two sexes. I need to create sex-specific networks for each tissue and then compare the networks with Alpaca
    
    Finished psedocode on:
    230109
    
    System which operations were done on: 
    Cheaha
    
    GitHub Repo:
    230321_JLF_Sex_bias_adverse_events
    
    Docker/Singularity: 
    rstudio_sex_bias_drugs
    
    Directory of operations: 
    "/data/project/lasseigne_lab/JLF_scratch/230321_JLF_Sex_bias_adverse_events"
    
    Scripts being edited for operations:
    created a new function in the sex_bias_functions.R for building the sex-specific networks
    
    Data being used: 
    the proccessed gene expression from the ./output/proccessed_normalized_expression and meta data from the 210806_GTEx_Age_Sex.Rmd script
    
    Papers and tools:
    panda and alpaca from the netzzor package

# STEPS
### Set working directory 
```{r}
#change this path to match your paths 
dir_path <- here::here()
```


```{r}
source(paste0(dir_path, "/src/sex_bias_drug_functions.R"))
library(Biobase)
library(netZooR)
```


### load in data
the transcription factor and protein-protein interaction networks for panda 
```{r}
motif_table <- readRDS(paste0(dir_path, "/data/panda_input_data/230104_symbol_motif_table.rds"))
ppi_net <- readRDS(paste0(dir_path, "/data/panda_input_data/230104_symbol_numeric_ppi_network_11_5_0.rds"))
```

loading in test data (liver tissue)
```{r}
#Liver_data <- readRDS("/data/project/lasseigne_lab/JLF_scratch/Sex_Bias_Adverse_Events/output/lioness/processed_normalized_expression/Liver_gtex_processed_expression.rds")
#Liver_gtex_metadata <- readRDS("/data/project/lasseigne_lab/JLF_scratch/Sex_Bias_Adverse_Events/data/gtex_qsmooth/Liver_gtex_metadata.rds")
```


```{r}
#test run 
#sex_networks(Liver_data, Liver_gtex_metadata, "/data/project/lasseigne_lab/JLF_scratch/Sex_Bias_Adverse_Events/output/alpaca/sex_specific_networks/Liver_test_M_panda.rds","/data/project/lasseigne_lab/JLF_scratch/Sex_Bias_Adverse_Events/output/alpaca/sex_specific_networks/Liver_test_F_panda.rds" )
```


### Analysis

```{r}
metadata <- readRDS(paste0(dir_path, "/data/metadata_gtex_filter_samples.rds"))

tissues <- unique(metadata$gtex.smtsd)
#remove cell tissues
tissues <- tissues[! grepl("Cell", tissues)]
```
note processed data had the white space removed 
```{r}
tissues_wo_ws <- gsub(" ", "_", tissues, fixed = TRUE)
```


```{r eval=FALSE}
for (i in 1:length(tissues_wo_ws)){
  
  print(tissues[i])
  file_name <- paste(paste0(dir_path, "/results/ac_results/lioness/processed_normalized_expression/"), tissues_wo_ws[i], "_gtex_processed_expression.rds",sep = "")
  #print(file_name)
  expression <- readRDS(file_name)
  
  tisssue_meta_file <- paste(paste0(dir_path, "/data/gtex_qsmooth/"), tissues[i], "_gtex_metadata.rds", sep = "")
  tisssue_meta <- readRDS(tisssue_meta_file)
  
  file_male <- paste(paste0(dir_path, "/results/ac_results/alpaca/sex_specific_networks/"), tissues_wo_ws[i], "_M_panda.rds", sep = "")
  
  file_female <- paste(paste0(dir_path, "/results/ac_results/alpaca/sex_specific_networks/"), tissues_wo_ws[i], "_F_panda.rds", sep = "")
  
  try(sex_networks(expression, 
                   tisssue_meta, 
                   file_male,
                   file_female))
}
```



### Save Data
Done in script
    
# END
    Location of final scripts:
    src
  
    
    Location of data produced:
    /data/project/lasseigne_lab/JLF_scratch/230321_JLF_Sex_bias_adverse_events/results/alpaca/sex_specific_networks
    
    Dates when operations were done:
    230104
    
## Versions
```{r}
sessionInfo()
```


