---
title: "230206_alpaca_results_exploration"
author: "Jennifer Fisher"
date: "2023-02-06"
output:
  html_document:
    toc: true
    theme: united
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# START 
    Goal/Purpose of operations: 
    I am going to be looking at the results from Alpaca for the sex-specific networks across the 43 different tissue by applying pathway analysis. 
    
    Finished psedocode on:
    230206
    
    System which operations were done on: 
    my laptop
    
    GitHub Repo:
    230321_JLF_Sex_bias_adverse_events
    
    Directory of operations:
    ~/Documents/230321_JLF_Sex_bias_adverse_events
    
    Scripts being edited for operations:
    NA
    
    Data being used:
    results/alpaca/alpaca_runs
    
    Papers and tools: 
    NA

# STEPS

### Set working directory 

```{r}
source("~/src/sex_bias_drug_functions.R")
library(gprofiler2)
library(viridis)
library(RColorBrewer)
library(ComplexHeatmap)
library(GOSemSim)
library(stringr)
library(ggplot2)
```


```{r}
metadata <- readRDS("~/data/metadata_gtex_filter_samples.rds")
```

```{r}
tissues <- unique(metadata$gtex.smtsd)
#remove cell tissues
tissues<- tissues[! grepl("Cell", tissues)]
```
note processed data had the white sapce removed 
```{r}
tissues_wo_ws <- gsub(" ", "_", tissues, fixed = TRUE)
```

```{r eval=FALSE}
for (j in 1:length(tissues_wo_ws)){
  alpaca_pathway_wrapper(tissues_wo_ws[j], "female")
  alpaca_pathway_wrapper(tissues_wo_ws[j], "male")
}
```
Bad request, response code 503<- might occur this is a server issue

# top pathways

```{r eval=FALSE}
all_files<- list.files("~/results/alpaca/alpaca_pathway_analysis")
pathway_files <- all_files[grep("pathways.csv", all_files )]
```

```{r eval=FALSE}
all_pathway_res<- read.csv2(paste0("~/results/alpaca/alpaca_pathway_analysis/", pathway_files[1]))
for (j in 2:length(pathway_files)){
  group <- read.csv2(paste0("~/results/alpaca/alpaca_pathway_analysis/", pathway_files[j]), sep = ",")
  all_pathway_res<- rbind(all_pathway_res, group)
}
saveRDS(all_pathway_res, "~/results/alpaca/alpaca_pathway_analysis/all_pathways.rds")
```

```{r eval=FALSE}
trim_1<- str_split(all_pathway_res$set, "_female", simplify = TRUE)[,1]
trim_2<- str_split(trim_1, "_male", simplify = TRUE)[,1]
all_pathway_res$tissue<- trim_2
```

```{r eval=FALSE}
sex_group<- grepl("female", all_pathway_res$set)
all_pathway_res$sex <- ifelse(sex_group, "female", "male")
```
```{r eval=FALSE}
all_pathway_res$comm_number <- as.numeric(gsub(".*?([0-9]+).*", "\\1", all_pathway_res$set))   
```

```{r eval=FALSE}
saveRDS(all_pathway_res, "~/results/alpaca/alpaca_pathway_analysis/all_pathways.rds")
```

```{r eval=FALSE}
all_pathway_res_filter<- cbind(all_pathway_res$term_name,all_pathway_res$source, all_pathway_res$term_id, all_pathway_res$tissue)
colnames(all_pathway_res_filter)<- c("term_name", "source", "id", "tissue")
all_pathway_res_filter<- all_pathway_res_filter[!duplicated(all_pathway_res_filter),]

all_pathway_res_filter_KEGG <- all_pathway_res_filter[all_pathway_res_filter[,2] %in% c("GO:BP"),]
test<- as.data.frame(table(all_pathway_res_filter_KEGG[,3]))
```

```{r eval=FALSE}
go_bp_all_tissues<- test[test$Freq == 43, 1]
```

```{r eval=FALSE}
hsGO <- godata('org.Hs.eg.db', ont="BP")
```

```{r eval=FALSE}
go_sim <- mgoSim(as.character(go_bp_all_tissues), as.character(go_bp_all_tissues), semData=hsGO, measure="Wang", combine=NULL)
#get the parent terms from rrvgo for the pathways
res <- reduceSimMatrix(go_sim , threshold = 0.85)
res_v2<- res[match(colnames(go_sim ), res$go),]
```

```{r eval=FALSE}
sem_res <- as.data.frame(table(res_v2$parentTerm))
```

```{r eval=FALSE}
sem_res$Var1<- factor(sem_res$Var1, levels=sem_res$Var1[order(sem_res$Freq)] )
ggplot(sem_res, aes(y= Var1, x= Freq, label=Freq)) + geom_bar(stat="identity", fill = "white", color= "black")+ 
    geom_text(size = 10, position = position_stack(vjust = 0.9)) + ylab("Parent BP terms") + xlab("number of GO:BP terms enriched once\nin all tissues" ) +theme(text = element_text(size = 10,  face="bold"))
```

look for some drug related pathways
```{r eval=FALSE}
all_pathway_res[grep("cytochrome", all_pathway_res$term_name),]
```

```{r eval=FALSE}
all_pathway_res[grep("p450", all_pathway_res$term_name),]
```

```{r eval=FALSE}
all_pathway_res[grep("xenobiotic", all_pathway_res$term_name),]
```
small intestine is interesting. no chromchrome p450 genes

```{r eval=FALSE}
all_pathway_res[grep("small molecule", all_pathway_res$term_name),]
```

look at the pathways around drug metabolism tissues
drug metabolism- tissue: liver, kidney, intestine ( small and large), lung, adrenal, blood, and skin

Skin_-_Sun_Exposed_(Lower_leg), Skin_-_Not_Sun_Exposed_(Suprapubic), Lung, Adrenal_Gland,

Liver, Small_Intestine_-_Terminal_Ileum,  Whole_Blood, Kidney_-_Cortex

```{r eval=FALSE}
all_pathway_res_drug_tissues<- all_pathway_res[all_pathway_res$tissue %in% c("Liver", "Small_Intestine_-_Terminal_Ileum",  "Whole_Blood", "Kidney_-_Cortex"),]
```

```{r eval=FALSE}
all_pathway_res_filter <- cbind(all_pathway_res_drug_tissues$term_name,all_pathway_res_drug_tissues$source, all_pathway_res_drug_tissues$term_id, all_pathway_res_drug_tissues$tissue)
colnames(all_pathway_res_filter) <- c("term_name", "source", "id", "tissue")
all_pathway_res_filter <- all_pathway_res_filter[!duplicated(all_pathway_res_filter),]

all_pathway_res_filter_KEGG <- all_pathway_res_filter[all_pathway_res_filter[,2] %in% c("GO:BP"),]
test<- as.data.frame(table(all_pathway_res_filter_KEGG[,3]))
```

```{r eval=FALSE}
go_bp_all_tissues <- test[test$Freq == 4, 1]
```

```{r eval=FALSE}
go_sim <- mgoSim(as.character(go_bp_all_tissues), as.character(go_bp_all_tissues), semData=hsGO, measure="Wang", combine=NULL)
#get the parent terms from rrvgo for the pathways
res <- reduceSimMatrix(go_sim , threshold = 0.9)
res_v2 <- res[match(colnames(go_sim ), res$go),]
```


```{r eval=FALSE}
sem_res <- as.data.frame(table(res_v2$parentTerm))
```

```{r eval=FALSE}
sem_res$Var1<- factor(sem_res$Var1, levels=sem_res$Var1[order(sem_res$Freq)] )
ggplot(sem_res, aes(y= Var1, x= Freq, label=Freq)) +
  geom_bar(stat="identity", fill = "white", color= "black") +
    geom_text(size = 10, position = position_stack(vjust = 0.9)) +
  ylab("Parent BP terms") + xlab("number of GO:BP terms enriched once\nin Liver, Small Intestine- Terminal Ileum, Whole Blood, & Kidney -Cortex" ) +
  theme(text = element_text(size = 10,  face="bold"))
```



# drug genes
KEGG_DRUG_METABOLISM_CYTOCHROME_P450.v2022.1.Hs.gmt contain a list of drug metabolism genes
```{r eval=FALSE}
drug_genes <- c("ADH1A"	,"ADH1B", "ADH1C", "ADH4",	"ADH5",	"ADH6",	"ADH7",	
                "ALDH1A3",	"ALDH3A1",	"ALDH3B1", "ALDH3B2", "AOX1", "CYP1A2",
                "CYP2A13", "CYP2A6", "CYP2A7",	"CYP2B6",	"CYP2C18",	"CYP2C19",
                "CYP2C8",	"CYP2C9",	"CYP2D6",	"CYP2E1",	"CYP3A4",	"CYP3A43",	"CYP3A5",
                "CYP3A7",	"FMO1",	"FMO2",	"FMO3", "FMO4",	"FMO5",  "GSTA1", "GSTA2", 	"GSTA3",	"GSTA4",
                "GSTA5",	"GSTK1",	"GSTM1",	"GSTM2",	"GSTM3",	"GSTM4",	"GSTM5",	"GSTO1",
                "GSTO2",	"GSTP1",	"GSTT1",	"GSTT2",	"GSTZ1",	"MAOA",	"MAOB",	"MGST1",
                "MGST2",	"MGST3",	"UGT1A1",	"UGT1A10",	"UGT1A3",	"UGT1A4",
                "UGT1A5",	"UGT1A6",	"UGT1A7",	"UGT1A8", "UGT1A9",	"UGT2A1",	"UGT2A3", "UGT2B10",
                "UGT2B11",	"UGT2B15",	"UGT2B17",	"UGT2B28",	"UGT2B4",	"UGT2B7")
```

```{r eval=FALSE}
test_p<- c()
test_change <- c()
for (j in 1:length(tissues_wo_ws)){ #
   test <- alpaca_gene_set_analysis( tissues_wo_ws[j], genes="drugs", gene_set_name= "Drug Metabolism Genes", path="~/results/alpaca/drug_metabolism_genes/")
  test_p[j]<- test[1]
  test_change[j]<- test[2]
}
```

```{r eval=FALSE}
sig_drug_gene_set <- as.data.frame(cbind(tissues_wo_ws, unlist(test_p), as.numeric(unlist(test_change))))
colnames(sig_drug_gene_set) <- c("Tissue", "p_value", "log2fold_Change")
```

```{r eval=FALSE}
sig_drug_gene_set$neg_log_p <- -log(as.numeric(sig_drug_gene_set$p_value))
sig_drug_gene_set$log2fold_Change <- as.numeric(sig_drug_gene_set$log2fold_Change)
```

volcano plot (not nice looking)
```{r eval=FALSE}
p_value_cut <- -log(0.05/43)
sig_drug_gene_set$Tissue_2 <- ifelse(sig_drug_gene_set$neg_log_p > p_value_cut, sig_drug_gene_set$Tissue, NA )

ggplot(sig_drug_gene_set, aes(x=log2fold_Change, y= neg_log_p, label=Tissue_2)) +
  geom_point() + xlab("log(Fold Change)") + 
  ylab("-log(p-value)") + geom_hline(yintercept=p_value_cut, linetype="dashed", color = "black") + 
  geom_text(size = 5, position = position_stack(vjust = 0.95)) + xlim(c(-3, 3))
```

```{r eval=FALSE}
sig_drug_gene_set$Tissue <- gsub("_", " ", sig_drug_gene_set$Tissue, fixed = TRUE)
sig_drug_gene_set$Tissue<- factor(sig_drug_gene_set$Tissue, levels= sig_drug_gene_set$Tissue[order(sig_drug_gene_set$neg_log_p)])
ggplot(sig_drug_gene_set, aes(y=Tissue, x= neg_log_p, fill= log2fold_Change)) + 
  geom_bar(stat="identity",  color= "black") +
  scale_fill_gradient2(low= "#21908CFF", mid= "white", high= "#440154FF") +
  theme(text = element_text(size = 20,  face="bold")) +
  geom_vline(xintercept=p_value_cut, linetype="dashed", color = "black") + xlab("-log(p-value)") + labs(fill = "log2 fold change")
ggsave("~/results/alpaca/drug_metabolism_genes/230220_differential_mod_drug_genes_wilcox_test.png", height=12, width=10)
```


Make a list of genes in each community
need to apply this to all the communities
```{r eval=FALSE}
female_com_list <- list()
male_com_list<- list()
for ( i in 1:length(tissues_wo_ws)){
  tissue_female <- gene_com_list(tissues_wo_ws[i], "female")
  female_com_list <- append(female_com_list, tissue_female)
  
  tissue_male <- gene_com_list(tissues_wo_ws[i], "male")
  male_com_list <- append(male_com_list, tissue_male)
}
#females
#males
```

```{r eval=FALSE}
all_com_list<- append(female_com_list, male_com_list)
```

calculate the jacard similarity for every community
```{r eval=FALSE}
res<- matrix(nrow= length(all_com_list), ncol= length(all_com_list))
for( i in 1:length(all_com_list)){ #i 
  for (j in 1:length(all_com_list)){ #j length(all_com_list)
    res[i,j]<- jaccard(all_com_list[[i]], all_com_list[[j]])
  }
}
rownames(res) <- names(all_com_list)
colnames(res) <- names(all_com_list)                     
```
look at the similiarty of the liver communities
```{r eval=FALSE}
res[grep("Liver",colnames(res)),grep("Liver",colnames(res))]
```

```{r eval=FALSE}
tissue_all <- str_split(rownames(res), "@" , simplify = TRUE)[,1]
sex_all <- str_split(rownames(res), "@" , simplify = TRUE)[,2]
com_all <- str_split(rownames(res), "@" , simplify = TRUE)[,3]

png("~/results/alpaca/alpaca_pathway_analysis/230220_jaccard_all_com_genes_clustering.png",width=20,height=10,units="in",res=1200)
set.seed(105)
row_ha<-  HeatmapAnnotation(Tissue=tissue_all, Sex=sex_all, Community_Number= com_all)
column_dend <- hclust(dist(t(res)), method = "ward.D2")
heatmap_v2 <- Heatmap(matrix(nrow = 0, ncol = ncol(res)),
          cluster_columns = column_dend, 
          show_column_names = FALSE,  
          show_row_names = FALSE, 
          top_annotation = row_ha) 
  draw(heatmap_v2,  annotation_legend_side = "bottom")
dev.off()
```

need to apply this to all the communities but core genes
```{r eval=FALSE}
female_core_list <- list()
male_core_list<- list()
for ( i in 1:length(tissues_wo_ws)){
  tissue_female <- gene_core_list(tissues_wo_ws[i], "female")
  female_core_list <- append(female_core_list, tissue_female)
  
  tissue_male <- gene_core_list(tissues_wo_ws[i], "male")
  male_core_list <- append(male_core_list, tissue_male)
}
#females
#males
```

```{r eval=FALSE}
all_core_list<- append(female_core_list, male_core_list)
```

```{r eval=FALSE}
res_core<- matrix(nrow= length(all_core_list), ncol= length(all_core_list))
for( i in 1:length(all_core_list)){ #i 
  for (j in 1:length(all_core_list)){ #j length(all_com_list)
    res_core[i,j]<- jaccard(all_core_list[[i]], all_core_list[[j]])
  }
}
rownames(res_core) <- names(all_core_list)
colnames(res_core) <- names(all_core_list)                     
```

```{r eval=FALSE}
tissue_all <- str_split(rownames(res_core), "@" , simplify = TRUE)[,1]
sex_all <- str_split(rownames(res_core), "@" , simplify = TRUE)[,2]
com_all <- str_split(rownames(res_core), "@" , simplify = TRUE)[,3]
png("~/results/alpaca/alpaca_pathway_analysis/230220_jaccard_core_com_genes_clustering.png",width=20,height=10,units="in",res=1200)
set.seed(105)
row_ha<-  HeatmapAnnotation(Tissue=tissue_all, Sex=sex_all, Community_Number= com_all)
  column_dend <-  hclust(dist(t(res_core)), method = "ward.D2")
heatmap_v2 <- Heatmap(matrix(nrow = 0, ncol = ncol(res_core)),
          cluster_columns = column_dend, 
          show_column_names = FALSE,  
          show_row_names = FALSE, 
          top_annotation = row_ha) 
draw(heatmap_v2,  annotation_legend_side = "bottom")
dev.off()
```


Just look at community number one across communities
```{r eval=FALSE}
res_core_1_3 <- res_core[com_all == 1, com_all == 1]
tissue_1_3 <- str_split(rownames(res_core_1_3), "@" , simplify = TRUE)[,1]
sex_1_3 <- str_split(rownames(res_core_1_3), "@" , simplify = TRUE)[,2]
#com_1_3 <- str_split(rownames(res_core_1_3), "@" , simplify = TRUE)[,3]

png("~/results/alpaca/alpaca_pathway_analysis/230220_jaccard_core_1_com_genes_clustering.png",width=20,height=10,units="in",res=1200)
set.seed(105)
row_ha<-  HeatmapAnnotation(Tissue=tissue_1_3, Sex=sex_1_3)
  column_dend <-  hclust(dist(t(res_core_1_3)), method = "ward.D2")
heatmap_v2 <- Heatmap(matrix(nrow = 0, ncol = ncol(res_core_1_3)),
          cluster_columns = column_dend, 
          top_annotation = row_ha) 
draw(heatmap_v2,  annotation_legend_side = "bottom")
dev.off()
```

look at the different brain regions
```{r eval=FALSE}
res_core_brain <- res_core_1_3[ grep("Brain", tissue_1_3), grep("Brain", tissue_1_3)]
tissue_brain<- str_split(rownames(res_core_brain), "@" , simplify = TRUE)[,1]
sex_brain <- str_split(rownames(res_core_brain), "@" , simplify = TRUE)[,2]
#com_1_3 <- str_split(rownames(res_core_1_3), "@" , simplify = TRUE)[,3]

png("~/results/alpaca/alpaca_pathway_analysis/230220_jaccard_core_com_genes_clustering_brain_regions.png",width=20,height=10,units="in",res=1200)
set.seed(120)
row_ha<-  HeatmapAnnotation(Tissue=tissue_brain, Sex=sex_brain)
  column_dend <-  hclust(dist(t(res_core_brain)), method = "ward.D2")
heatmap_v2 <- Heatmap(matrix(nrow = 0, ncol = ncol(res_core_brain)),
          cluster_columns = column_dend, 
          top_annotation = row_ha) 
draw(heatmap_v2,  annotation_legend_side = "bottom")
dev.off()
```

try the drug metabolism tissues only
```{r eval=FALSE}
res_core_drug_tissues <- res_core[ grep(paste(c("Liver", "Small_Intestine_-_Terminal_Ileum",  "Whole_Blood", "Kidney_-_Cortex"), collapse =  "|"), tissue_all), grep(paste(c("Liver", "Small_Intestine_-_Terminal_Ileum",  "Whole_Blood", "Kidney_-_Cortex"), collapse =  "|"), tissue_all)]
tissue_drug<- str_split(rownames(res_core_drug_tissues), "@" , simplify = TRUE)[,1]
sex_drug <- str_split(rownames(res_core_drug_tissues), "@" , simplify = TRUE)[,2]
com_drug <- str_split(rownames(res_core_drug_tissues), "@" , simplify = TRUE)[,3]

png("~/results/alpaca/alpaca_pathway_analysis/230220_jaccard_core_com_genes_clustering_drug_metabolism_tissues.png",width=20,height=10,units="in",res=1200)
set.seed(110)
row_ha<-  HeatmapAnnotation(Tissue=tissue_drug, Sex=sex_drug, Community= com_drug)
  column_dend <-  hclust(dist(t(res_core_drug_tissues)), method = "ward.D2")
heatmap_v2 <- Heatmap(matrix(nrow = 0, ncol = ncol(res_core_drug_tissues)),
          cluster_columns = column_dend, 
          top_annotation = row_ha) 
draw(heatmap_v2,  annotation_legend_side = "bottom")
dev.off()
```

community one is clustering for both sex and tissue

# END
    Location of final scripts:
    src
    Location of data produced:
    ~/results/alpaca/
    Dates when operations were done:
    230206
    
## Versions
```{r}
sessionInfo()
```



