---
title: "230206_alpaca_results_exploration"
author: "Jennifer Fisher"
date: "2023-02-06"
output:
  html_document:
    toc: true
    theme: united
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# START 
    Goal/Purpose of operations: 
    I am going to be looking at the results from Alpaca for the sex-specific networks across the 43 different tissue s
    
    Finished psedocode on:
    230206
    
    System which operations were done on: 
    my laptop
    
    GitHub Repo:
    230321_JLF_Sex_bias_adverse_events
    
    Directory of operations:
    ~/Documents/230321_JLF_Sex_bias_adverse_events
    
    Scripts being edited for operations:
    NA
    
    Data being used:
    results/alpaca/alpaca_runs
    
    Papers and tools: 
    NA

# STEPS

### Set working directory 

```{r}
library(gprofiler2)
library(viridis)
library(RColorBrewer)
library(ComplexHeatmap)
library(GOSemSim)
```


```{r}
metadata <- readRDS("~/data/metadata_gtex_filter_samples.rds")
```

```{r}
tissues <- unique(metadata$gtex.smtsd)
#remove cell tissues
tissues<- tissues[! grepl("Cell", tissues)]
```
note processed data had the white sapce removed 
```{r}
tissues_wo_ws <- gsub(" ", "_", tissues, fixed = TRUE)
```


```{r eval= FALSE}
for (j in 1:length(tissues_wo_ws)){
  alpaca_pathway_wrapper(tissues_wo_ws[j], "female")
  alpaca_pathway_wrapper(tissues_wo_ws[j], "male")
}
```

# top pathways

```{r}
all_files<- list.files("~/results/alpaca/alpaca_pathway_analysis")
pathway_files <- all_files[grep("pathways.csv", all_files )]
```

```{r}
all_pathway_res<- read.csv2(paste0("~/results/alpaca/alpaca_pathway_analysis/", pathway_files[1]))
for (j in 2:length(pathway_files)){
  group <- read.csv2(paste0("~/results/alpaca/alpaca_pathway_analysis/", pathway_files[j]))
  all_pathway_res<- rbind(all_pathway_res, group)
}
saveRDS(all_pathway_res, "~/results/alpaca/alpaca_pathway_analysis/all_pathways.rds")
```

```{r}
trim_1<- str_split(all_pathway_res$set, "_female", simplify = TRUE)[,1]
trim_2<- str_split(trim_1, "_male", simplify = TRUE)[,1]
all_pathway_res$tissue<- trim_2
```

```{r}
sex_group<- grepl("female", all_pathway_res$set)
all_pathway_res$sex <- ifelse(sex_group, "female", "male")
```
```{r}
all_pathway_res$comm_number <- as.numeric(gsub(".*?([0-9]+).*", "\\1", all_pathway_res$set))   
```

```{r}
saveRDS(all_pathway_res, "~/results/alpaca/alpaca_pathway_analysis/all_pathways.rds")
```

```{r}
all_pathway_res_filter<- cbind(all_pathway_res$term_name,all_pathway_res$source, all_pathway_res$term_id, all_pathway_res$tissue)
colnames(all_pathway_res_filter)<- c("term_name", "source", "id", "tissue")
all_pathway_res_filter<- all_pathway_res_filter[!duplicated(all_pathway_res_filter),]

all_pathway_res_filter_KEGG <- all_pathway_res_filter[all_pathway_res_filter[,2] %in% c("GO:BP"),]
test<- as.data.frame(table(all_pathway_res_filter_KEGG[,3]))
```

```{r}
go_bp_all_tissues<- test[test$Freq == 43,1]
```



```{r}
hsGO <- godata('org.Hs.eg.db', ont="BP")
```



```{r}
go_sim <- mgoSim(as.character(go_bp_all_tissues), as.character(go_bp_all_tissues), semData=hsGO, measure="Wang", combine=NULL)
#get the parent terms from rrvgo for the pathways
res <- reduceSimMatrix(go_sim , threshold = 0.85)
res_v2<- res[match(colnames(go_sim ), res$go),]
```


```{r}
sem_res <- as.data.frame(table(res_v2$parentTerm))
```

```{r}
sem_res$Var1<- factor(sem_res$Var1, levels=sem_res$Var1[order(sem_res$Freq)] )
ggplot(sem_res, aes(y= Var1, x= Freq, label=Freq)) + geom_bar(stat="identity", fill = "white", color= "black")+ 
    geom_text(size = 10, position = position_stack(vjust = 0.9)) + ylab("Parent BP terms") + xlab("number of GO:BP terms enriched once\nin all tissues" ) +theme(text = element_text(size = 10,  face="bold"))
```


```{r}
all_pathway_res[grep("cytochrome", all_pathway_res$term_name),]
```


```{r}
all_pathway_res[grep("p450", all_pathway_res$term_name),]
```

```{r}
all_pathway_res[grep("xenobiotic", all_pathway_res$term_name),]
```
small intestine and whole blood are interesting. no chromchrome p450 genes

```{r}
all_pathway_res[grep("small molecule", all_pathway_res$term_name),]
```



drug metabolism- tissue: liver, kidney, intestine ( small and large), lung, adrenal, blood, and skin

Skin_-_Sun_Exposed_(Lower_leg), Skin_-_Not_Sun_Exposed_(Suprapubic), Lung, Adrenal_Gland,

Liver, Small_Intestine_-_Terminal_Ileum,  Whole_Blood, Kidney_-_Cortex

```{r}
all_pathway_res_drug_tissues<- all_pathway_res[all_pathway_res$tissue %in% c("Liver", "Small_Intestine_-_Terminal_Ileum",  "Whole_Blood", "Kidney_-_Cortex"),]
```



```{r}

all_pathway_res_filter<- cbind(all_pathway_res_drug_tissues$term_name,all_pathway_res_drug_tissues$source, all_pathway_res_drug_tissues$term_id, all_pathway_res_drug_tissues$tissue)
colnames(all_pathway_res_filter)<- c("term_name", "source", "id", "tissue")
all_pathway_res_filter<- all_pathway_res_filter[!duplicated(all_pathway_res_filter),]

all_pathway_res_filter_KEGG <- all_pathway_res_filter[all_pathway_res_filter[,2] %in% c("GO:BP"),]
test<- as.data.frame(table(all_pathway_res_filter_KEGG[,3]))
```

```{r}
go_bp_all_tissues<- test[test$Freq == 4,1]
```

```{r}
go_sim <- mgoSim(as.character(go_bp_all_tissues), as.character(go_bp_all_tissues), semData=hsGO, measure="Wang", combine=NULL)
#get the parent terms from rrvgo for the pathways
res <- reduceSimMatrix(go_sim , threshold = 0.9)
res_v2<- res[match(colnames(go_sim ), res$go),]
```


```{r}
sem_res <- as.data.frame(table(res_v2$parentTerm))
```

```{r}
sem_res$Var1<- factor(sem_res$Var1, levels=sem_res$Var1[order(sem_res$Freq)] )
ggplot(sem_res, aes(y= Var1, x= Freq, label=Freq)) + geom_bar(stat="identity", fill = "white", color= "black")+ 
    geom_text(size = 10, position = position_stack(vjust = 0.9)) + ylab("Parent BP terms") + xlab("number of GO:BP terms enriched once\nin Liver, Small Intestine- Terminal Ileum, Whole Blood, & Kidney -Cortex" ) +theme(text = element_text(size = 10,  face="bold"))
```



# drug genes
KEGG_DRUG_METABOLISM_CYTOCHROME_P450.v2022.1.Hs.gmt contain a list of drug metabolism genes
```{r}
drug_genes <- c("ADH1A"	,"ADH1B", "ADH1C", "ADH4",	"ADH5",	"ADH6",	"ADH7",	
                "ALDH1A3",	"ALDH3A1",	"ALDH3B1", "ALDH3B2", "AOX1", "CYP1A2",
                "CYP2A13", "CYP2A6", "CYP2A7",	"CYP2B6",	"CYP2C18",	"CYP2C19",
                "CYP2C8",	"CYP2C9",	"CYP2D6",	"CYP2E1",	"CYP3A4",	"CYP3A43",	"CYP3A5",
                "CYP3A7",	"FMO1",	"FMO2",	"FMO3", "FMO4",	"FMO5",  "GSTA1", "GSTA2", 	"GSTA3",	"GSTA4",
                "GSTA5",	"GSTK1",	"GSTM1",	"GSTM2",	"GSTM3",	"GSTM4",	"GSTM5",	"GSTO1",
                "GSTO2",	"GSTP1",	"GSTT1",	"GSTT2",	"GSTZ1",	"MAOA",	"MAOB",	"MGST1",
                "MGST2",	"MGST3",	"UGT1A1",	"UGT1A10",	"UGT1A3",	"UGT1A4",
                "UGT1A5",	"UGT1A6",	"UGT1A7",	"UGT1A8", "UGT1A9",	"UGT2A1",	"UGT2A3", "UGT2B10",
                "UGT2B11",	"UGT2B15",	"UGT2B17",	"UGT2B28",	"UGT2B4",	"UGT2B7")
```


```{r}
#alpaca_gene_set_analysis( "Liver", genes="drugs", gene_set_name= "Drug Metabolism Genes", path="~/results/alpaca/drug_metabolism_genes/")
#alpaca_gene_set_analysis( "Lung", genes="drugs", gene_set_name= "Drug Metabolism Genes", path="~/results/alpaca/drug_metabolism_genes/")
```

```{r}
test_p<- c()
test_change <- c()
for (j in 1:length(tissues_wo_ws)){ #
   test <- alpaca_gene_set_analysis( tissues_wo_ws[j], genes="drugs", gene_set_name= "Drug Metabolism Genes", path="~/results/alpaca/drug_metabolism_genes/")
  test_p[j]<- test[1]
  test_change[j]<- test[2]
}
```

```{r}
sig_drug_gene_set <- as.data.frame(cbind(tissues_wo_ws, unlist(test_p), as.numeric(unlist(test_change))))
colnames(sig_drug_gene_set)<- c("Tissue", "p_value", "log2fold_Change")
```

```{r}
sig_drug_gene_set$neg_log_p <- -log(as.numeric(sig_drug_gene_set$p_value))
sig_drug_gene_set$log2fold_Change<- as.numeric(sig_drug_gene_set$log2fold_Change)
```

```{r}
p_value_cut<- -log(0.05/43)
sig_drug_gene_set$Tissue_2 <- ifelse(sig_drug_gene_set$neg_log_p > p_value_cut,sig_drug_gene_set$Tissue, NA )

ggplot(sig_drug_gene_set, aes(x=log2fold_Change, y= neg_log_p, label=Tissue_2  ))+ geom_point()+ xlab("log(Fold Change)") + ylab("-log(p-value)") + geom_hline(yintercept=p_value_cut, linetype="dashed", color = "black") + 
    geom_text(size = 5, position = position_stack(vjust = 0.95)) + xlim(c(-3, 3))
```

```{r}
sig_drug_gene_set$Tissue<- factor(sig_drug_gene_set$Tissue, levels= sig_drug_gene_set$Tissue[order(sig_drug_gene_set$neg_log_p)])
ggplot(sig_drug_gene_set, aes(y=Tissue, x= neg_log_p, fill= log2fold_Change))+ geom_bar(stat="identity",  color= "black") + scale_fill_gradient2(low= "#21908CFF", mid= "white", high= "#440154FF")  +theme(text = element_text(size = 20,  face="bold"))+ geom_vline(xintercept=p_value_cut, linetype="dashed", color = "black")+ xlab("-log(p-value)")
ggsave("~/results/alpaca/drug_metabolism_genes/230220_differential_mod_drug_genes_wilcox_test.png", height=12, width=10)
```

#make new function for jaccard similarity


need to apply this to all the communities
```{r}
female_com_list <- list()
male_com_list<- list()
for ( i in 1:length(tissues_wo_ws)){
  tissue_female <- gene_com_list(tissues_wo_ws[i], "female")
  female_com_list <- append(female_com_list, tissue_female)
  
  tissue_male <- gene_com_list(tissues_wo_ws[i], "male")
  male_com_list <- append(male_com_list, tissue_male)
}
#females
#males
```

```{r}
all_com_list<- append(female_com_list, male_com_list)
```

```{r}
res<- matrix(nrow= length(all_com_list), ncol= length(all_com_list))
for( i in 1:length(all_com_list)){ #i 
  for (j in 1:length(all_com_list)){ #j length(all_com_list)
    res[i,j]<- jaccard(all_com_list[[i]], all_com_list[[j]])
  }
}
rownames(res) <- names(all_com_list)
colnames(res) <- names(all_com_list)                     
```


```{r}
res[grep("Liver",colnames(res)),grep("Liver",colnames(res))]
```



```{r}
tissue_all <- str_split(rownames(res), "@" , simplify = TRUE)[,1]
sex_all <- str_split(rownames(res), "@" , simplify = TRUE)[,2]
com_all <- str_split(rownames(res), "@" , simplify = TRUE)[,3]

png("~/results/alpaca/alpaca_pathway_analysis/230220_jaccard_all_com_genes_clustering.png",width=20,height=10,units="in",res=1200)
#Your heat-map code
set.seed(105)
row_ha<-  HeatmapAnnotation(Tissue=tissue_all, Sex=sex_all, Community_Number= com_all)
column_dend <- hclust(dist(t(res)), method = "ward.D2")
heatmap_v2 <- Heatmap(matrix(nrow = 0, ncol = ncol(res)),
          cluster_columns = column_dend, 
          show_column_names = FALSE,  
          show_row_names = FALSE, 
          top_annotation = row_ha) 
  draw(heatmap_v2,  annotation_legend_side = "bottom")
dev.off()
```


need to apply this to all the communities
```{r}
female_core_list <- list()
male_core_list<- list()
for ( i in 1:length(tissues_wo_ws)){
  tissue_female <- gene_core_list(tissues_wo_ws[i], "female")
  female_core_list <- append(female_core_list, tissue_female)
  
  tissue_male <- gene_core_list(tissues_wo_ws[i], "male")
  male_core_list <- append(male_core_list, tissue_male)
}
#females
#males
```

```{r}
all_core_list<- append(female_core_list, male_core_list)
```

```{r}
res_core<- matrix(nrow= length(all_core_list), ncol= length(all_core_list))
for( i in 1:length(all_core_list)){ #i 
  for (j in 1:length(all_core_list)){ #j length(all_com_list)
    res_core[i,j]<- jaccard(all_core_list[[i]], all_core_list[[j]])
  }
}
rownames(res_core) <- names(all_core_list)
colnames(res_core) <- names(all_core_list)                     
```

```{r}
tissue_all <- str_split(rownames(res_core), "@" , simplify = TRUE)[,1]
sex_all <- str_split(rownames(res_core), "@" , simplify = TRUE)[,2]
com_all <- str_split(rownames(res_core), "@" , simplify = TRUE)[,3]
png("~/results/alpaca/alpaca_pathway_analysis/230220_jaccard_core_com_genes_clustering.png",width=20,height=10,units="in",res=1200)
set.seed(105)
row_ha<-  HeatmapAnnotation(Tissue=tissue_all, Sex=sex_all, Community_Number= com_all)
  column_dend <-  hclust(dist(t(res_core)), method = "ward.D2")
heatmap_v2 <- Heatmap(matrix(nrow = 0, ncol = ncol(res_core)),
          cluster_columns = column_dend, 
          show_column_names = FALSE,  
          show_row_names = FALSE, 
          top_annotation = row_ha) 
draw(heatmap_v2,  annotation_legend_side = "bottom")
dev.off()
```



```{r}

res_core_1_3 <- res_core[com_all == 1, com_all == 1]
tissue_1_3 <- str_split(rownames(res_core_1_3), "@" , simplify = TRUE)[,1]
sex_1_3 <- str_split(rownames(res_core_1_3), "@" , simplify = TRUE)[,2]
#com_1_3 <- str_split(rownames(res_core_1_3), "@" , simplify = TRUE)[,3]

#png("~/results/alpaca/alpaca_pathway_analysis/230220_jaccard_core_com_genes_clustering.png",width=20,height=10,units="in",res=1200)
set.seed(105)
row_ha<-  HeatmapAnnotation(Tissue=tissue_1_3, Sex=sex_1_3)
  column_dend <-  hclust(dist(t(res_core_1_3)), method = "ward.D2")
heatmap_v2 <- Heatmap(matrix(nrow = 0, ncol = ncol(res_core_1_3)),
          cluster_columns = column_dend, 
          top_annotation = row_ha) 
draw(heatmap_v2,  annotation_legend_side = "bottom")
#dev.off()
```


```{r}
res_core_brain <- res_core_1_3[ grep("Brain", tissue_1_3), grep("Brain", tissue_1_3)]
tissue_brain<- str_split(rownames(res_core_brain), "@" , simplify = TRUE)[,1]
sex_brain <- str_split(rownames(res_core_brain), "@" , simplify = TRUE)[,2]
#com_1_3 <- str_split(rownames(res_core_1_3), "@" , simplify = TRUE)[,3]

#png("~/results/alpaca/alpaca_pathway_analysis/230220_jaccard_core_com_genes_clustering.png",width=20,height=10,units="in",res=1200)
set.seed(120)
row_ha<-  HeatmapAnnotation(Tissue=tissue_brain, Sex=sex_brain)
  column_dend <-  hclust(dist(t(res_core_brain)), method = "ward.D2")
heatmap_v2 <- Heatmap(matrix(nrow = 0, ncol = ncol(res_core_brain)),
          cluster_columns = column_dend, 
          top_annotation = row_ha) 
draw(heatmap_v2,  annotation_legend_side = "bottom")
#dev.off()
```

try the drug metabolism tissues only
```{r}

#all_pathway_res_drug_tissues<- all_pathway_res[all_pathway_res$tissue %in% c("Liver", "Small_Intestine_-_Terminal_Ileum",  "Whole_Blood", "Kidney_-_Cortex"),]

res_core_drug_tissues <- res_core[ grep(paste(c("Liver", "Small_Intestine_-_Terminal_Ileum",  "Whole_Blood", "Kidney_-_Cortex"), collapse =  "|"), tissue_all), grep(paste(c("Liver", "Small_Intestine_-_Terminal_Ileum",  "Whole_Blood", "Kidney_-_Cortex"), collapse =  "|"), tissue_all)]
tissue_drug<- str_split(rownames(res_core_drug_tissues), "@" , simplify = TRUE)[,1]
sex_drug <- str_split(rownames(res_core_drug_tissues), "@" , simplify = TRUE)[,2]
com_drug <- str_split(rownames(res_core_drug_tissues), "@" , simplify = TRUE)[,3]

#png("~/results/alpaca/alpaca_pathway_analysis/230220_jaccard_core_com_genes_clustering.png",width=20,height=10,units="in",res=1200)
set.seed(110)
row_ha<-  HeatmapAnnotation(Tissue=tissue_drug, Sex=sex_drug, Community= com_drug)
  column_dend <-  hclust(dist(t(res_core_drug_tissues)), method = "ward.D2")
heatmap_v2 <- Heatmap(matrix(nrow = 0, ncol = ncol(res_core_drug_tissues)),
          cluster_columns = column_dend, 
          top_annotation = row_ha) 
draw(heatmap_v2,  annotation_legend_side = "bottom")
#dev.off()
```

communitiy one is cllustering for both sexx and tissue

```{r}
liver_pathway_drug_com <- all_pathway_res[all_pathway_res$set %in% c("Liver_female_3_community", "Liver_male_2_community"),]
```

```{r}
go_bp_3_2_liver<- unique(liver_pathway_drug_com$term_id[liver_pathway_drug_com$source == "GO:BP"])
```

```{r}
counts<- as.data.frame(table(liver_pathway_drug_com$term_name))
```

```{r}
#counts$Var1[counts$Freq == 2]
```


```{r}
go_sim <- mgoSim(as.character(go_bp_3_2_liver), as.character(go_bp_3_2_liver), semData=hsGO, measure="Wang", combine=NULL)
#get the parent terms from rrvgo for the pathways
res <- reduceSimMatrix(go_sim , threshold = 0.6)
res_v2<- res[match(colnames(go_sim ), res$go),]
```

```{r}
#all_pathway_res[all_pathway_res$set %in% c("Liver_female_3_community", "Liver_male_5_community"),]
go_term_heatmap(all_pathway_res[all_pathway_res$set=="Liver_female_3_community",], all_pathway_res[all_pathway_res$set=="Liver_male_3_community",], threshold = 0.9, file_name= "~/results/alpaca/alpaca_pathway_analysis/liver_female_3_male_2_go_sem_bp.csv" )
```


```{r}
unique(liver_female_3_male_5_go_sem_bp$parentTerm)
```
```{r}
sim_go_groups<- c()
enriched_group<- c()
for (i in 1:10){
  female_3<- liver_female_3_male_5_go_sem_bp$term[liver_female_3_male_5_go_sem_bp$parentTerm == unique(liver_female_3_male_5_go_sem_bp$parentTerm)[i] & liver_female_3_male_5_go_sem_bp$female_drugs]
  male_5<- liver_female_3_male_5_go_sem_bp$term[liver_female_3_male_5_go_sem_bp$parentTerm == unique(liver_female_3_male_5_go_sem_bp$parentTerm)[i] & liver_female_3_male_5_go_sem_bp$male_drugs]
  sim_go_groups[i]<- jaccard(female_3, male_5)
  enriched_group[i]<- length(female_3) > length(male_5)
}
names(sim_go_groups)<- unique(liver_female_3_male_5_go_sem_bp$parentTerm)
```

```{r}
sim_go_groups<- as.data.frame(sim_go_groups)
```

```{r}
sim_go_groups$Jaccard<- sim_go_groups$sim_go_groups
sim_go_groups$sim_go_groups<- rownames(sim_go_groups)
sim_go_groups$sex<- ifelse(enriched_group, "female", "male")
```
```{r}
#fill = "#0D0887B3",
sim_go_groups$sim_go_groups <- factor(sim_go_groups$sim_go_groups, levels= sim_go_groups$sim_go_groups[order(sim_go_groups$Jaccard)])
ggplot(sim_go_groups, aes(x=Jaccard, y=sim_go_groups, fill =sex )) + geom_bar(stat="identity",  color= "black", alpha=0.7) + scale_fill_manual(values= c("#0D0887B3","#F0F921B3" ))
```


```{r}
Whole_Blood_male_core_genes <- readRDS("~/results/alpaca/alpaca_core_gene_lists/Whole_Blood_male_core_genes.rds")
Whole_Blood_female_core_genes <- readRDS("~/results/alpaca/alpaca_core_gene_lists/Whole_Blood_female_core_genes.rds")
```

```{r}
table(Whole_Blood_male_core_genes$symbol %in% drug_genes)
table(Whole_Blood_female_core_genes$symbol %in% drug_genes)
```


```{r}
> `Small_Intestine_-_Terminal_Ileum_male_core_genes` <- readRDS("~/results/alpaca/alpaca_core_gene_lists/Small_Intestine_-_Terminal_Ileum_male_core_genes.rds")
> `Small_Intestine_-_Terminal_Ileum_female_core_genes` <- readRDS("~/results/alpaca/alpaca_core_gene_lists/Small_Intestine_-_Terminal_Ileum_female_core_genes.rds")
```


```{r}
table(`Small_Intestine_-_Terminal_Ileum_male_core_genes`$symbol %in% drug_genes)
table(`Small_Intestine_-_Terminal_Ileum_male_core_genes`$symbol %in% drug_genes)
```


```{r}
scores_file <- paste0("~/results/alpaca/alpaca_run/", tissues_wo_ws[41], "__ALPACA_scores.txt")
comms_file <- paste0("~/results/alpaca/alpaca_run/", tissues_wo_ws[41], "__ALPACA_final_memb.txt")
  file_scores <- read.delim(scores_file, header=FALSE)
  file_comms<- read.delim(comms_file, header=FALSE)
test<- alpaca_output_dw_all(file_scores, file_comms)
```

```{r}
drug_subset <- test[test$symbol %in% drug_genes,]
```

```{r}
drug_subset_list <- factor(drug_subset$com, levels=1:length(unique(test$com)))
drug_gene_count<- as.data.frame(table(drug_subset_list))
drug_gene_count$drug_subset_list <-  paste0( "female", drug_gene_count$drug_subset_list)
ggplot(drug_gene_count, aes(x= drug_subset_list, y= Freq, label=Freq)) + geom_bar(stat="identity", fill = "#0D0887B3", color= "black")+                                                     geom_text(size = 10, position = position_stack(vjust = 0.9)) + xlab("Communities") + ylab("Number of Drug Metabolism Genes")
```


```{r}
scores_file <- paste0("~/results/alpaca/alpaca_run/", tissues_wo_ws[41], "femalemale__ALPACA_scores.txt")
comms_file <- paste0("~/results/alpaca/alpaca_run/", tissues_wo_ws[41], "femalemale__ALPACA_final_memb.txt")
  file_scores <- read.delim(scores_file, header=FALSE)
  file_comms<- read.delim(comms_file, header=FALSE)
test2<- alpaca_output_dw_all(file_scores, file_comms)
```

```{r}
drug_subset2 <- test2[test2$symbol %in% drug_genes,]
```

```{r}
table(drug_subset2$com)
```

```{r}
drug_subset_list <- factor(drug_subset2$com, levels=1:length(unique(test2$com)))
drug_gene_count<- as.data.frame(table(drug_subset_list))
drug_gene_count$drug_subset_list <-  paste0( "male", drug_gene_count$drug_subset_list)
ggplot(drug_gene_count, aes(x= drug_subset_list, y= Freq, label=Freq)) + geom_bar(stat="identity", fill = "#F0F921B3", color= "black")+                                                     geom_text(size = 10, position = position_stack(vjust = 0.9)) + xlab("Communities") + ylab("Number of Drug Metabolism Genes")
```


```{r}
drug_subset2$sex<- rep("male", 64)
drug_subset$sex<- rep("female",64)
```

```{r}
liver_drug_subset<- rbind(drug_subset, drug_subset2)
```

```{r}
ggplot(liver_drug_subset, aes(y=score, x= sex, fill = sex)) + geom_violin() +geom_point()
```

```{r}
liver_drug_subset<- liver_drug_subset[order(liver_drug_subset$symbol),]
liver_drug_subset_female<- liver_drug_subset[liver_drug_subset$sex == "female",]
liver_drug_subset_male<- liver_drug_subset[liver_drug_subset$sex == "male",]
```



```{r}
res <- wilcox.test(liver_drug_subset_female$score, liver_drug_subset_male$score, paired = TRUE)
res
```
```{r}
ggplot(liver_drug_subset, aes(y=score, x= sex, fill = sex)) + geom_violin() +geom_point() + geom_text(x="female", y=0.002, label= "Paired Wilcoxon signed rank test\np-value = 1.949e-09") +ylab("Differential modularity score") + xlab("Sex") + ggtitle("Differential modularity scoreof drug metabolism genes\nin liver sex-specific gene regulatory networks") + scale_fill_viridis( alpha=0.7, option= "C", discrete = TRUE)
```



```{r}
res<- matrix(nrow= length(unique(test$com)), ncol= length(unique(test2$com)))
for( i in 1:length(unique(test$com))){ #i is female 
  for (j in 1:length(unique(test2$com)) ){ #j is male
    female_1<- test$symbol[test$com == i]
    male_1<- test2$symbol[test2$com == j]
    res[i,j]<- jaccard(female_1, male_1)
  }
}
rownames(res) <- paste0("female ", 1:length(unique(test$com)))
colnames(res) <- paste0("male ", 1:length(unique(test2$com)))                        
```


```{r}
col_fun <- colorRamp2(c(0,  1), c( "black", "yellow"))
Heatmap(res, nam= "Jaccard Similarity", col = col_fun, 
          clustering_distance_rows= "euclidean",
          clustering_distance_columns=  "euclidean",
          clustering_method_rows = "ward.D2" , column_names_rot = 45,
          clustering_method_columns="ward.D2", heatmap_legend_param = list( title_gp = gpar(fontsize = 15 , fontface = "bold"), labels_gp = gpar(fontsize = 12, fontface = "bold")))
```




```{r}
jaccard(female_1, male_5)
```



