---
title: "230313_sbae_drug_target"
author: "Jennifer Fisher"
date: "2023-03-13"
output:
  html_document:
    toc: true
    theme: united
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# START 
    Goal/Purpose of operations: 
    I am going to be looking at the enriched drug targets from the SBAE drugs and looking at gene expression and core genes of sex-specific communitites. 
    
    Finished psedocode on:
    230308
    
    System which operations were done on: 
    my laptop
    
    GitHub Repo:
    Sex_Bias_Adverse_Events
    
    Directory of operations:
    ~/Documents/Sex_Bias_Adverse_Events
    
    Scripts being edited for operations:
    NA
    
    Data being used:
    output/alpaca/alpaca_runs
    
    Papers and tools: 
    NA

# STEPS

### Set working directory 
```{r}
library(ggplot2)
library(circlize)
library(ComplexHeatmap)
library(drugbankR)
library(DBI)
library(hash)
library(org.Hs.eg.db)
library(EnsDb.Hsapiens.v75)
```

```{r}
drug_info_sbae_drugs <- readRDS("~/output/FARES_plots/drug_info_sbae_drugs.rds")
drug_info_all <- readRDS("~/output/FARES_plots/drug_info_all.rds")
sbae_targets_df_top <- readRDS("~/output/FARES_plots/sbae_targets_df_top.rds")
```



## Analysis

```{r}
table(sbae_targets_df_top$p_adj < 0.001)
```


```{r}
SBAE_drug_targets<- unfactor(sbae_targets_df_top$sbae_targets)
```


```{r}
all_targets <- unlist(str_split(drug_info_all[,2], ";"))
all_targets <- all_targets[!is.na(all_targets )]
all_targets <- all_targets[!all_targets == "NA"]
```

```{r}
all_targets_df<- as.data.frame(table(all_targets))
```

```{r}
all_drug_targets<- unfactor(all_targets_df$all_targets)
```


core gene analysis 

```{r}
#BA9_female_core_genes <- readRDS("~/output/alpaca/alpaca_core_gene_lists/Brain_-_Frontal_Cortex_(BA9)_female_core_genes.rds")
#BA9_male_core_genes <- readRDS("~/output/alpaca/alpaca_core_gene_lists/Brain_-_Frontal_Cortex_(BA9)_male_core_genes.rds")
#BA9_list <- unique(BA9_female_core_genes$symbol, BA9_male_core_genes$symbol)
#table(SBAE_drug_targets %in% BA9_list)
#table(all_drug_targets %in% BA9_list)
```


```{r}
drug_gene_core_test<- function( count, all_drug_targets_list, core_list, seed= 101){
  res<- c()
  set.seed(seed)
  for (i in 1:1000){
    #t<- table(grepl("",))
    sub<- sample( all_drug_targets_list, 84, replace = FALSE)

    res1 <- table(sub %in% core_list)
    res2<- res1[names(res1) == "TRUE"]
    if (length(res2) ==1){
      res[i]<-res2
    }else{
      res[i]<- 0
    }
    
  }
  #print(res)
  test<- wilcox.test(res, mu= count, alternative = "less")
  #print(count)
  #print(hist(res))
  plot_data<- as.data.frame(cbind(1:1000, res))
  plot1 <- ggplot(plot_data, aes(x=res)) + geom_histogram() + geom_vline(xintercept = count)
  
  res_list<- list(test$p.value, plot1)
  return(res_list)
}
```


```{r}
#drug_gene_core_test(5, all_drug_targets,BA9_list )
```


```{r}
files <- list.files(path = "~/output/alpaca/alpaca_core_gene_lists")
```


```{r}
tissues<- str_split(files, "_female", simplify = TRUE)[,1]
tissues<- str_split(tissues, "_male", simplify = TRUE)[,1]
tissues<- unique(tissues)
```


male core genes
```{r}
pvalue<- c()
plots<- list()
core_gene_count <- c()
core_gene_list <- list()
for( i in 1:length(tissues)){
  print(tissues[i])
  male_core_genes <- readRDS(paste("~/output/alpaca/alpaca_core_gene_lists/", tissues[i],"_male_core_genes.rds", sep=""))
  
  sbae_res <- table(SBAE_drug_targets %in% unique(male_core_genes$symbol))
  
  sbae_count<- sbae_res[names(sbae_res) == "TRUE"]
  
  outcome<- drug_gene_core_test(sbae_count, all_drug_targets, core_list)
  
  pvalue[i]<- outcome[1]
  plots[i]<- outcome[2]
  core_gene_count[i]<- sbae_count
  core_gene_list[[i]] <- SBAE_drug_targets[SBAE_drug_targets %in% unique(male_core_genes$symbol)]
}
```


```{r}
p_adj<-  p.adjust(pvalue, method= "BH")
```

```{r}
male_sex_core_gene <- data.frame(tissue= tissues, count_core_gene= core_gene_count) #, p_value= pvalue, BH_adj_p= p_adj)
```

```{r}
male_sex_core_gene<- cbind(male_sex_core_gene, unlist(pvalue), unlist(p_adj ))
```

```{r}
male_sex_core_gene$fraction<- male_sex_core_gene$count_core_gene/84
```

```{r}
male_sex_core_gene_df_top<- male_sex_core_gene[order(male_sex_core_gene$count_core_gene),]
male_sex_core_gene_df_top$tissue<- factor(male_sex_core_gene_df_top$tissue, levels= male_sex_core_gene_df_top$tissue)
```

```{r}
male_sex_core_gene_df_top$sig <- ifelse(male_sex_core_gene_df_top$`unlist(p_adj)` < 0.05, "Sig.", "Not Sig.")
```

```{r}
male_sex_core_gene_df_top$tissue2 <- str_replace_all(male_sex_core_gene_df_top$tissue, "_", " ")
```

```{r}

male_sex_core_gene_df_top$tissue2<- factor(male_sex_core_gene_df_top$tissue2, levels= male_sex_core_gene_df_top$tissue2)
```

```{r}
ggplot(male_sex_core_gene_df_top, aes(y=tissue2, x= fraction , label= count_core_gene, fill=sig))+ geom_bar(stat="identity",color= "black") + 
    geom_text(size = 5, position = position_stack(vjust = 0.8)) + ylab("Tissue") + xlab("The fraction of SBAE drug targets\nthat are core genes" ) +theme(text = element_text(size = 20,  face="bold")) + scale_fill_viridis_d(option = "E", alpha= 0.7) +labs(fill="Significant?")
ggsave("~/output/alpaca/drug_target/230313_male_drug_target_core_gene_barplot.png", height=15, width = 10)
```

```{r}
male_drug_target_core_count <- as.data.frame(table(unlist(core_gene_list)))
```

make this plot

```{r}
male_drug_target_core_count<- male_drug_target_core_count[order(male_drug_target_core_count$Freq),]
male_drug_target_core_count$Var1 <- factor(male_drug_target_core_count$Var1, levels= male_drug_target_core_count$Var1)
```

```{r}
ggplot(male_drug_target_core_count, aes(y=Var1, x= Freq , label= Freq))+ geom_bar(stat="identity",color= "black", fill = "white") + 
    geom_text(size = 5, position = position_stack(vjust = 0.8)) + ylab("SBAE Drug Target") + xlab("The number of tissues with drug target as core gene" ) +theme(text = element_text(size = 20,  face="bold")) 
ggsave("~/output/alpaca/drug_target/230313_male_drug_target_core_gene_barplot_just_genes.png", height=15, width = 10)
```


female core genes

```{r}
pvalue<- c()
plots<- list()
core_gene_count <- c()
core_gene_list <- list()
for( i in 1:length(tissues)){
  print(tissues[i])
  female_core_genes <- readRDS(paste("~/output/alpaca/alpaca_core_gene_lists/", tissues[i],"_female_core_genes.rds", sep=""))
  
  sbae_res <- table(SBAE_drug_targets %in% unique(female_core_genes$symbol))
  
  sbae_count<- sbae_res[names(sbae_res) == "TRUE"]
  
  outcome<- drug_gene_core_test(sbae_count, all_drug_targets, core_list)
  
  pvalue[i]<- outcome[1]
  plots[i]<- outcome[2]
  core_gene_count[i]<- sbae_count
  core_gene_list[[i]] <- SBAE_drug_targets[SBAE_drug_targets %in% unique(female_core_genes$symbol)]
}
```


```{r}
p_adj<-  p.adjust(pvalue, method= "BH")
```

```{r}
female_sex_core_gene <- data.frame(tissue= tissues, count_core_gene= core_gene_count) #, p_value= pvalue, BH_adj_p= p_adj)
```

```{r}
female_sex_core_gene<- cbind(female_sex_core_gene, unlist(pvalue), unlist(p_adj ))
```

```{r}
female_sex_core_gene$fraction<- female_sex_core_gene$count_core_gene/84
```

```{r}
female_sex_core_gene_df_top<- female_sex_core_gene[order(female_sex_core_gene$count_core_gene),]
female_sex_core_gene_df_top$tissue<- factor(female_sex_core_gene_df_top$tissue, levels= female_sex_core_gene_df_top$tissue)
```

```{r}
female_sex_core_gene_df_top$sig <- ifelse(female_sex_core_gene_df_top$`unlist(p_adj)` < 0.05, "Sig.", "Not Sig.")
```

```{r}
female_sex_core_gene_df_top$tissue2 <- str_replace_all(female_sex_core_gene_df_top$tissue, "_", " ")
```

```{r}

female_sex_core_gene_df_top$tissue2<- factor(female_sex_core_gene_df_top$tissue2, levels= female_sex_core_gene_df_top$tissue2)
```

```{r}
ggplot(female_sex_core_gene_df_top, aes(y=tissue2, x= fraction , label= count_core_gene, fill=sig))+ geom_bar(stat="identity",color= "black") + 
    geom_text(size = 5, position = position_stack(vjust = 0.8)) + ylab("Tissue") + xlab("The fraction of SBAE drug targets\nthat are core genes") +theme(text = element_text(size = 20,  face="bold")) + scale_fill_viridis_d(option = "E", alpha= 0.7)+labs(fill="Significant?")
ggsave("~/output/alpaca/drug_target/230313_female_drug_target_core_gene_barplot.png", height=15, width = 10)
```

```{r}
female_drug_target_core_count <- as.data.frame(table(unlist(core_gene_list)))
```

make this plot

```{r}
female_drug_target_core_count<- female_drug_target_core_count[order(female_drug_target_core_count$Freq),]
female_drug_target_core_count$Var1 <- factor(female_drug_target_core_count$Var1, levels= female_drug_target_core_count$Var1)
```

```{r}
ggplot(female_drug_target_core_count, aes(y=Var1, x= Freq , label= Freq))+ geom_bar(stat="identity",color= "black", fill = "white") + 
    geom_text(size = 5, position = position_stack(vjust = 0.8)) + ylab("SBAE Drug Target") + xlab("The number of tissues with drug target as core gene" ) +theme(text = element_text(size = 20,  face="bold")) 
ggsave("~/output/alpaca/drug_target/230313_female_drug_target_core_gene_barplot_just_genes.png", height=15, width = 10)
```



Now look at sex-bias expression 

```{r}
sex_bias_genes_gtex_science_paper <- read.table("~/data/GTEx_Analysis_v8_sbgenes/signif.sbgenes.txt", header=TRUE)
```



```{r}
table(sex_bias_genes_gtex_science_paper$tissue)
```

```{r}
#for each tissue find the male and female biased genes
tissues<- unique(sex_bias_genes_gtex_science_paper$tissue)
male_tissues_names <- paste("M", tissues, sep="_")
female_tissues_names<- paste("F", tissues, sep="_")
for (i in 1:length(tissues)){
  assign(male_tissues_names[i], sex_bias_genes_gtex_science_paper[sex_bias_genes_gtex_science_paper$tissue ==tissues[i] & sex_bias_genes_gtex_science_paper$effsize < 0,])
  assign(female_tissues_names[i], sex_bias_genes_gtex_science_paper[sex_bias_genes_gtex_science_paper$tissue ==tissues[i] & sex_bias_genes_gtex_science_paper$effsize > 0,])
}
```

```{r}
both_sex_tissue_names<- c(male_tissues_names, female_tissues_names)

geneset<- data.frame(id =both_sex_tissue_names, name= both_sex_tissue_names, genes= I(rep(list(c("a","b")), 88)))
sex_genes <- both_sex_tissue_names
for (i in 1:length(both_sex_tissue_names)){
  data <- get(both_sex_tissue_names[i])
  trim_names<- strtrim(data[,1],15)
  assign( sex_genes[i], trim_names)
  
}

sex_gene_set <- list(M_Adipose_Subcutaneous= M_Adipose_Subcutaneous, M_Adipose_Visceral_Omentum= M_Adipose_Visceral_Omentum)
```

```{r}
for (j in 1:88){
  sex_gene_set[[j]]<- get(sex_genes[j])
  #names(sex_gene_set[j])<- sex_genes[j]
}
names(sex_gene_set)<-sex_genes
```


need to convert symbols to ensmble ids 

```{r}
SBAE_drug_targets_v2 <- gprofiler2::gconvert(SBAE_drug_targets, mthreshold= 1)
```


```{r}
all_drug_targets_v2<- gprofiler2::gconvert(all_drug_targets, mthreshold= 1)
```




```{r}
drug_gene_expression_test<- function( count, all_drug_targets_list, expression_list, seed= 101){
  res<- c()
  set.seed(seed)
  for (i in 1:1000){
    #t<- table(grepl("",))
    sub<- sample( all_drug_targets_list, 84, replace = FALSE)

    res1 <- table(sub %in% expression_list)
    res2<- res1[names(res1) == "TRUE"]
    if (length(res2) ==1){
      res[i]<-res2
    }else{
      res[i]<- 0
    }
    
  }
  #print(res)
  test<- wilcox.test(res, mu= count, alternative = "less")
  #print(count)
  #print(hist(res))
  plot_data<- as.data.frame(cbind(1:1000, res))
  plot1 <- ggplot(plot_data, aes(x=res)) + geom_histogram() + geom_vline(xintercept = count)
  
  res_list<- list(test$p.value, plot1)
  return(res_list)
}
```

```{r}
pvalue<- c()
plots<- list()
expression_count <- c()
expression_gene_list <- list()

for( i in 1:length(sex_gene_set)){
  print(names(sex_gene_set)[i])
  
  expression_list <- sex_gene_set[[i]]

  sbae_res <- table(SBAE_drug_targets_v2$target %in% expression_list)
  
  sbae_count<- sbae_res[names(sbae_res) == "TRUE"]
  print(sbae_count)
  sbae_count<- ifelse( length(sbae_count)== 0, 0, sbae_count)
  outcome<- drug_gene_expression_test(sbae_count, all_drug_targets_v2$target, expression_list)
  
  pvalue[i]<- outcome[1]
  plots[i]<- outcome[2]
  expression_count[i]<- sbae_count
  expression_gene_list[[i]] <- SBAE_drug_targets_v2$target[SBAE_drug_targets_v2$target %in% expression_list]
}
```

```{r}
p_adj<-  p.adjust(pvalue, method= "BH")
```

```{r}
sex_expression_gene <- data.frame(tissue= names(sex_gene_set), expression_count= expression_count) #, p_value= pvalue, BH_adj_p= p_adj)
```

```{r}
sex_expression_gene<- cbind(sex_expression_gene, unlist(pvalue), unlist(p_adj ))
```

```{r}
sex_expression_gene$fraction<- sex_expression_gene$expression_count/84
```



```{r}
sex_expression_gene$sig <- ifelse(sex_expression_gene$`unlist(p_adj)` < 0.05, "Sig.", "Not Sig.")
```

```{r}
female_sex_expression_gene<- sex_expression_gene[grep("F_", sex_expression_gene$tissue),]
```
```{r}
female_sex_expression_gene$tissue2 <- str_split(female_sex_expression_gene$tissue, "F_", simplify = TRUE)[,2]
female_sex_expression_gene$tissue2 <- str_replace_all(female_sex_expression_gene$tissue2, "_", " ")
```

```{r}
female_sex_expression_gene_df_top<- female_sex_expression_gene[order(female_sex_expression_gene$expression_count),]
female_sex_expression_gene_df_top$tissue2<- factor(female_sex_expression_gene$tissue2, levels= female_sex_expression_gene$tissue2)
```


```{r}
ggplot(female_sex_expression_gene_df_top, aes(y=tissue2, x= fraction , label= expression_count, fill=sig))+ geom_bar(stat="identity",color= "black") + 
    geom_text(size = 5, position = position_stack(vjust = 0.8)) + ylab("Tissue") + xlab("The fraction of SBAE drug targets that have\nfemale-bias expression" ) +theme(text = element_text(size = 18,  face="bold")) + scale_fill_viridis_d(option = "E", alpha= 0.7)+labs(fill="Significant?")
ggsave("~/output/FARES_plots/230313_drug_target_female_bias_expression_barplot.png", height=15, width = 10)
```


```{r}
male_sex_expression_gene<- sex_expression_gene[grep("M_", sex_expression_gene$tissue),]
```
```{r}
male_sex_expression_gene$tissue2 <- str_split(male_sex_expression_gene$tissue, "M_", simplify = TRUE)[,2]
male_sex_expression_gene$tissue2 <- str_replace_all(male_sex_expression_gene$tissue2, "_", " ")
```

```{r}
male_sex_expression_gene_df_top<- male_sex_expression_gene[order(male_sex_expression_gene$expression_count),]
male_sex_expression_gene_df_top$tissue2<- factor(male_sex_expression_gene$tissue2, levels= male_sex_expression_gene$tissue2)
```


```{r}
ggplot(male_sex_expression_gene_df_top, aes(y=tissue2, x= fraction , label= expression_count, fill=sig))+ geom_bar(stat="identity",color= "black") + 
    geom_text(size = 5, position = position_stack(vjust = 0.8)) + ylab("Tissue") + xlab("The fraction of SBAE drug targets that have\nmale-bias expression" ) +theme(text = element_text(size = 18,  face="bold")) + scale_fill_viridis_d(option = "E", alpha= 0.7) +labs(fill="Significant?")
ggsave("~/output/FARES_plots/230313_drug_target_male_bias_expression_barplot.png", height=15, width = 10)
```


```{r}
expression_drug_targets <- SBAE_drug_targets_v2[ SBAE_drug_targets_v2$target %in% unique(unlist(expression_gene_list)),]
```






```{r}
drug_genes <- c("ADH1A"	,"ADH1B", "ADH1C", "ADH4",	"ADH5",	"ADH6",	"ADH7",	
                "ALDH1A3",	"ALDH3A1",	"ALDH3B1", "ALDH3B2", "AOX1", "CYP1A2",
                "CYP2A13", "CYP2A6", "CYP2A7",	"CYP2B6",	"CYP2C18",	"CYP2C19",
                "CYP2C8",	"CYP2C9",	"CYP2D6",	"CYP2E1",	"CYP3A4",	"CYP3A43",	"CYP3A5",
                "CYP3A7",	"FMO1",	"FMO2",	"FMO3", "FMO4",	"FMO5",  "GSTA1", "GSTA2", 	"GSTA3",	"GSTA4",
                "GSTA5",	"GSTK1",	"GSTM1",	"GSTM2",	"GSTM3",	"GSTM4",	"GSTM5",	"GSTO1",
                "GSTO2",	"GSTP1",	"GSTT1",	"GSTT2",	"GSTZ1",	"MAOA",	"MAOB",	"MGST1",
                "MGST2",	"MGST3",	"UGT1A1",	"UGT1A10",	"UGT1A3",	"UGT1A4",
                "UGT1A5",	"UGT1A6",	"UGT1A7",	"UGT1A8", "UGT1A9",	"UGT2A1",	"UGT2A3", "UGT2B10",
                "UGT2B11",	"UGT2B15",	"UGT2B17",	"UGT2B28",	"UGT2B4",	"UGT2B7")
```


```{r}
drug_genes_v2 <- gprofiler2::gconvert(drug_genes, mthreshold= 1)
```

```{r}
table(drug_genes_v2$target %in% sex_gene_set$F_Liver)
```

```{r}
table(drug_genes_v2$target %in% sex_gene_set$M_Liver)
```

```{r}
drug_genes_v2$name[drug_genes_v2$target %in% sex_gene_set$F_Liver]
```

```{r}
drug_genes_v2$name[drug_genes_v2$target %in% sex_gene_set$M_Liver]
```





```{r}
drug_enzymes_gene_expression_test<- function( count, all_drug_targets_list, expression_list, seed= 101){
  res<- c()
  set.seed(seed)
  for (i in 1:1000){
    #t<- table(grepl("",))
    sub<- sample( all_drug_targets_list, 64, replace = FALSE)

    res1 <- table(sub %in% expression_list)
    res2<- res1[names(res1) == "TRUE"]
    if (length(res2) ==1){
      res[i]<-res2
    }else{
      res[i]<- 0
    }
    
  }
  #print(res)
  test<- wilcox.test(res, mu= count, alternative = "less")
  #print(count)
  #print(hist(res))
  plot_data<- as.data.frame(cbind(1:1000, res))
  plot1 <- ggplot(plot_data, aes(x=res)) + geom_histogram() + geom_vline(xintercept = count)
  
  res_list<- list(test$p.value, plot1)
  return(res_list)
}
```


```{r}
pvalue<- c()
plots<- list()
expression_count <- c()
expression_gene_list <- list()

for( i in 1:length(sex_gene_set)){
  print(names(sex_gene_set)[i])
  
  expression_list <- sex_gene_set[[i]]

  sbae_res <- table(SBAE_drug_targets_v2$target %in% expression_list)
  
  sbae_count<- sbae_res[names(sbae_res) == "TRUE"]
  print(sbae_count)
  sbae_count<- ifelse( length(sbae_count)== 0, 0, sbae_count)
  outcome<- drug_enzymes_gene_expression_test(sbae_count, all_drug_targets_v2$, expression_list)
  
  pvalue[i]<- outcome[1]
  plots[i]<- outcome[2]
  expression_count[i]<- sbae_count
  expression_gene_list[[i]] <- SBAE_drug_targets_v2$target[SBAE_drug_targets_v2$target %in% expression_list]
}
```
