---
title: "230308_drugbank_info"
author: "Jennifer Fisher"
date: "2023-03-08"
output:
  html_document:
    toc: true
    theme: united
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# START 
    Goal/Purpose of operations: 
    I am going to be looking at the results from FARES sex-bias adverse event but with the addition of the drug bank infomation
    
    Finished psedocode on:
    230308
    
    System which operations were done on: 
    my laptop
    
    GitHub Repo:
    230321_JLF_Sex_bias_adverse_events
    
    Docker: 
    jenfisher7/rstudio_sex_bias_drugs
    
    Directory of operations:
    ~/Documents/230321_JLF_Sex_bias_adverse_events
    
    Scripts being edited for operations:
    NA
    
    Data being used:
    ~/results/FARES_plots/drug_ae_res_info.rds - fisher extact test results from FARES database
    ~/data/full database.xml- drugbank database; version="5.1.8"
    
    Papers and tools: 
    NA

# STEPS

### Set working directory 
```{r}
source("~/src/sex_bias_drug_functions.R")
library(ggplot2)
library(circlize)
library(ComplexHeatmap)
library(drugbankR)
library(DBI)
library(hash)
library(org.Hs.eg.db)
```

```{r}
drug_ae_res_info <- readRDS("~/results/FARES_plots/drug_ae_res_info.rds")
```

## Analysis

```{r eval=FALSE}
drugbank_dataframe <- dbxml2df(xmlfile="~/data/full database.xml", version="5.1.8")
```
Extracting data for column names. This may take 20 minutes.
Successfully convert DrugBank database (xml file) into dataframe.

```{r eval=FALSE}
saveRDS(drugbank_dataframe, "~/data/drug_bank_1_5_8_df.rds")
```

```{r eval=FALSE}
drugbank_dataframe[1:10,]
```


```{r eval=FALSE}
setwd("~/data")
df2SQLite(dbdf=drugbank_dataframe, version="5.1.8")
```

identify drug information on only the significant drugs 
```{r}
drug_ae_res_info_sig <- drug_ae_res_info[drug_ae_res_info$BH < 0.05 & abs(drug_ae_res_info$logROR) > 1,]
```

```{r}
drug_ids <- unique(drug_ae_res_info_sig$drug)
```

```{r}
targets <- queryDB(drug_ids ,type = "getTargets", db_path = "~/data/drugbank_5.1.8.db") [,5]
print("done target")

carrier <- getCarriers(drug_ids, db_path = "~/data/drugbank_5.1.8.db")
print("done carrier")
  
transporter <- getTransporter(drug_ids, db_path = "~/data/drugbank_5.1.8.db")
print("done transporter")
  
enzyme <- getEnzymes(drug_ids, db_path = "~/data/drugbank_5.1.8.db")
print("done enzymes")
  
drug_info <- cbind(drug_ids, targets, enzyme, transporter, carrier)
```

```{r}
drug_info<- as.data.frame(drug_info)
```

get information for all the drugs in FARES database
```{r}
drug_ae_df_inital <- readRDS("~/data/drug_ae_df.rds")
```

```{r}
inital_drug_list <- unique(drug_ae_df_inital$drug)
```
repeat the process
```{r}
targets <- queryDB(inital_drug_list ,type = "getTargets", db_path = "~/data/drugbank_5.1.8.db") [,5]
print("done target")
  
carrier <- getCarriers(inital_drug_list, db_path = "~/data/drugbank_5.1.8.db")
print("done carrier")
  
transporter <- getTransporter(inital_drug_list, db_path = "~/data/drugbank_5.1.8.db")
print("done transporter")
  
enzyme <- getEnzymes(inital_drug_list, db_path = "~/data/drugbank_5.1.8.db")
print("done enzymes")
  
drug_info_all <- cbind(inital_drug_list, targets, enzyme, transporter, carrier)
```

```{r}
drug_info_all<- as.data.frame(drug_info_all)
```

permutation testing 

count table of drug target in sbaes
```{r}
sbae_targets <- unlist(str_split(drug_info[,2], ";"))
sbae_targets <- sbae_targets[!is.na(sbae_targets )]
sbae_targets <- sbae_targets[!sbae_targets == "NA"]
```

```{r}
sbae_targets_df<- as.data.frame(table(sbae_targets))
```

count table of drug target in all targets
```{r}
all_targets <- unlist(str_split(drug_info_all[,2], ";"))
all_targets <- all_targets[!is.na(all_targets )]
all_targets <- all_targets[!all_targets == "NA"]
```

```{r}
all_targets_df<- as.data.frame(table(all_targets))
```

```{r}
length(all_targets_df$all_targets)
```
2082- total targets

```{r}
not_sbae_targets_list <- all_targets_df$all_targets[!all_targets_df$all_targets %in% sbae_targets_df$sbae_targets]
```

add 0 counts to target targets that are not used by SBAE drugs
```{r}
count <- rep(0, length(not_sbae_targets_list))
other_targets <- cbind(as.character(not_sbae_targets_list), count)
colnames(other_targets) <- colnames(sbae_targets_df)
```

```{r}
sbae_targets_test_df<- rbind(sbae_targets_df, other_targets)
```

test all the drug targets in FARES for enrichment in SBAE drugs
```{r}
pvalue<- c()
plots<- list()
test<- list()
for ( j in 1:nrow(sbae_targets_test_df)){ #
  print(j)
  res <- drug_gene_test(as.character(sbae_targets_test_df[j,1]),as.numeric(sbae_targets_test_df[j,2]), 2, drug_info_all )
  pvalue[j]<- res[1]
  plots[j]<- res[2]
  test[j]<- res[3]
}
```

```{r}
names(test)<- sbae_targets_test_df[,1]
```

```{r}
test_df<- as.data.frame(test)
```

Plot the test results
```{r}
test_long_df <- pivot_longer(test_df, cols=everything(), names_to = "Drug_Target", values_to = "Count")
```

```{r}
sbae_targets_test_df$p_adj<-  p.adjust(pvalue, method= "BH")
```

```{r}
table(sbae_targets_test_df$p_adj < 0.001)
```

Focus on SBAE drug targets that are shared by 5 SBAE drugs
```{r}
sbae_targets_df_filter <- sbae_targets_test_df[as.numeric(sbae_targets_test_df$Freq) > 5,]
```

```{r}
dim(sbae_targets_df_filter)
```


```{r}
sbae_targets_df_top <-sbae_targets_df_filter[sbae_targets_df_filter$p_adj< 0.001,]
```

```{r}
dim(sbae_targets_df)
```

```{r}
dim(sbae_targets_df_top)
```


```{r}
test_long_df_top <- test_long_df[test_long_df$Drug_Target %in% sbae_targets_df_top$sbae_targets, ]
```


```{r}
sbae_counts<- c()
for(i in 1:nrow(test_long_df_top)){
  sbae_counts[i] <- sbae_targets_df_top$Freq[sbae_targets_df_top$sbae_targets == test_long_df_top$Drug_Target[i]]
}
```

```{r}
test_long_df_top$Drug_Target <- factor(test_long_df_top$Drug_Target, levels= sbae_targets_df_top$sbae_targets)
```


```{r}
test_long_df_top$sbae_number <- as.numeric(sbae_counts)
```

```{r}
test_long_df_top$Drug_Target <- factor(test_long_df_top$Drug_Target, levels =as.character(sbae_targets_df_top$sbae_targets)[order(-as.numeric(sbae_targets_df_top$Freq))])
```


```{r}
ggplot(test_long_df_top, aes(x=Count, y= Drug_Target ))+ geom_boxplot() + geom_point( aes(y=Drug_Target, x= sbae_number ), color= "forestgreen",size =5) +
  xlab("Number of drugs with drug target after random selection" ) +
  theme(text = element_text(size = 20,  face="bold")) + ylab("Drug Target")
ggsave("~/results/FARES_plots/230313_drug_target_box_plot.png", height=20, width = 10)
```


different plot looking at the fraction of drug targets
```{r}
sbae_targets_df_top$neg_log_p <- -log(sbae_targets_df_top$p_adj)
```


```{r}
sbae_targets_df_top <- sbae_targets_df_top[order(-as.numeric(sbae_targets_df_top$Freq)), ]
sbae_targets_df_top$sbae_targets <- factor(sbae_targets_df_top$sbae_targets, levels= sbae_targets_df_top$sbae_targets)
```

```{r}
all_counts<- c()
for (i in 1:nrow(sbae_targets_df_top)){
  all_counts[i] <- all_targets_df$Freq[unfactor(all_targets_df$all_targets) == sbae_targets_df_top$sbae_targets[i]]
}
```

```{r}
sbae_targets_df_top$all_count<- all_counts
```

```{r}
sbae_targets_df_top$fraction <- as.numeric(sbae_targets_df_top$Freq)/sbae_targets_df_top$all_count
```

```{r}
sbae_targets_df_top$sbae_fraction <- as.numeric(sbae_targets_df_top$Freq)/416 # number of sbae drugs 
```


```{r}
sbae_targets_df_top$all_fraction <- as.numeric(sbae_targets_df_top$all_count)/2945 # number of all. drugs
```


```{r}
table(sbae_targets_df_top$sbae_fraction > sbae_targets_df_top$all_fraction )
```

```{r}
fract_plot_data <- sbae_targets_df_top[, c(1,7,8)]
```

```{r}
fract_plot_data_long <- pivot_longer(fract_plot_data, col = !sbae_targets)
```


```{r}
ggplot(fract_plot_data_long, aes())
```



```{r}
ggplot(sbae_targets_df_top[1:30,], aes(y=sbae_targets, x= fraction , label= text))+ geom_bar(stat="identity",color= "black", fill= "white") + 
    geom_text(size = 8, position = position_stack(vjust = 0.8)) + ylab("Drug Target") + xlab("The fraction of SBAE drugs over all drugs\nwith same target" ) +theme(text = element_text(size = 20,  face="bold")) 
```


```{r}
sbae_targets_df_top<- sbae_targets_df_top[order(-sbae_targets_df_top$fraction),]
sbae_targets_df_top$sbae_targets<- factor(sbae_targets_df_top$sbae_targets, levels= sbae_targets_df_top$sbae_targets)
```

```{r}
sbae_targets_df_top$text <- paste0(sbae_targets_df_top$Freq, "/", sbae_targets_df_top$all_count)
```


```{r}
ggplot(sbae_targets_df_top[1:30,], aes(y=sbae_targets, x= fraction , label= text))+ geom_bar(stat="identity",color= "black", fill= "white") + 
    geom_text(size = 8, position = position_stack(vjust = 0.8)) + ylab("Drug Target") + xlab("The fraction of SBAE drugs over all drugs\nwith same target" ) +theme(text = element_text(size = 20,  face="bold")) 
#ggsave("~/results/FARES_plots/230313_drug_target_barplot.png", height=15, width = 10)
```


count table of drug enzymes
```{r}
sbae_enz <- unlist(str_split(drug_info[,3], ";"))
sbae_enz <- sbae_enz[!is.na(sbae_enz )]
sbae_enz <- sbae_enz[!sbae_enz == "NA"]
sbae_enz <- sbae_enz[!sbae_enz == ""]
```

```{r}
sbae_enz_df<- as.data.frame(table(sbae_enz))
```

count table of drug target in all targets
```{r}
all_enz <- unlist(str_split(drug_info_all[,3], ";"))
all_enz <- all_enz[!is.na(all_enz )]
all_enz <- all_enz[!all_enz == "NA"]
```

```{r}
all_enz_df<- as.data.frame(table(all_enz))
```

```{r}
length(all_enz_df$all_enz)
```
2082- total targets

```{r}
not_sbae_enz_list <- all_enz_df$all_enz[!all_enz_df$all_enz %in% sbae_enz_df$sbae_enz]
```

add 0 counts to target targets that are not used by SBAE drugs
```{r}
count <- rep(0, length(not_sbae_enz_list))
other_enz <- cbind(as.character(not_sbae_targets_list), count)
colnames(other_enz) <- colnames(sbae_enz_df)
```

```{r}
sbae_enz_test_df<- rbind(sbae_enz_df, other_enz)
```


matrix for each adverse event and drug target
```{r}
pvalue<- c()
plots2<- list()
test2<- list()
for ( j in 1:nrow(sbae_enz_df)){ #
   res <- drug_gene_test(as.character(sbae_enz_df[j,1]),as.numeric(sbae_enz_df[j,2]), 3, drug_info_all )
  pvalue[j]<- res[1]
  plots2[j]<- res[2]
  test2[j]<- res[3]
  }
```

```{r}
sbae_enz_df$p_adj<-  p.adjust(pvalue, method= "BH")
```

```{r}
sbae_enz_df_top <- sbae_enz_df[sbae_enz_df$p_adj < 0.001, ]
```

```{r}
names(test2) <- sbae_enz_df[, 1]
```

```{r}
test2_df <- as.data.frame(test2)
```

```{r}
test2_long_df <- pivot_longer(test2_df, cols=everything(), names_to = "Drug_Enzyme", values_to = "Count")
```

```{r}
test2_long_df$Drug_Enzyme<- ifelse(test2_long_df$Drug_Enzyme == "CYP3A7.CYP3A51P", "CYP3A7-CYP3A51P", test2_long_df$Drug_Enzyme)
```

```{r}
sbae_counts<- c()
for(i in 1:nrow(test2_long_df)){
  sbae_counts[i] <- sbae_enz_df$Freq[sbae_enz_df$sbae_enz == test2_long_df$Drug_Enzyme[i]]
}
```

```{r}
test2_long_df$sbae_number<- sbae_counts
```

```{r}
sbae_enz_df_top<- sbae_enz_df_top[order(-sbae_enz_df_top$Freq),]
test2_long_df$Drug_Enzyme<- factor(test2_long_df$Drug_Enzyme, levels= sbae_enz_df_top$sbae_enz)
```


```{r}
ggplot(test2_long_df, aes(x=Count, y= Drug_Enzyme ))+ geom_boxplot() +geom_point( aes(y=Drug_Enzyme,x= sbae_number ), color= "forestgreen",size =5) + xlab("Number of drugs with drug enzyme\nafter random selection" ) +theme(text = element_text(size = 20,  face="bold")) + ylab("Drug Enzyme")
ggsave("~/results/FARES_plots/230313_drug_enzyme_box_plot.png", height=20, width = 10)
```


```{r}
all_enz <- unlist(str_split(drug_info_all[,3], ";"))
all_enz <- all_enz[!is.na(all_enz )]
all_enz <- all_enz[!all_enz == "NA"]
```

```{r}
all_enz_df<- as.data.frame(table(all_enz))
```

```{r}
all_counts<- c()
for (i in 1:nrow(sbae_enz_df_top)){
  all_counts[i] <- all_enz_df$Freq[ unfactor(all_enz_df$all_enz) == sbae_enz_df_top$sbae_enz[i]]
}
```

```{r}
sbae_enz_df_top$all_count<- all_counts
```

```{r}
sbae_enz_df_top$fraction <- sbae_enz_df_top$Freq/sbae_enz_df_top$all_count
```

```{r}
sbae_enz_df_top<- sbae_enz_df_top[order(-sbae_enz_df_top$fraction),]
sbae_enz_df_top$sbae_enz<- factor(sbae_enz_df_top$sbae_enz, levels= sbae_enz_df_top$sbae_enz)
```

```{r}
sbae_enz_df_top$text <- paste0(sbae_enz_df_top$Freq, "/", sbae_enz_df_top$all_count)
```

```{r}
ggplot(sbae_enz_df_top, aes(y=sbae_enz, x= fraction , label= text))+ geom_bar(stat="identity",color= "black", fill= "white") + 
    geom_text(size = 8, position = position_stack(vjust = 0.7)) + ylab("Drug Metabolism Enzyme") + xlab("The fraction of SBAE drugs over all drugs\nwith same drug metabolism enzyme" ) +theme(text = element_text(size = 20,  face="bold")) 
ggsave("~/results/FARES_plots/230313_drug_metabolism_enzyme_barplot.png", height=15, width = 10)
```

save everything here... 
```{r}
saveRDS(sbae_targets_df_top, "~/results/FARES_plots/sbae_targets_df_top.rds")
saveRDS(sbae_enz_df_top, "~/results/FARES_plots/sbae_enz_df_top.rds")
saveRDS(drug_info_all,"~/results/FARES_plots/drug_info_all.rds" )
saveRDS(drug_info, "~/results/FARES_plots/drug_info_sbae_drugs.rds")
```

```{r}
saveRDS(plots,"~/results/FARES_plots/sbae_targets_plots.rds" )
saveRDS(test,"~/results/FARES_plots/sbae_targets_test.rds" )
saveRDS(plots2,"~/results/FARES_plots/sbae_enzyme_plots.rds" )
saveRDS(test2,"~/results/FARES_plots/sbae_enzyme_test.rds" )
```


# END
    Location of final scripts:
    src
    Location of data produced:
    ~/results/FARES_plots
    Dates when operations were done:
    230418
    
## Versions
```{r}
sessionInfo()
```



