---
title: "230323_sbae_heatmaps"
author: "Jennifer Fisher"
date: "2023-03-23"
output:
  html_document:
    toc: true
    theme: united
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# START 
    Goal/Purpose of operations: 
    I am going to be looking at the enriched drug targets from the SBAE drugs and looking at gene expression and core genes of sex-specific communities. 
    
    Finished psedocode on:
    230308
    
    System which operations were done on: 
    my laptop
    
    GitHub Repo:
    Sex_Bias_Adverse_Events
    
    Directory of operations:
    ~/Documents/Sex_Bias_Adverse_Events
    
    Scripts being edited for operations:
    NA
    
    Data being used:
    output/alpaca/alpaca_runs
    
    Papers and tools: 
    NA

# STEPS

### Set working directory 
```{r}
library(ggplot2)
library(circlize)
library(ComplexHeatmap)
library(drugbankR)
library(DBI)
library(hash)
library(org.Hs.eg.db)
library(EnsDb.Hsapiens.v75)
library(stringr)
```

```{r}
sbae_targets_df_top <- readRDS("~/results/FARES_plots/sbae_targets_df_top.rds")
```

```{r}
sex_bias_genes_gtex_science_paper <- read.table("~/data/GTEx_Analysis_v8_sbgenes/signif.sbgenes.txt", header=TRUE)
```


## Analysis


```{r}
#for each tissue find the male and female biased genes
tissues<- unique(sex_bias_genes_gtex_science_paper$tissue)
male_tissues_names <- paste("M", tissues, sep="_")
female_tissues_names<- paste("F", tissues, sep="_")
for (i in 1:length(tissues)){
  assign(male_tissues_names[i], sex_bias_genes_gtex_science_paper[sex_bias_genes_gtex_science_paper$tissue ==tissues[i] & sex_bias_genes_gtex_science_paper$effsize < 0,])
  assign(female_tissues_names[i], sex_bias_genes_gtex_science_paper[sex_bias_genes_gtex_science_paper$tissue ==tissues[i] & sex_bias_genes_gtex_science_paper$effsize > 0,])
}
```

```{r}
both_sex_tissue_names<- c(male_tissues_names, female_tissues_names)

geneset<- data.frame(id =both_sex_tissue_names, name= both_sex_tissue_names, genes= I(rep(list(c("a","b")), 88)))
sex_genes <- both_sex_tissue_names
for (i in 1:length(both_sex_tissue_names)){
  data <- get(both_sex_tissue_names[i])
  trim_names<- strtrim(data[,1],15)
  assign( sex_genes[i], trim_names)
  
}

sex_gene_set <- list(M_Adipose_Subcutaneous= M_Adipose_Subcutaneous, M_Adipose_Visceral_Omentum= M_Adipose_Visceral_Omentum)
```

```{r}
for (j in 1:88){
  sex_gene_set[[j]]<- get(sex_genes[j])
  #names(sex_gene_set[j])<- sex_genes[j]
}
names(sex_gene_set)<-sex_genes
```

```{r}
SBAE_drug_targets_v2 <- gprofiler2::gconvert(sbae_targets_df_top$sbae_targets)
```

```{r}
SBAE_drug_targets_v2 <- SBAE_drug_targets_v2[SBAE_drug_targets_v2$input == SBAE_drug_targets_v2$name,]
```

```{r}
male_gene_set <- sex_gene_set[1:44]
female_gene_set <- sex_gene_set[45:88]
```


```{r}
sbae_target_exp <- matrix(nrow=84, ncol=44)
for ( i in 1:84){
  m_exp <- as.numeric(grepl(SBAE_drug_targets_v2$target[i], male_gene_set))
  f_exp <- as.numeric(grepl(SBAE_drug_targets_v2$target[i], female_gene_set))
  sbae_target_exp[i,] <- f_exp- m_exp
}
```

```{r}
sbae_target_exp<- as.data.frame(sbae_target_exp)
```

```{r}
rownames(sbae_target_exp) <- SBAE_drug_targets_v2$name
colnames(sbae_target_exp)<- tissues
```


```{r}
sbae_target_exp_top <- sbae_target_exp[rowSums(sbae_target_exp==0, na.rm=TRUE)<ncol(sbae_target_exp), ]
```


```{r}

#colors2 <- structure(1:4, names = c("-1", "0", "1", "2")) # black, red, green, blue
colors<- 1:4
names(colors)<- c("-1", "0", "1")
colors[1]<- "#21908CFF"
colors[2]<- "white"
colors[3]<- "#440154FF"

png("~/results/FARES_plots/drug_target_sex_bias_expression_heatmap.png",width=15,height=10,units="in",res=1400)
Heatmap(t(sbae_target_exp_top), name = "Sex-bias Gene Expression", col = colors, heatmap_legend_param = list(at = c("-1", "0", "1"),labels= c("male", "not sex-bias", "female"), column_names_gp =gpar(fontsize = 18 , fontface = "bold"), labels_gp = gpar(fontsize = 12, fontface = "bold") ))
dev.off()
```


look at core genes 

```{r}
files <- list.files(path = "~/results/alpaca/alpaca_core_gene_lists")
```


```{r}
tissues<- str_split(files, "_female", simplify = TRUE)[,1]
tissues<- str_split(tissues, "_male", simplify = TRUE)[,1]
tissues<- unique(tissues)
```


```{r}
SBAE_drug_targets<- SBAE_drug_targets_v2$name
```

male core genes
```{r}
core_gene_list <- list()
for( i in 1:length(tissues)){
  print(tissues[i])
  male_core_genes <- readRDS(paste("~/results/alpaca/alpaca_core_gene_lists/", tissues[i],"_male_core_genes.rds", sep=""))
  core_gene_list[[i]] <- unique(male_core_genes$symbol)
}
```

```{r}
female_core_gene_list <- list()
for( i in 1:length(tissues)){
  print(tissues[i])
  female_core_genes <- readRDS(paste("~/results/alpaca/alpaca_core_gene_lists/", tissues[i],"_female_core_genes.rds", sep=""))
  female_core_gene_list[[i]] <- unique(female_core_genes$symbol)
}
```

```{r}
names(female_core_gene_list)<- tissues
names(core_gene_list)<- tissues
```



```{r}
sbae_target_core <- matrix(nrow=84, ncol=43)
for ( i in 1:84){
  
  m_core <- which(lapply(core_gene_list, function(x) grep(paste("^",SBAE_drug_targets[i],"$", sep=""), x)) !=0)
  m_core<- 1:43 %in% m_core
  
  f_core <- which(lapply(female_core_gene_list, function(x) grep(paste("^",SBAE_drug_targets[i],"$", sep=""), x))!=0)
  f_core<- 1:43 %in% f_core
  
  sub <- rep(0, 43)
  
  sub <- ifelse(m_core, -1, sub)
  sub <- ifelse(f_core, 1, sub)
  sbae_target_core[i,]<- ifelse(f_core & m_core, 1e-15, sub)

}
```

```{r}
rownames(sbae_target_core) <- SBAE_drug_targets_v2$name
colnames(sbae_target_core)<- tissues
```


```{r}
sbae_target_core_top <- sbae_target_core[rowSums(sbae_target_core == 0, na.rm=TRUE)<ncol(sbae_target_core), ]
```


```{r}

#colors2 <- structure(1:4, names = c("-1", "0", "1", "2")) # black, red, green, blue
colors<- 1:4
names(colors)<- c("-1", "0","1e-15", "1")
colors[1] <- "#21908CFF"
colors[2] <- "white"
colors[3] <- "gray"
colors[4] <- "#440154FF"

png("~/results/FARES_plots/drug_target_sex_bias_core_genes_heatmap.png",width=15,height=10,units="in",res=1400)
Heatmap(t(sbae_target_core_top), name = "Sex-bias Core Gene", col = colors, heatmap_legend_param = list(at = c("-1", "0","1e-15", "1"),labels= c("male", "not sex-bias","both", "female"), column_names_gp =gpar(fontsize = 18 , fontface = "bold"), labels_gp = gpar(fontsize = 12, fontface = "bold") ))
dev.off()
```



```{r}
setdiff(rownames(sbae_target_core_top), rownames(sbae_target_exp_top))
```

```{r}
setdiff(rownames(sbae_target_exp_top), rownames(sbae_target_core_top))
```


```{r}
intersect(rownames(sbae_target_exp_top), rownames(sbae_target_core_top))
```

```{r}
unique(c(rownames(sbae_target_exp_top), rownames(sbae_target_core_top)))
```

```{r}
68/84
```


```{r}
sex_bias_genes<- unique(c(rownames(sbae_target_exp_top), rownames(sbae_target_core_top)))
```

```{r}
drug_ae_res_info <- readRDS("~/results/FARES_plots/drug_ae_res_info.rds")
```

```{r}
drug_info_sbae_drugs <- readRDS("~/results/FARES_plots/drug_info_sbae_drugs.rds")
```

```{r}
dim(drug_info_sbae_drugs)
```

```{r}
length(grep(paste0(sex_bias_genescollapse="|")  , drug_info_sbae_drugs$targets ))
```

```{r}
389/416
```


look at top targets

```{r}
drug_target_drugs <- drug_info_sbae_drugs$drug_ids[grep("ADRA1A", drug_info_sbae_drugs$targets)]
```

```{r}
length(drug_target_drugs)
```
26

```{r}
drug_ae_res_info <- readRDS("~/results/FARES_plots/drug_ae_res_info.rds")
```

```{r}
res_drug_target_subset <- drug_ae_res_info[drug_ae_res_info$drug %in% drug_target_drugs,]
```


```{r}
res_drug_target_subset<- res_drug_target_subset[res_drug_target_subset$BH < 0.05 & res_drug_target_subset$logROR < -1,]
```


```{r}
adverse_event_counts <- as.data.frame(table(res_drug_target_subset$soc_term))
```


```{r}
ggplot(adverse_event_counts, aes(x= Freq, y= Var1, label= Freq)) + geom_bar(stat="identity", fill =  "#21908CFF", color= "black", alpha=0.7)+ 
    geom_text(size = 10, position = position_stack(vjust = 0.9)) + xlab("Number of male-bias drug-adverse event pairs with drugs with drug target") + ylab("Drug")+theme(text = element_text(size = 12,  face="bold"))
```

```{r}
drug_target_drugs <- drug_info_sbae_drugs$drug_ids[grep("ADRA2A", drug_info_sbae_drugs$targets)]
```

```{r}
length(drug_target_drugs)
```
24


```{r}
res_drug_target_subset <- drug_ae_res_info[drug_ae_res_info$drug %in% drug_target_drugs,]
```


```{r}
res_drug_target_subset<- res_drug_target_subset[res_drug_target_subset$BH < 0.05 & res_drug_target_subset$logROR < -1,]
```


```{r}
adverse_event_counts <- as.data.frame(table(res_drug_target_subset$soc_term))
```


```{r}
ggplot(adverse_event_counts, aes(x= Freq, y= Var1, label= Freq)) + geom_bar(stat="identity", fill =  "#21908CFF", color= "black", alpha=0.7)+ 
    geom_text(size = 10, position = position_stack(vjust = 0.9)) + xlab("Number of male-bias drug-adverse event pairs with drugs with drug target") + ylab("Drug")+theme(text = element_text(size = 12,  face="bold"))
```


```{r}
drug_target_drugs <- drug_info_sbae_drugs$drug_ids[grep("CHRNA4", drug_info_sbae_drugs$targets)]
```

```{r}
length(drug_target_drugs)
```
8


```{r}
res_drug_target_subset <- drug_ae_res_info[drug_ae_res_info$drug %in% drug_target_drugs,]
```


```{r}
res_drug_target_subset<- res_drug_target_subset[res_drug_target_subset$BH < 0.05 & res_drug_target_subset$logROR < -1,]
```


```{r}
adverse_event_counts <- as.data.frame(table(res_drug_target_subset$pt_term))
```


```{r}
ggplot(adverse_event_counts, aes(x= Freq, y= Var1, label= Freq)) + geom_bar(stat="identity", fill =  "#21908CFF", color= "black", alpha=0.7)+ 
    geom_text(size = 10, position = position_stack(vjust = 0.9)) + xlab("Number of male-bias drug-adverse event pairs with drugs with drug target") + ylab("Drug")+theme(text = element_text(size = 12,  face="bold"))
```


```{r}
female_ae_test <- drug_ae_res_info[drug_ae_res_info$pt_term == "Alopecia" & abs(drug_ae_res_info$logROR) > 1 & drug_ae_res_info$BH < 0.05,]
```
only 153 have a female-bias for Alopecia

```{r}
hist(female_ae_test$c /(female_ae_test$a+female_ae_test$b+female_ae_test$c+female_ae_test$d))
```

```{r}
hist((female_ae_test$a-female_ae_test$c)/(female_ae_test$a+female_ae_test$b+female_ae_test$c+female_ae_test$d))
```



```{r}
female_ae_test <- drug_ae_res_info[drug_ae_res_info$pt_term == "Urinary tract infection" & abs(drug_ae_res_info$logROR) > 1 & drug_ae_res_info$BH < 0.05,]
```
only 88 have a female-bias for Alopecia

```{r}
hist(female_ae_test$c /(female_ae_test$a+female_ae_test$b+female_ae_test$c+female_ae_test$d))
```

```{r}
hist((female_ae_test$a-female_ae_test$c)/(female_ae_test$a+female_ae_test$b+female_ae_test$c+female_ae_test$d))
```

```{r}
drug_ae_res_info[drug_ae_res_info$pt_term == "Alopecia" & drug_ae_res_info$logROR > 1 & drug_ae_res_info$BH < 0.05,]
```



```{r}
Liver_F_panda <- readRDS("~/results/alpaca/sex_specific_networks/Liver_F_panda.rds")
```


```{r}
drug_gene_core_test<- function( count = 5, all_drug_targets_list= colnames(Liver_F_panda@regNet), core_list= c(core_gene_list$Liver, female_core_gene_list$Liver), seed= 101){
  res<- c()
  set.seed(seed)
  for (i in 1:1000){
    #t<- table(grepl("",))
    sub<- sample( all_drug_targets_list, 64, replace = FALSE)

    res1 <- table(sub %in% core_list)
    res2<- res1[names(res1) == "TRUE"]
    if (length(res2) ==1){
      res[i]<-res2
    }else{
      res[i]<- 0
    }
    
  }
  #print(res)
  test<- wilcox.test(res, mu= count, alternative = "less")
  #print(count)
  #print(hist(res))
  plot_data<- as.data.frame(cbind(1:1000, res))
  plot1 <- ggplot(plot_data, aes(x=res)) + geom_histogram() + geom_vline(xintercept = count) + xlab("Number of genes that are core genes")
  
  res_list<- list(test$p.value, plot1)
  return(res_list)
}
```

```{r}
drug_gene_core_test()
```

```{r}
drug_gene_core_test( count = 5, all_drug_targets_list= colnames(Liver_F_panda@regNet), core_list= core_gene_list$Liver, seed= 101)
```

```{r}
drug_gene_core_test( count = 0, all_drug_targets_list= colnames(Liver_F_panda@regNet), core_list= female_core_gene_list$Liver, seed= 101)
```


