---
title: "230417_liver_drug_edges"
output: html_document
date: "2023-04-17"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# START 
    Goal/Purpose of operations: 
    I am going to be looking at the drug metabloism gene edges for the sex-specific networks across the 43 different tissues. 
    
    Finished psedocode on:
    230206
    
    System which operations were done on: 
    my laptop
    
    GitHub Repo:
    230321_JLF_Sex_bias_adverse_events
    
    Docker: 
    jenfisher7/rstudio_sex_bias_drugs
    
    Directory of operations:
    ~/Documents/230321_JLF_Sex_bias_adverse_events
    
    Scripts being edited for operations:
    NA
    
    Data being used:
    results/alpaca/alpaca_runs
    
    Papers and tools: 
    NA

# STEPS

### Set working directory 

```{r}
source("~/src/sex_bias_drug_functions.R")
library(gprofiler2)
library(viridis)
library(RColorBrewer)
library(ComplexHeatmap)
library(netZooR)
library(ggplot2)
library(GOSemSim)
library(parallel)
```

Start by looking at Liver by itself
```{r}
Liver_M_panda <- readRDS("~/results/alpaca/sex_specific_networks/Liver_M_panda.rds")
Liver_F_panda <- readRDS("~/results/alpaca/sex_specific_networks/Liver_F_panda.rds")
Liver__male_female_long <- readRDS("~/results/alpaca/alpaca_adjusted_inputs/Liver__topNet_regnet_male_female_long.rds")
```


Now look at edges of drug metabolism genes in the liver
 
```{r}
Liver_tf_gene_sex_drug_only <- Liver__male_female_long[Liver__male_female_long$gene %in% drug_genes,]
```

```{r}
Liver_tf_gene_sex_drug_only<- Liver_tf_gene_sex_drug_only[ Liver_tf_gene_sex_drug_only$male ==1 | Liver_tf_gene_sex_drug_only$female == 1,]
```

```{r}
all_male_tfs <- unique(Liver_tf_gene_sex_drug_only$TF[Liver_tf_gene_sex_drug_only$male == 1])
all_female_tfs <- unique(Liver_tf_gene_sex_drug_only$TF[Liver_tf_gene_sex_drug_only$female == 1])
```

```{r}
setdiff(all_male_tfs, all_female_tfs)
```

```{r}
setdiff(all_female_tfs, all_male_tfs)
```
only one transcription factor is unique to male the rest are shared. 

```{r}
Liver_sex_bias_tf_list <- readRDS("~/results/alpaca/panda_out_degree/Liver_sex_bias_tf_list.rds")
```

```{r}
all_tfs <- unique(all_female_tfs, all_male_tfs)
sex_div_tf <- Liver_sex_bias_tf_list[[1]][Liver_sex_bias_tf_list[[1]] %in% all_tfs]
length(sex_div_tf)
f_bias_tf <- Liver_sex_bias_tf_list[[2]][Liver_sex_bias_tf_list[[2]] %in% all_tfs]
length(f_bias_tf)
m_bias_tf <- Liver_sex_bias_tf_list[[3]][Liver_sex_bias_tf_list[[3]] %in% all_tfs]
length(m_bias_tf)
```

```{r}
Liver_tf_gene_sex_drug_only[Liver_tf_gene_sex_drug_only$TF == "ETS1",]
```
This gene encodes a member of the aldehyde dehydrogenase protein family. Aldehyde dehydrogenases are a family of isozymes that may play a major role in the detoxification of aldehydes generated by alcohol metabolism and lipid peroxidation. The encoded protein is able to oxidize long-chain fatty aldehydes in vitro, and may play a role in protection from oxidative stress.

To plot the edge relationships on a heatmap and have it cluster 
"1e-10", "-1", "0", "1" == "no edge", "male", "both", "female"
```{r}
Liver_tf_gene_sex_drug_only$edge_diff <- (Liver_tf_gene_sex_drug_only$female - Liver_tf_gene_sex_drug_only$male)
```

```{r}
liver_drug_edges <- pivot_wider(data= as.data.frame( Liver_tf_gene_sex_drug_only[,-c(3,4,6)] ), names_from = gene, values_from=edge_diff, values_fill= 1e-10)
```

```{r}
liver_drug_edges<- as.data.frame(liver_drug_edges)
rownames(liver_drug_edges)<- liver_drug_edges$TF
```

```{r}
liver_drug_edges<- liver_drug_edges[,-1]
```

```{r}
colors<- 1:4
names(colors)<- c("1e-10", "-1", "0", "1")
colors[1]<- "white"
colors[2]<- "#21908CFF" #"#440154FF", "#21908CFF"
colors[3]<- "lightgray"
colors[4]<- "#440154FF"

png("~/results/alpaca/drug_metabolism_genes/liver_drug_metabolism_edge_heatmap.png",width=15,height=10,units="in",res=1400)
Heatmap(liver_drug_edges, name = "Edge Present", col = colors, show_row_names = FALSE,
    column_title = "Male and female edges in liver networks",heatmap_legend_param = list(at = c("1e-10", "-1", "0", "1"),labels= c("no edge", "male", "both", "female"), column_names_gp =gpar(fontsize = 18 , fontface = "bold"), labels_gp = gpar(fontsize = 12, fontface = "bold") ))
dev.off()
```

annotating some of the tfs based on the number of sex-bias edges
```{r}
tf_count_table<- matrix(0, nrow= nrow(liver_drug_edges), ncol=4 )
for (i in 1:nrow(liver_drug_edges)){
  tf_count_sub <- table(unlist(liver_drug_edges[i,]))
  try(tf_count_table[i,1]<- tf_count_sub[names(tf_count_sub) == -1])
  try(tf_count_table[i,2]<- tf_count_sub[names(tf_count_sub) == 0])
  try(tf_count_table[i,3]<- tf_count_sub[names(tf_count_sub) == 1e-10 ])
  try(tf_count_table[i,4]<- tf_count_sub[names(tf_count_sub) == 1 ])
}
rownames(tf_count_table)<- rownames(liver_drug_edges)
```

```{r}
hist(tf_count_table[,4]- tf_count_table[,1])
```

```{r}
print("male_tf_subset_drug_genes")
rownames(tf_count_table)[(tf_count_table[,4]- tf_count_table[,1]) < -2]
male_tf <- rownames(tf_count_table)[(tf_count_table[,4]- tf_count_table[,1]) < -2]
print("female_tf_subset_drug_genes")
rownames(tf_count_table)[(tf_count_table[,4]- tf_count_table[,1]) > 4]
female_tf <- rownames(tf_count_table)[(tf_count_table[,4]- tf_count_table[,1]) > 4]
```

```{r}
tf_sub  <- c(female_tf, male_tf )
tf_location <- (1:nrow(liver_drug_edges))[rownames(liver_drug_edges) %in% tf_sub]
Tf_anno<- anno_mark(at = tf_location, labels= rownames(liver_drug_edges)[tf_location], which = "row")
```


```{r}
png("~/results/alpaca/drug_metabolism_genes/liver_drug_metabolism_edge_heatmap_v2.png",width=15,height=10,units="in",res=1400)
Heatmap(liver_drug_edges, name = "Edge Present", col = colors, show_row_names = FALSE,
    column_title = "Male and female edges in liver networks",heatmap_legend_param = list(at = c("1e-10", "-1", "0", "1"),labels= c("no edge", "male", "both", "female"), column_names_gp =gpar(fontsize = 18 , fontface = "bold"), labels_gp = gpar(fontsize = 12, fontface = "bold") )) + rowAnnotation(mark = Tf_anno)
dev.off()
```


For the liver sex-specific gene-regulatory networks, we determined the predicted regulation of the transcription factor and targeted gene by finding the Pearson correlation between a given transcription factor gene’s expression and the targeted drug gene metabolism gene’s expression. If the Pearson correlation was significant after Benjamini-Hochberg (BH) p-value adjustment and positive, we defined these edges as activator edges. If the Pearson correlation was significant after BH p-value adjustment and negative, we described these edges as repressor edges. If the correlation was not significant, we defined these relationships as undefined. We applied this in a sex-specific manner where we only correlated the transcription factor and target gene’s expression from one sex, and in a non-sex-specific manner as well. 
```{r}
Liver_gtex_qsmooth_expression <- readRDS("~/results/lioness/processed_normalized_expression/Liver_gtex_processed_expression.rds")
Liver_gtex_metadata <- readRDS("~/data/gtex_qsmooth/Liver_gtex_metadata.rds")
```

```{r}
identical(colnames(Liver_gtex_qsmooth_expression), Liver_gtex_metadata$external_id)
```


```{r}
Liver_gtex_qsmooth_expression<- as.data.frame(Liver_gtex_qsmooth_expression)
Liver_gtex_qsmooth_expression$genes <- rownames(Liver_gtex_qsmooth_expression)
Liver_gtex_qsmooth_expression_long <- pivot_longer(as.data.frame(Liver_gtex_qsmooth_expression), col=!genes, names_to = "sample", values_to = "expression")
```


```{r}
male_samples<- Liver_gtex_metadata$external_id[Liver_gtex_metadata$gtex.sex ==  "M"]
Liver_gtex_qsmooth_expression_long$sex <- Liver_gtex_qsmooth_expression_long$sample %in% male_samples
Liver_gtex_qsmooth_expression_long$sex <- ifelse(Liver_gtex_qsmooth_expression_long$sex, "male", "female")
```

```{r}
table(Liver_gtex_qsmooth_expression_long$sex)
```

Calculating correaltion for all the TF and drug genes
```{r}
all_cor<- c()
all_p<- c()
f_cor<- c()
f_p<- c()
m_cor <- c()
m_p<- c()
for (i in 1:nrow(Liver_tf_gene_sex_drug_only)){ 
  tf_name<- Liver_tf_gene_sex_drug_only[i,1]
  gene_name<- Liver_tf_gene_sex_drug_only[i,2]
  #get tf expression
  tf <- Liver_gtex_qsmooth_expression_long$expression[Liver_gtex_qsmooth_expression_long$genes == tf_name]
  gene<-  Liver_gtex_qsmooth_expression_long$expression[Liver_gtex_qsmooth_expression_long$genes == gene_name]
  res<- cor.test(tf, gene, method= "pearson",exact=FALSE)
  all_cor[i]<- res$estimate
  all_p[i]<- res$p.value
  
  tf <- Liver_gtex_qsmooth_expression_long$expression[Liver_gtex_qsmooth_expression_long$genes == tf_name & Liver_gtex_qsmooth_expression_long$sex == "female"]
  gene<-  Liver_gtex_qsmooth_expression_long$expression[Liver_gtex_qsmooth_expression_long$genes == gene_name & Liver_gtex_qsmooth_expression_long$sex == "female"]
  res<- cor.test(tf, gene, method= "pearson",exact=FALSE)
  f_cor[i]<- res$estimate
  f_p[i]<- res$p.value
  
   tf <- Liver_gtex_qsmooth_expression_long$expression[Liver_gtex_qsmooth_expression_long$genes == tf_name & Liver_gtex_qsmooth_expression_long$sex == "male"]
  gene<-  Liver_gtex_qsmooth_expression_long$expression[Liver_gtex_qsmooth_expression_long$genes == gene_name & Liver_gtex_qsmooth_expression_long$sex == "male"]
  res<- cor.test(tf, gene, method= "pearson",exact=FALSE)
  m_cor[i]<- res$estimate
  m_p[i]<- res$p.value
}
Liver_tf_gene_sex_drug_only<- cbind(Liver_tf_gene_sex_drug_only, all_cor,all_p, m_cor, m_p, f_cor, f_p)
```

note the type of relationships between TF and gene
```{r}
Liver_tf_gene_sex_drug_only$all_rel <- ifelse(Liver_tf_gene_sex_drug_only$all_cor >0, "pos", "neg")
Liver_tf_gene_sex_drug_only$f_rel <- ifelse(Liver_tf_gene_sex_drug_only$f_cor >0, "pos", "neg")
Liver_tf_gene_sex_drug_only$m_rel <- ifelse(Liver_tf_gene_sex_drug_only$m_cor >0, "pos", "neg")
```

pvalue correction (BH)

```{r}
Liver_tf_gene_sex_drug_only$all_p_adj <- p.adjust(Liver_tf_gene_sex_drug_only$all_p, method= "BH")
Liver_tf_gene_sex_drug_only$f_p_adj <- p.adjust(Liver_tf_gene_sex_drug_only$f_p, method= "BH")
Liver_tf_gene_sex_drug_only$m_p_adj <- p.adjust(Liver_tf_gene_sex_drug_only$m_p, method= "BH")
```

```{r}
Liver_tf_gene_sex_drug_only$all_p_adj_tf <- ifelse( Liver_tf_gene_sex_drug_only$all_p_adj < 0.05, "yes", "no")
Liver_tf_gene_sex_drug_only$f_p_adj_tf <- ifelse( Liver_tf_gene_sex_drug_only$f_p_adj < 0.05, "yes", "no")
Liver_tf_gene_sex_drug_only$m_p_adj_tf <- ifelse( Liver_tf_gene_sex_drug_only$m_p_adj < 0.05, "yes", "no")

```

```{r}
Liver_tf_gene_sex_drug_only$summary <- paste(Liver_tf_gene_sex_drug_only$f_rel, Liver_tf_gene_sex_drug_only$f_p_adj_tf, Liver_tf_gene_sex_drug_only$m_rel, Liver_tf_gene_sex_drug_only$m_p_adj_tf, sep = "_")
```

```{r}
saveRDS(Liver_tf_gene_sex_drug_only, "~/results/alpaca/drug_metabolism_genes/liver_drug_metabolism_gene_edges.rds")
```


```{r}
Liver_tf_gene_sex_drug_both <- Liver_tf_gene_sex_drug_only[Liver_tf_gene_sex_drug_only$edge_diff == 0,]

df_edge_cor_counts<- as.data.frame( table(Liver_tf_gene_sex_drug_both$summary ))
```

```{r}
df_edge_cor_counts
```



```{r}
Liver_tf_gene_sex_drug_only$edge <- paste(Liver_tf_gene_sex_drug_only$TF, Liver_tf_gene_sex_drug_only$gene, sep= "_")
```



tfs activating in females
```{r}
f_act_edge<- unique(Liver_tf_gene_sex_drug_only$edge[Liver_tf_gene_sex_drug_only$f_cor > 0 & Liver_tf_gene_sex_drug_only$f_p_adj < 0.05 & Liver_tf_gene_sex_drug_only$female== 1 ])
```


tfs repressing in females
```{r}
f_repress_edge<-unique(Liver_tf_gene_sex_drug_only$edge[Liver_tf_gene_sex_drug_only$f_cor < 0 & Liver_tf_gene_sex_drug_only$f_p_adj < 0.05 & Liver_tf_gene_sex_drug_only$female== 1])
```

tfs activating in males
```{r}
m_act_edge<-unique(Liver_tf_gene_sex_drug_only$edge[Liver_tf_gene_sex_drug_only$m_cor > 0 & Liver_tf_gene_sex_drug_only$m_p_adj < 0.05  & Liver_tf_gene_sex_drug_only$male== 1])
```

```{r}
m_repress_edge<-unique(Liver_tf_gene_sex_drug_only$edge[Liver_tf_gene_sex_drug_only$m_cor < 0 & Liver_tf_gene_sex_drug_only$m_p_adj < 0.05  & Liver_tf_gene_sex_drug_only$male== 1])
```

m3 = make_comb_mat(lt, mode = "union")

```{r}
comb_mat<- make_comb_mat(list("Female Activator Edges"=f_act_edge, "Female Repressor Edges"=f_repress_edge, "Male Activator Edges"=m_act_edge, "Male Repressor Edges"=m_repress_edge), mode = "distinct")
```

```{r}
png("~/results/alpaca/drug_metabolism_genes/liver_drug_metabolism_edge_upset_plot.png",width=15,height=10,units="in",res=1400)
cs <-  comb_size(comb_mat)
plot<- UpSet(comb_mat, comb_order = order(-comb_size(comb_mat)), top_annotation = HeatmapAnnotation(
        "Intersection size" = anno_barplot(cs, 
            ylim = c(0, max(cs)*1.1),
            border = FALSE, 
            gp = gpar(fill = "black"), 
            height = unit(4, "cm")
        ), 
        annotation_name_side = "left", 
        annotation_name_rot = 90))
plot_draw <- draw(plot)
od <- column_order(plot_draw)
decorate_annotation("Intersection size", {
    grid.text(cs[od], x = seq_along(cs), y = unit(cs[od], "native") + unit(2, "pt"), 
        default.units = "native", just = c("left", "bottom"), 
        gp = gpar(fontsize = 11, face="bold",col = "#404040"), rot = 45)
})
dev.off()
```


pathway analysis on the different edges

```{r}
only_male_act_edges<- setdiff(m_act_edge, c(m_repress_edge, f_repress_edge, f_act_edge ))
```

```{r}
only_male_act_tf_genes <- unique(unlist(str_split(only_male_act_edges, "_")))
```


```{r}
set.seed(101)
  pathway_results_male_act <- gost(query = only_male_act_tf_genes , 
                               organism = "hsapiens", ordered_query = FALSE, 
                               multi_query = FALSE, significant = TRUE, exclude_iea= FALSE, measure_underrepresentation = FALSE, evcodes = FALSE,
                               user_threshold = 0.05, correction_method = "g_SCS", 
                               domain_scope = "annotated", custom_bg = NULL, 
                               numeric_ns = "", as_short_link = FALSE)
   pathway_results_male_res_act <-  pathway_results_male_act$result
```

```{r}
only_male_repress_edges<- setdiff(m_repress_edge, c(f_repress_edge, f_act_edge ,m_act_edge))
```

```{r}
only_male_repress_tf_genes <- unique(unlist(str_split(only_male_repress_edges, "_")))
```


```{r}
set.seed(101)
  pathway_results_male_repress <- gost(query = only_male_repress_tf_genes , 
                               organism = "hsapiens", ordered_query = FALSE, 
                               multi_query = FALSE, significant = TRUE, exclude_iea= FALSE, measure_underrepresentation = FALSE, evcodes = FALSE,
                               user_threshold = 0.05, correction_method = "g_SCS", 
                               domain_scope = "annotated", custom_bg = NULL, 
                               numeric_ns = "", as_short_link = FALSE)
   pathway_results_male_res_repress <-  pathway_results_male_repress$result
```


```{r}
only_female_act_edges<- setdiff(f_act_edge, c(f_repress_edge, m_repress_edge, m_act_edge ))
```

```{r}
only_female_act_tf_genes <- unique(unlist(str_split(only_female_act_edges, "_")))
```


```{r}
set.seed(101)
  pathway_results_female_act <- gost(query = only_female_act_tf_genes , 
                               organism = "hsapiens", ordered_query = FALSE, 
                               multi_query = FALSE, significant = TRUE, exclude_iea= FALSE, measure_underrepresentation = FALSE, evcodes = FALSE,
                               user_threshold = 0.05, correction_method = "g_SCS", 
                               domain_scope = "annotated", custom_bg = NULL, 
                               numeric_ns = "", as_short_link = FALSE)
   pathway_results_female_res_act <-  pathway_results_female_act$result
```

```{r}
only_female_repress_edges<- setdiff(f_repress_edge, c(m_repress_edge, m_act_edge ,f_act_edge))
```

```{r}
only_female_repress_tf_genes <- unique(unlist(str_split(only_female_repress_edges, "_")))
```


```{r}
set.seed(101)
  pathway_results_female_repress <- gost(query = only_female_repress_tf_genes , 
                               organism = "hsapiens", ordered_query = FALSE, 
                               multi_query = FALSE, significant = TRUE, exclude_iea= FALSE, measure_underrepresentation = FALSE, evcodes = FALSE,
                               user_threshold = 0.05, correction_method = "g_SCS", 
                               domain_scope = "annotated", custom_bg = NULL, 
                               numeric_ns = "", as_short_link = FALSE)
   pathway_results_female_res_repress <-  pathway_results_female_repress$result
```




```{r}
hsGO <- godata('org.Hs.eg.db', ont="BP")
```

```{r}
go_term_heatmap_v2(pathway_results_female_res_act, pathway_results_female_res_repress, pathway_results_male_res_act , pathway_results_male_res_repress, threshold = 0.94, "~/results/alpaca/drug_metabolism_genes/liver_edge_differences_pathways.csv")
```


```{r}
list("Female Activator Edges"=f_act_edge, "Female Repressor Edges"=f_repress_edge, "Male Activator Edges"=m_act_edge, "Male Repressor Edges"=m_repress_edge)
```

```{r}
f_act_edge_gene_count <- as.data.frame(table(str_split(f_act_edge, "_", simplify = TRUE)[,2]))
f_repress_edge_gene_count <- as.data.frame(table(str_split(f_repress_edge, "_", simplify = TRUE)[,2]))
m_act_edge_gene_count <- as.data.frame(table(str_split(m_act_edge, "_", simplify = TRUE)[,2]))
m_repress_edge_gene_count <- as.data.frame(table(str_split(m_repress_edge, "_", simplify = TRUE)[,2]))
```

```{r}
count_act_repress <- merge(f_act_edge_gene_count, f_repress_edge_gene_count, by= "Var1", all =TRUE)
count_act_repress <- merge(count_act_repress, m_act_edge_gene_count, by= "Var1", all =TRUE)
count_act_repress <- merge(count_act_repress, m_repress_edge_gene_count, by= "Var1", all =TRUE)
```

```{r}
rownames(count_act_repress) <- count_act_repress[,1]
```

```{r}
count_act_repress<- count_act_repress[,-1]
```

```{r}
colnames(count_act_repress)<- c("Female_Activator_Edges", "Female_Repressor_Edges", "Male_Activator_Edges", "Male_Repressor_Edges")
```

```{r}
count_act_repress$Female_Activator_Edges<- ifelse( is.na(count_act_repress$Female_Activator_Edges), 0, count_act_repress$Female_Activator_Edges)
count_act_repress$Female_Repressor_Edges<- ifelse( is.na(count_act_repress$Female_Repressor_Edges), 0, count_act_repress$Female_Repressor_Edges)
count_act_repress$Male_Activator_Edges<- ifelse( is.na(count_act_repress$Male_Activator_Edges), 0, count_act_repress$Male_Activator_Edges)
count_act_repress$Male_Repressor_Edges<- ifelse( is.na(count_act_repress$Male_Repressor_Edges), 0, count_act_repress$Male_Repressor_Edges)
```


```{r}
col_fun = colorRamp2(c(0,  45), c( "black", "yellow"))
Heatmap(t(count_act_repress), nam= "Number of Edges", col = col_fun, show_column_names = TRUE,  show_row_names = TRUE,
          clustering_distance_rows= "euclidean",
          clustering_distance_columns=  "euclidean",
          clustering_method_rows = "ward.D2" ,
          clustering_method_columns="ward.D2", heatmap_legend_param = list( title_gp = gpar(fontsize = 15 , fontface = "bold"), labels_gp = gpar(fontsize = 12, fontface = "bold")))
```

```{r}
female_diff<-  count_act_repress$Female_Activator_Edges- count_act_repress$Female_Repressor_Edges
male_diff<-  count_act_repress$Male_Activator_Edges- count_act_repress$Male_Repressor_Edges
```

```{r}
diff_matrix<- cbind(female_diff, male_diff)
```

```{r}
rownames(diff_matrix)<- rownames(count_act_repress)
colnames(diff_matrix)<- c("Female_Edge_Differences", "Male_Edge_Differences")
```

```{r}
diff_matrix[,1]<- as.numeric(diff_matrix[,1])
diff_matrix[,2]<- as.numeric(diff_matrix[,2])
```


```{r}
#col_fun <- colorRamp2(c(-30, 0,  30), c( "blue", "white", "yellow"))
Heatmap(t(diff_matrix), nam= "Activator - Repressor\nEdges",col=colorRamp2(c(-26, 0,  26), c( "blue", "black", "yellow")), show_column_names = TRUE,  show_row_names = TRUE,
          clustering_distance_rows= "euclidean",
          clustering_distance_columns=  "euclidean",
          clustering_method_rows = "ward.D2" ,
          clustering_method_columns="ward.D2", heatmap_legend_param = list( title_gp = gpar(fontsize = 15 , fontface = "bold", at = c(-26,0,26)), labels_gp = gpar(fontsize = 12, fontface = "bold")))
```



