---
title: "230426_more_drug_metabolism_testing"
author: "Jennifer Fisher"
date: "2023-03-13"
output:
  html_document:
    toc: true
    theme: united
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# START 
    Goal/Purpose of operations: 
    I am going to be looking at the enriched drug targets from the SBAE drugs and looking at gene expression and core genes of sex-specific communities. 
    
    Finished psedocode on:
    230308
    
    System which operations were done on: 
    my laptop
    
    GitHub Repo:
    230321_JLF_Sex_bias_adverse_events
    
    Docker: 
    jenfisher7/rstudio_sex_bias_drugs
    
    Directory of operations:
    ~/Documents/230321_JLF_Sex_bias_adverse_events
    
    Scripts being edited for operations:
    NA
    
    Data being used:
    output/alpaca/alpaca_runs
    
    Papers and tools: 
    "~/results/FARES_plots/drug_info_sbae_drugs.rds"
    "~/results/FARES_plots/drug_info_all.rds"
    "~/results/FARES_plots/sbae_targets_df_top.rds"

# STEPS

### Set working directory 
```{r}
source("~/src/sex_bias_drug_functions.R")
library(ggplot2)
library(circlize)
library(ComplexHeatmap)
library(drugbankR)
library(DBI)
library(hash)
library(org.Hs.eg.db)
library(EnsDb.Hsapiens.v75)
library(ggpubr)
library(stringr)
```


```{r}
drug_info_sbae_drugs <- readRDS("~/results/FARES_plots/drug_info_sbae_drugs.rds")
drug_info_all <- readRDS("~/results/FARES_plots/drug_info_all.rds")
sbae_targets_df_top <- readRDS("~/results/FARES_plots/sbae_targets_df_top.rds")
```

```{r}
sex_bias_genes_gtex_science_paper <- read.table("~/data/GTEx_Analysis_v8_sbgenes/signif.sbgenes.txt", header=TRUE)
```

```{r}
Liver_F_panda <- readRDS("~/results/alpaca/sex_specific_networks/Liver_F_panda.rds")
```

## Analysis

```{r}
#for each tissue find the male and female biased genes
tissues<- unique(sex_bias_genes_gtex_science_paper$tissue)
male_tissues_names <- paste("M", tissues, sep="_")
female_tissues_names<- paste("F", tissues, sep="_")
for (i in 1:length(tissues)){
  assign(male_tissues_names[i], sex_bias_genes_gtex_science_paper[sex_bias_genes_gtex_science_paper$tissue ==tissues[i] & sex_bias_genes_gtex_science_paper$effsize < 0,])
  assign(female_tissues_names[i], sex_bias_genes_gtex_science_paper[sex_bias_genes_gtex_science_paper$tissue ==tissues[i] & sex_bias_genes_gtex_science_paper$effsize > 0,])
}
```

```{r}
both_sex_tissue_names<- c(male_tissues_names, female_tissues_names)

geneset<- data.frame(id =both_sex_tissue_names, name= both_sex_tissue_names, genes= I(rep(list(c("a","b")), 88)))
sex_genes <- both_sex_tissue_names
for (i in 1:length(both_sex_tissue_names)){
  data <- get(both_sex_tissue_names[i])
  trim_names<- strtrim(data[,1],15)
  assign( sex_genes[i], trim_names)
  
}

sex_gene_set <- list(M_Adipose_Subcutaneous= M_Adipose_Subcutaneous, M_Adipose_Visceral_Omentum= M_Adipose_Visceral_Omentum)
```

```{r}
for (j in 1:88){
  sex_gene_set[[j]]<- get(sex_genes[j])
}
names(sex_gene_set) <- sex_genes
```


look at core genes 

```{r}
files <- list.files(path = "~/results/alpaca/alpaca_core_gene_lists")
```

remove the underscores
```{r}
tissues <- str_split(files, "_female", simplify = TRUE)[, 1]
tissues <- str_split(tissues, "_male", simplify = TRUE)[, 1]
tissues <- unique(tissues)
```

male core genes
```{r}
core_gene_list <- list()
for( i in 1:length(tissues)){
  print(tissues[i])
  male_core_genes <- readRDS(paste("~/results/alpaca/alpaca_core_gene_lists/", tissues[i],"_male_core_genes.rds", sep=""))
  core_gene_list[[i]] <- unique(male_core_genes$symbol)
}
```

```{r}
female_core_gene_list <- list()
for( i in 1:length(tissues)){
  print(tissues[i])
  female_core_genes <- readRDS(paste("~/results/alpaca/alpaca_core_gene_lists/", tissues[i],"_female_core_genes.rds", sep=""))
  female_core_gene_list[[i]] <- unique(female_core_genes$symbol)
}
```

```{r}
names(female_core_gene_list)<- tissues
names(core_gene_list)<- tissues
```



check the drug genes
```{r}
drug_genes <- c("ADH1A"	,"ADH1B", "ADH1C", "ADH4",	"ADH5",	"ADH6",	"ADH7",	
                "ALDH1A3",	"ALDH3A1",	"ALDH3B1", "ALDH3B2", "AOX1", "CYP1A2",
                "CYP2A13", "CYP2A6", "CYP2A7",	"CYP2B6",	"CYP2C18",	"CYP2C19",
                "CYP2C8",	"CYP2C9",	"CYP2D6",	"CYP2E1",	"CYP3A4",	"CYP3A43",	"CYP3A5",
                "CYP3A7",	"FMO1",	"FMO2",	"FMO3", "FMO4",	"FMO5",  "GSTA1", "GSTA2", 	"GSTA3",	"GSTA4",
                "GSTA5",	"GSTK1",	"GSTM1",	"GSTM2",	"GSTM3",	"GSTM4",	"GSTM5",	"GSTO1",
                "GSTO2",	"GSTP1",	"GSTT1",	"GSTT2",	"GSTZ1",	"MAOA",	"MAOB",	"MGST1",
                "MGST2",	"MGST3",	"UGT1A1",	"UGT1A10",	"UGT1A3",	"UGT1A4",
                "UGT1A5",	"UGT1A6",	"UGT1A7",	"UGT1A8", "UGT1A9",	"UGT2A1",	"UGT2A3", "UGT2B10",
                "UGT2B11",	"UGT2B15",	"UGT2B17",	"UGT2B28",	"UGT2B4",	"UGT2B7")
```


```{r}
drug_genes_v2 <- gprofiler2::gconvert(drug_genes, mthreshold= 1)
```

```{r}
table(drug_genes_v2$target %in% sex_gene_set$F_Liver)
```

```{r}
table(drug_genes_v2$target %in% sex_gene_set$M_Liver)
```

```{r}
drug_genes_v2$name[drug_genes_v2$target %in% sex_gene_set$F_Liver]
```

```{r}
drug_genes_v2$name[drug_genes_v2$target %in% sex_gene_set$M_Liver]
```

Permutation testing
of sex bias genes of drug metabolism genes
```{r}
network_genes <- gprofiler2::gconvert(colnames(Liver_F_panda@regNet), mthreshold= 1)
```


```{r}
drug_enz_gene_test1 <- drug_enz_gene_test(count = 8, all_drug_targets_list = network_genes$target,
                        core_list = c(sex_gene_set$F_Liver, sex_gene_set$M_Liver), seed = 101)
drug_enz_gene_test1 
```

```{r}
drug_enz_gene_test2 <- drug_enz_gene_test(count = 3, all_drug_targets_list = network_genes$target,
                        core_list = sex_gene_set$F_Liver, seed = 101)
drug_enz_gene_test2 
```

```{r}
drug_enz_gene_test3 <- drug_enz_gene_test(count = 5, all_drug_targets_list = network_genes$target,
                        core_list = sex_gene_set$M_Liver, seed = 101)
drug_enz_gene_test3
```
```{r}
ggarrange(drug_enz_gene_test1[[2]], drug_enz_gene_test2[[2]], drug_enz_gene_test3[[2]], nrow=2, ncol= 2)
ggsave("~/results/FARES_plots_drug_enzymes_expression_histograms.png", height=15, width = 10)
```




Permutation testing
Both male and female -core genes 
```{r}
drug_genes_v2$name[drug_genes_v2$name %in% core_gene_list$Liver]
```

```{r}
drug_genes_v2$name[drug_genes_v2$name %in% female_core_gene_list$Liver]
```

```{r}
#function defaults for this test 
drug_enz_gene_test4 <- drug_enz_gene_test()
drug_enz_gene_test4
```
male core genes 
```{r}
drug_enz_gene_test5 <- drug_enz_gene_test(count = 4, all_drug_targets_list = colnames(Liver_F_panda@regNet),
                        core_list = core_gene_list$Liver, seed = 101)
drug_enz_gene_test5
```
female core genes
```{r}
drug_enz_gene_test6 <- drug_enz_gene_test(count = 0, all_drug_targets_list = colnames(Liver_F_panda@regNet), 
                        core_list = female_core_gene_list$Liver, seed = 101)
drug_enz_gene_test6
```

```{r}
ggarrange(drug_enz_gene_test4[[2]], drug_enz_gene_test5[[2]], drug_enz_gene_test6[[2]], nrow =2, ncol = 2)
ggsave("~/results/FARES_plots_drug_enzymes_core_histograms.png", height=15, width = 10)
```


# END
    Location of final scripts:
    src
    Location of data produced:
    ~/results/FARES_plots
    Dates when operations were done:
    230418
    
## Versions
```{r}
sessionInfo()
```



